---
author: "Abigail Tyrell"
date: "`r format(Sys.time(), '%d %b %Y')`"
output: html_document
params:
  State:
    label: "State:"
    value: CONNECTICUT
    input: select
    choices: [CONNECTICUT, DELAWARE, MAINE, MARYLAND, MASSACHUSETTS, NEW HAMPSHIRE, NEW JERSEY, NEW YORK, RHODE ISLAND, VIRGINIA]
title: "Recreational fishing effort in`r params$state`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# 
```{r data cleaning, echo=FALSE,message=FALSE,warning=FALSE}
library(tidyverse)
#Load the required dataset from local csv
rec_effort <- read.csv(
  here::here("Recreational","mrip_effort_1981_2020_rec_trips_MA_NE.csv"), 
  stringsAsFactors = T, 
  fileEncoding = 'UTF-8-BOM')
#levels(rec_effort$State)


```
## Visualizing Recreational fishing effort

These records are Recreational Fisheries Statistics sourced from the [ Marine Recreational Information Program (MRIP) ](https://www.st.nmfs.noaa.gov/recreational-fisheries/data-and-documentation/queries/index) via NOAA Office of Science and Technology and are inclusive of all north and mid-Atlantic states. The records span from `r min(rec_effort$Year )` to `r max(rec_effort$Year) ` and are specific to the mode and location of effort. Effort is quantified as angler trips. Mode is divided into `r stringr::str_to_sentence(levels(rec_effort$Fishing.Mode))`.




```{r effort by single state, echo=FALSE, message=FALSE,warning=FALSE}
by_state_mode <- rec_effort %>% 
  dplyr::group_by(State, Fishing.Mode, Year) %>% 
  dplyr::summarise (total_trips = sum(Angler.Trips))

ggplot(by_state_mode,
       aes(x = Year,
           y = total_trips))+
  geom_line()+
  facet_grid(cols = vars(State), rows = vars(Fishing.Mode))+
  scale_y_continuous(labels = scales::comma)+
  ylab("Annual trips")

```

```{r effort by location, echo=FALSE, message=FALSE}

by_location <- rec_effort %>% 
  dplyr::group_by(State, Year, Fishing.Area) %>% 
  dplyr::summarise(Total_effort = sum(Angler.Trips))

ggplot(by_location,
       aes(x = Year,
           y = Total_effort))+
  geom_line()+
  facet_grid(cols = vars(State), 
             rows = vars(Fishing.Area))+
  scale_y_continuous(labels = scales::comma)+
  ylab("Annual trips")

```

```{r summary by state, echo=FALSE, message=FALSE,error=FALSE}
#group effort by state mode and year for first summary

by_state <- rec_effort %>% 
  dplyr::group_by(State, Year) %>%
  dplyr::summarise(total_trips= sum(Angler.Trips))

by_state %>% 
  ggplot2::ggplot(aes(x=Year, 
                      y=total_trips/100000))+
  tidyquant::geom_ma(aes(color = State, 
                         linetype = State), 
                     size=1.3,
                     n = 5)+
  ggtitle("Recreational Fishing effort by state ", 
          subtitle = "5 year rolling average")+
  xlab("Year")+
  ylab("Annual recreational trips in 100,000 units")+
  theme_minimal()

```

## Facet everything
```{r, echo=FALSE, message=FALSE,warning=FALSE}
all <- rec_effort %>% 
  dplyr::group_by(State, Fishing.Mode, Fishing.Area, Year) %>% 
  dplyr::summarise (total_trips = sum(Angler.Trips))

ggplot(all, 
       aes(x = Year,
           y = total_trips/10^5,
           color = State))+
  geom_line()+
  facet_grid(rows = vars(Fishing.Mode),
             cols = vars(Fishing.Area),
             scales = "free_y")+
  theme_bw()+
  ylab("Total trips / 100,000")


```


