---
title: "Recreational fishing effort in mid-atlantic and northeast states"
author: "Ricky Tabandera"
date: "`r format(Sys.time(), '%d %b %Y')`"
output: html_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# 
```{r data cleaning, echo=FALSE,message=FALSE,warning=FALSE}
library(tidyverse)
ggplot2::theme_set(theme_minimal())
#Load the required dataset from local csv
rec_effort<-read.csv(here::here("Recreational","mrip_effort_1981_2020_rec_trips_MA_NE.csv"), stringsAsFactors = T, fileEncoding = 'UTF-8-BOM')
#levels(rec_effort$State)


```
## Visualizing Recreational fishing effort

These records are Recreational Fisheries Statistics sourced from the [ Marine Recreational Information Program (MRIP) ](https://www.st.nmfs.noaa.gov/recreational-fisheries/data-and-documentation/queries/index) via NOAA Office of Science and Technology and are inclusive of all north and mid-Atlantic states. The records span from `r min(rec_effort$Year )` to `r max(rec_effort$Year) ` and are specific to the mode and location of effort. Effort is quantified as angler trips. Mode is divided into `r stringr::str_to_sentence(levels(rec_effort$Fishing.Mode))`.




```{r effort by  state, echo=FALSE, message=FALSE,warning=FALSE,fig.width=11,fig.height =8.5}
by_state_mode<-rec_effort %>% dplyr::group_by(State, Fishing.Mode, Year) %>% dplyr::summarise (total_trips= sum(Angler.Trips))

by_state_mode$Fishing.Mode<- stringr::str_to_sentence(by_state_mode$Fishing.Mode)

by_state_mode %>% ggplot2::ggplot(aes(x=Year, y=(total_trips/100000)))+
  ggplot2::geom_line(aes(color = Fishing.Mode),size=1.3)+
  ggtitle("Fishing effort by mode across states")+
  xlab("Year")+
  ylab("Annual recreational trips in 100,000 units")+
  facet_grid(cols = vars(State), rows = vars(Fishing.Mode))+
  theme(axis.text.x = element_text(angle = 90), strip.text.x = element_text(size = 7,angle = 30))


```

```{r effort by location, echo=FALSE, message=FALSE,warning=FALSE,fig.width=11,fig.height =8.5}

by_location<- rec_effort %>% dplyr::group_by(Year, Fishing.Area, State) %>% dplyr::summarise(Total_effort=sum(Angler.Trips))

by_location %>% ggplot2::ggplot(aes(x=Year, y=(Total_effort/100000)))+
  tidyquant::geom_ma(aes(color = Fishing.Area),size=1.3,n = 5)+
  ggtitle("Recreational Fishing effort by location ", subtitle = " 4 year rolling average")+
  xlab("Year")+ylab("Annual recreational trips in 100,000 units")+theme_minimal()+
  facet_grid(rows = vars(Fishing.Area), cols = vars(State) )+
  theme(axis.text.x = element_text(angle = 90), strip.text.x = element_text(size = 7,angle = 30))




```

```{r summary by state, echo=FALSE, message=FALSE,error=FALSE,fig.width=11,fig.height =8.5}
#group effort by state mode and year for first summary

by_state<-rec_effort %>% dplyr::group_by(State, Year) %>% dplyr::summarise (total_trips= sum(Angler.Trips))

by_state %>% ggplot2::ggplot(aes(x=Year, y=(total_trips/100000)))+
               tidyquant::geom_ma(aes(color = State,linetype = State),size=1.3, n=5)+
               ggtitle("Recreational Fishing effort by state ", subtitle = "5 year rolling average")+xlab("Year")+
               ylab("Annual recreational trips in 100,000 units")
```

