---
title: "recreational catch 1950-2019"
author: "Ricky Tabandera"
date: "11/18/2020
"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


```



```{r cars}

library("tidyverse")
rec_landings_1950_2019<-read.csv("foss_landings_REC_1950_2019_MA_NE.csv", stringsAsFactors = T)
fish_list<-unique(rec_landings_1950_2019$NMFS.Name)
stock_list_all_strata <- read.csv("https://raw.githubusercontent.com/NOAA-EDAB/ECSA/master/data/stock_list.csv")
stock_list<-unique(stock_list_all_strata$common_name)
which(fish_list == stock_list)
rec_by_sp<-rec_landings_1950_2019 %>% group_by(NMFS.Name)





```

## Recreational catches time serise 
This data was sourced from [FOSS - Fisheries One Stop Shop](https://foss.nmfs.noaa.gov/apexfoss/f?p=215:200:4615327020711::NO:::) and is inclusive of `r min(rec_landings_1950_2019$Year) ` to `r max(rec_landings_1950_2019$Year)`. The entire data set contains `r length(unique(rec_landings_1950_2019$NMFS.Name))` species. 

You can also embed plots, for example:

```{r pressure, echo=FALSE}




```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
