---
title: "recreational catch 1950-2019"
author: "Ricky Tabandera"
date: "11/18/2020
"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


```



```{r }

library("tidyverse")


#read in recreational landings data
rec_landings_1950_2019<-readr::read_csv("foss_landings_REC_1950_2019_MA_NE.csv")
#extract the names of species
fish_list<-unique(rec_landings_1950_2019$`NMFS Name`)
#load in the stocks managed in NEFSC
stock_list_all_strata <- read.csv("https://raw.githubusercontent.com/NOAA-EDAB/ECSA/master/data/stock_list.csv")
#extract out the names
stock_list<-unique(stock_list_all_strata$common_name)
stock_list<-stringr::str_to_lower(stock_list)
#
rec_by_sp<-rec_landings_1950_2019 %>% group_by(`NMFS Name`,Year) %>% summarise(mean= mean(Pounds))
rec_by_sp$`NMFS Name`<-stringr::str_to_lower(rec_by_sp$`NMFS Name`)
cod<-rec_by_sp$`NMFS Name`[913]
t2<-rec_w_comma$`NMFS Name` %>% str_split("\\,\\s")
str_split(cod,  "\\,\\s")

rec_w_comma<-rec_by_sp[str_detect(rec_by_sp$`NMFS Name`, "\\,\\s"),]

str_glue_data()

if(stringr::str_detect(cod, "\\,\\s")== T){
  print("has comma")
}





ggplot2::ggplot(rec_by_sp, aes(x=Pounds,))

length(unique(rec_landings_1950_2019$`NMFS Name`))

```

## Recreational catches time serise 
This data was sourced from [FOSS - Fisheries One Stop Shop](https://foss.nmfs.noaa.gov/apexfoss/f?p=215:200:4615327020711::NO:::) and is inclusive of `r min(rec_landings_1950_2019$Year) ` to `r max(rec_landings_1950_2019$Year)`. The entire data set contains `r length(unique(rec_landings_1950_2019$`NMFS Name`))` species. 

You can also embed plots, for example:

```{r pressure, echo=FALSE}




```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
