---
title: "Recreational fishing effort in`r params$state`"
author: "Abigail Tyrell"
date: "`r format(Sys.time(), '%d %b %Y')`"
output: html_document
params:
  State:
    label: 'State:'
    value: CONNECTICUT
    input: select
    choices:
    - CONNECTICUT
    - DELAWARE
    - MAINE
    - MARYLAND
    - MASSACHUSETTS
    - NEW HAMPSHIRE
    - NEW JERSEY
    - NEW YORK
    - RHODE ISLAND
    - VIRGINIA
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# 
```{r data cleaning, echo=FALSE,message=FALSE,warning=FALSE}
library(tidyverse)
#Load the required dataset from local csv
rec_effort<-read.csv(here::here("Recreational","mrip_effort_1981_2020_rec_trips_MA_NE.csv"), stringsAsFactors = T, fileEncoding = 'UTF-8-BOM')
#levels(rec_effort$State)


```
## Visualizing Recreational fishing effort

These records are Recreational Fisheries Statistics sourced from the [ Marine Recreational Information Program (MRIP) ](https://www.st.nmfs.noaa.gov/recreational-fisheries/data-and-documentation/queries/index) via NOAA Office of Science and Technology and are inclusive of all north and mid-Atlantic states. The records span from `r min(rec_effort$Year )` to `r max(rec_effort$Year) ` and are specific to the mode and location of effort. Effort is quantified as angler trips. Mode is divided into `r stringr::str_to_sentence(levels(rec_effort$Fishing.Mode))`.




```{r effort by single state, echo=FALSE, message=FALSE,warning=FALSE}
by_state_mode <- rec_effort %>% 
  dplyr::group_by(State, Fishing.Mode, Year) %>% 
  dplyr::summarise (total_trips= sum(Angler.Trips))

selected_state <- by_state_mode %>% 
  dplyr::filter(State == params$State)

selected_state$Fishing.Mode <- stringr::str_to_sentence(selected_state$Fishing.Mode)

selected_state %>% 
  ggplot2::ggplot(aes(x=Year, 
                      y=(total_trips/100000)))+ 
  ggplot2::geom_line(aes(color = Fishing.Mode, 
                         linetype = Fishing.Mode),
                     size=1.3)+
  ggtitle(params$State)+
  xlab("Year")+
  ylab("Annual recreational trips in 100,000 units")+
  theme_minimal()


```

```{r effort by location, echo=FALSE, message=FALSE}

by_location<- rec_effort %>% 
  dplyr::group_by(Year, Fishing.Area) %>%
  dplyr::summarise(Total_effort= sum(Angler.Trips))

by_location %>%
  ggplot2::ggplot(aes(x=Year, 
                      y = Total_effort/100000,
                      group = Fishing.Area))+ 
  ggplot2::geom_line()+
  tidyquant::geom_ma(aes(color = Fishing.Area,
                         lty = Fishing.Area),
                     n = 5,
                     lwd = 1.3)+
  ggtitle("Recreational Fishing effort by location ", 
          subtitle = "5 year rolling average")+
  xlab("Year")+
  ylab("Annual recreational trips in 100,000 units")+
  theme_minimal()

  # + facet_grid(rows = vars(Fishing.Area))




```

```{r summary by state, echo=FALSE, message=FALSE,error=FALSE}
#group effort by state mode and year for first summary

by_state<-rec_effort %>% dplyr::group_by(State, Year) %>% dplyr::summarise (total_trips= sum(Angler.Trips))

by_state %>% ggplot2::ggplot(aes(x=Year, y=(zoo::rollmean(total_trips/100000,4, na.pad=TRUE))))+ggplot2::geom_line(aes(color = State, linetype = State), size=1.3)+ggtitle("Recreational Fishing effort by state ", subtitle = " 4 year rolling average")+xlab("Year")+ylab("Annual recreational trips in 100,000 units")+theme_minimal()
```

