
### Commercial fishing gear 

Ricky Tabandera

Fishing gear can range widely in their catch ability, selectivity, and effective areas of operations. Using `comlandr` to extract reported gear used in the many commercial fisheries in the northeast and any trends and/or shifts that may have occurred as one potential explanatory factor on changes in fisheries landings/revenue.  



```{r com-gear-setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(comlandr)
library(dbutils)
library(remotes)
suppressPackageStartupMessages(library(tidyverse))
library(rio)

```



```{r gear-wrangling}

com_gear<-rio::import(here::here("com_landings_test.rdata"))
sp_group<-rio::import(here::here("data", "species_groupings_V2.csv"))

sp_group<-sp_group %>% filter(!is.na(SVSPP)) %>% select(COMNAME, SVSPP, ITISSPP, NESPP3 , SCINAME) 
sp_group<-sp_group  %>%  mutate_at(c("COMNAME", "SCINAME"), .funs = str_to_sentence)



key_join<-dplyr::left_join(com_gear,sp_group,by="NESPP3" )

```

### Gear usage by year

Each landing record included gear type used, size class of gear, ecological production unit, species landed, weight of catch

```{r, gear-raw}

key_join <-key_join %>% drop_na()
if(nrow(key_join)>6){
  key_join %>% 
  group_by(YEAR, GEAR) %>%
  dplyr::summarise(gear.count=length(GEAR))%>%
  ggplot2::ggplot(aes(x= YEAR,y= gear.count,color=GEAR))+
  ggplot2::geom_point()+
  geom_line()+
  nmfspalette::scale_color_nmfs(palette = "regional web")+
  theme(axis.text.x = element_text(angle = 90))+
  xlab("Year")+
  ylab("# gear type records")
}else("No Data")


```
#### Gear diversity all species 
```{r, gear-diversity}

gear.diversity<-key_join %>% group_by(YEAR, EPU, GEAR) %>% summarise(gear.count=(length(GEAR))) 

if(nrow(gear.diversity)>5){
  gear.diversity<-gear.diversity %>%
  group_by(YEAR,EPU) %>%
  mutate( gear.prop= (gear.count/sum(gear.count)))


gear.H<-gear.diversity %>% group_by(YEAR, EPU) %>% dplyr::summarise(shannon= -sum(gear.prop*log(gear.prop))) 

gear.H %>% ggplot(aes(x=YEAR, y= shannon, color=EPU))+
  geom_point()+
  geom_line()+
  nmfspalette::scale_color_nmfs(palette = "regional web")+
  xlab("Year")+
  ylab("gear diversity index (H')")



}else("No data")
```



#### Commercial gear types used to land 'r params$species'

```{r, gear-by-species}

gear.spp<-key_join%>% filter(COMNAME == params$species)

if(nrow(gear.spp)>5){
  gear.spp %>% 
  group_by(YEAR, EPU) %>% 
  mutate(gear.count=length(GEAR)) %>%  
  ggplot2::ggplot(aes(x=YEAR, y=gear.count, fill=GEAR))+
  ggplot2::geom_bar(stat="identity")+
  theme(axis.text.x = element_text(angle = 90))+
  labs(fill="Gear Type")+
  nmfspalette::scale_fill_nmfs(palette = "regional web")+
  xlab("Year")+
  ylab("# of landings")+
  ggtitle(label = params$species)
  # scale_fill_discrete(name = "Gear Type")
  }else("No data")

```
#### Diversity index of gear type across EPUs




#### Diversity index of gear used to land 'rparams$species' across EPUs

```{r, gear-diversity-by-species}


spp.diversity<-gear.spp %>% group_by(YEAR, GEAR, EPU) %>% summarise(gear.count=(length(GEAR))) 

spp.diversity<-spp.diversity %>%
  group_by(YEAR, EPU) %>%
  mutate( gear.prop= (gear.count/sum(gear.count)))


spp.H<-spp.diversity %>% group_by(YEAR, EPU) %>% dplyr::summarise(shannon= -sum(gear.prop*log(gear.prop))) 
if(nrow(spp.H)>5){
  spp.H %>% 
  ggplot(aes(x=YEAR,y=shannon, color=EPU))+
  geom_point()+
  geom_line()+
  nmfspalette::scale_color_nmfs(palette = "regional web")+
  xlab("Year")+
  ylab("gear diversity index (H')")+
  ggtitle(label=params$species)
}else("No data")
```

