# Habitat information

```{r, habitat_functions, include = FALSE, cache = TRUE}
source(here::here("R/full_report_functions", "get_latlong.R"))
source(here::here("R/full_report_functions", "get_survdat_info.R"))
source(here::here("R/full_report_functions", "map_strata_ecsa.R"))
```

```{r, habitat_data, include = FALSE}
latlong_data <- params$latlong_data[params$latlong_data$Species == species, ]
 
survey_data <- params$survey_data[params$survey_data$Species == species, ]
```

## Map of distribution
Strata maps were pulled and compiled using code from [NOAA/EDAB ECSA](https://github.com/NOAA-EDAB/ECSA). Please note, only fall and spring strata are shown on the map.
```{r, ecsa_map}
map_strata_ecsa(data = latlong_data,
                species_name = species)
```

## Latitude and longitude ranges
Latitude and longitude ranges were calculated from NOAA/EDAB ECSA [seasonal species strata](https://github.com/NOAA-EDAB/ECSA/blob/master/data/seasonal_stock_strata.csv) and [Bottom Trawl Survey (BTS) shapefiles](https://github.com/NOAA-EDAB/ECSA/tree/master/data/strata_shapefiles). The coordinate system is WGS84.
```{r, latlong}
data <- get_latlong(species, latlong_data, shapefile = params$shape) %>%
  character_to_factor()

make_html_table(data, col_names = colnames(data))
```

## Temperature

Surface and bottom temperature data were pulled from ``r params$survey_source``. 

### Figures

Separate geom_gls() functions were fit for fall and spring measurements; trend lines are only shown when the trend was statistically significant, so some plots may have fewer than two trend lines. Fall has solid trend lines, and spring has dashed trend lines. Please note, sometimes the survey observed a small number of fish outside of the defined stock area.
```{r, temp}
generate_plot(survey_data,
              variable = "SURFTEMP", 
              ytitle = "Surface temperature")

generate_plot(survey_data,
              variable = "BOTTEMP", 
              ytitle = "Bottom temperature")
```

### Summary
```{r, temp_summary}
generate_table(survey_data,
              variable = "SURFTEMP", 
              cap = "Surface temperature")

generate_table(survey_data,
              variable = "BOTTEMP", 
              cap = "Bottom temperature")
```

### Data
```{r, temp_data}
data <- survey_data %>% 
                dplyr::select(YEAR, SEASON, Region, date, 
                              fish_id, SURFTEMP, BOTTEMP) %>%
                dplyr::filter(SURFTEMP > 0 | BOTTEMP > 0) %>%
                dplyr::group_by(YEAR, SEASON, Region, date, fish_id) %>%
                dplyr::distinct() %>% # remove repeated row info
                dplyr::summarise(day_mean_surf = mean(SURFTEMP),
                                 day_mean_bot = mean(BOTTEMP)) %>% # mean by day
                dplyr::ungroup() %>%
                dplyr::group_by(YEAR, SEASON, Region) %>%
                dplyr::summarise(Mean_surface_temperature = mean(day_mean_surf) %>%
                                   round(digits = 2),
                                 Mean_bottom_temperature = mean(day_mean_bot) %>%
                                   round(digits = 2)) %>%
  character_to_factor() # mean by season-year

make_html_table(data, col_names = c("Year", "Season", "Region", 
                       "Mean surface temperature", "Mean bottom temperature"))
```

## Salinity

Surface and bottom salinity data were pulled from ``r params$survey_source``. 

### Figures

Separate geom_gls() functions were fit for fall and spring measurements; trend lines are only shown when the trend was statistically significant, so some plots may have fewer than two trend lines. Fall has solid trend lines, and spring has dashed trend lines. Please note, sometimes the survey observed a small number of fish outside of the defined stock area.
```{r, surfsalin}
generate_plot(survey_data,
              variable = "SURFSALIN", 
              ytitle = "Surface salinity")

generate_plot(survey_data,
              variable = "BOTSALIN", 
              ytitle = "Bottom salinity")
```

### Summary
```{r, salin_summary}
generate_table(survey_data,
              variable = "SURFSALIN", 
              cap = "Surface salinity")

generate_table(survey_data,
              variable = "BOTSALIN", 
              cap = "Bottom salinity")
```

### Data
```{r, salin_data}
data <- survey_data %>% 
                dplyr::select(YEAR, SEASON, Region, date, 
                              fish_id, SURFSALIN, BOTSALIN) %>%
                dplyr::filter(SURFSALIN > 0 | BOTSALIN > 0) %>%
                dplyr::group_by(YEAR, SEASON, Region, date, fish_id) %>%
                dplyr::distinct() %>% # remove repeated row info
                dplyr::summarise(day_mean_surf = mean(SURFSALIN),
                                 day_mean_bot = mean(BOTSALIN)) %>% # mean by day
                dplyr::ungroup() %>%
                dplyr::group_by(YEAR, SEASON, Region) %>%
                dplyr::summarise(Mean_surface_salinity = mean(day_mean_surf) %>%
                                   round(digits = 2),
                                 Mean_bottom_salinity = mean(day_mean_bot) %>%
                                   round(digits = 2)) %>%
  character_to_factor() # mean by season-year

make_html_table(data, col_names = c("Year", "Season", "Region", 
                       "Mean surface salinity", "Mean bottom salinity"))
```
