## Diet

```{r, diet_functions, cache = params$cache}
source(here::here("R/full_report_functions", "get_diet.R"))
```

```{r, diet_raw_data}
diet_data <- params$diet_data[params$diet_data$Species == species, ]
```

Diet data were compiled from [existing data](https://github.com/Laurels1/Condition/blob/master/data/allfh.RData). For analysis, all geographic samples were grouped by season, year, and region, and only year-season-region combinations with more than 20 predators sampled were considered. Prey items that made up more than 5% of the predator's diet in at least one year-season-region were identified to the broad category level; all other prey are grouped into the "other" category.

### Figure
```{r, diet}
get_diet_plot(data = diet_data)
```

### Summary
```{r, diet_summary}
get_diet_table(data = diet_data)
```

### Data
```{r, diet_data}
diet_data2 <- diet_data %>%
  dplyr::filter(pyamtw > 0) %>%

  # only look at season/year combinations with >20 predator samples
  dplyr::group_by(year, season, Region) %>%
  dplyr::mutate(n_predators = fish_id %>% unique() %>% length()) %>%
  dplyr::filter(n_predators > 20) %>%
  dplyr::group_by(year, season, Region, n_predators, gensci) %>%
  dplyr::summarise(total_weight = sum(pyamtw)) %>%
  dplyr::mutate(
    proportion = total_weight / sum(total_weight),
    proportion = proportion %>%
      round(digits = 3),
    total_weight = total_weight %>%
      round(digits = 2),
    gensci = gensci %>%
      stringr::str_replace("_", " ") %>%
      stringr::str_to_sentence()
  ) %>%
  character_to_factor()

make_html_table(diet_data2, col_names = c(
  "Year", "Season", "Region", "Number of fish",
  "Prey category", "Prey weight", "Prey proportion"
))
```
