## Age diversity

Ricky Tabandera

```{r setup_age_diversity,  cache = params$cache}
# knitr::opts_chunk$set(echo = TRUE)

suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(here))
suppressPackageStartupMessages(library(survdat))
suppressPackageStartupMessages(library(dbutils))
```


```{r data_wrangling_age, cache = params$cache, cache.lazy = FALSE}
# load survdat
survdata <- readRDS(here::here("data", "survdat_pull_bio.rds"))
# change from char to numeric
survdata <- survdata %>% mutate(SVSPP = as.numeric(SVSPP))
# load ECSA data with common names
stock_list_all_strata <- read.csv("https://raw.githubusercontent.com/NOAA-EDAB/ECSA/master/data/stock_list.csv")
stock_list_all_strata <- stock_list_all_strata %>% rename(SVSPP = svspp)
stock_list <- stock_list_all_strata %>% dplyr::distinct(SVSPP, .keep_all = T)
# link common names to  this dataset
survdata.w.codes <- inner_join(survdata, stock_list, by = "SVSPP")
# filter down to a single stock
selected.spp <- survdata.w.codes %>% filter(common_name %>%
  stringr::str_to_sentence() %>%
  stringr::str_replace("Goosefish", "Monkfish") == params$species)
```

Diversity in age measurements of a stock is a useful indicator of several factors relating to fishing pressure and recruitment. A decrease in diversity can be due to either truncation, the lack of older or younger ages. Diversity changes as a function of an increase of a single/few ages relative to the usual stock age structure or as more ages become less represented. Diagnostic plots of age are constructed below using fisheries independent data from `survdat`.

### Figures

#### Age diversity {-}

`r if(params$species == "Monkfish"){"The age determination method for monkfish has not been validated, and the anatomic structure used has changed through time. In addition these stocks display a high degree of sexual dimorphisms making fitting of single growth curves unreliable.  This results a high degree of uncertainty in assessing the age structure of the stock and the effects of fishing pressure."}`

```{r, age diversity calculations}

if (selected.spp$AGE %>% unique() %>% length() >= 3) {
  selected.age <- selected.spp %>% filter(!is.na(AGE))

  age.freq <- selected.age %>%
    group_by(YEAR, AGE) %>%
    summarise(age.n = n())
  age.freq <- age.freq %>%
    group_by(YEAR) %>%
    mutate(prop = (age.n / sum(age.n))) %>%
    mutate(prop.ln = (prop * log(prop)))


  age.freq <- age.freq %>%
    group_by(YEAR) %>%
    summarise(shanon.h = (-1 * (sum(prop.ln))))
} else {
  print("NOT ENOUGH DATA TO GENERATE METRIC")
  age.freq <- tibble::tibble() # make empty tibble for next logical test
}

if (nrow(age.freq) > 0) {
  fig <- age.freq %>%
    ggplot(aes(x = YEAR, y = shanon.h)) +
    geom_path(group = 1, size = 1.2, color = "blue") +
    geom_point(size = 3, shape = 23, fill = "Black") +
    xlab("Year") +
    ylab("Shannon diversity index (H)") +
    theme(axis.text.x = element_text(angle = 90)) +
    ggtitle("Age diversity of", subtitle = params$species)

  return(fig)
} else {
  print("NO DATA")
}
```

#### Density plots of age {-}

Age distribution across years of survey data of `r params$species`. These plots can help identify strong year classes of recruits and how these classes persist in the fishery.  

```{r age density, fig.height = 12}

if (selected.spp$AGE %>% unique() %>% length() >= 3) {
  fig <- selected.spp %>%
    drop_na(AGE) %>%
    group_by(YEAR) %>%
    ggplot(aes(x = AGE, y = YEAR, fill = YEAR %>% as.numeric())) +
    scale_fill_gradientn(
      colors = nmfspalette::nmfs_palette("regional web")(4),
      name = "Year"
    ) +
    ggridges::geom_density_ridges2() +
    scale_x_continuous(
      limits = c(0, (max(selected.spp$AGE, na.rm = TRUE))),
      breaks = seq(0, max(selected.spp$AGE, na.rm = TRUE), by = 5)
    )

  return(fig)
} else {
  print("NO DATA")
}
```
