---
title: "`r paste(params$species_ID, 'Preliminary ESP regressions')`"
author: "Abigail Tyrell"
date: "`r format(Sys.time(), '%d %b %Y')`"
site: bookdown::bookdown_site
output: 
  bookdown::gitbook:
    split_by: section
documentclass: book
github-repo: NOAA-EDAB/esp_data_aggregation
params:
  lag: 0 # lag for correlations
  remove_recent: FALSE
  cache: FALSE
  stock: 
  epu: MAB # epu options: GB, GOM, MAB, SS, All, OTHER, MA, NE, all 
  region: "" # the stock region
  path: "" # figure path
---

# `r params$stock` {-}
Stock region: `r params$region`

EPU: `r params$epu`

# Introduction

These are preliminary regressions that compare `r params$region` `r params$stock` catch, abundance, recruitment, and F to various indicators in the `r paste(params$epu, collapse = ", ")` Environmental Protection Units (EPUs) taken from the `ecodata` package. The indicators are lagged by `r params$lag` years.

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      message = FALSE,
                      warning = FALSE, 
                      cache = params$cache,
                      results = "asis",
                      fig.width = 8, 
                      fig.align = 'center',
                      fig.path = params$path)

library(ggplot2)
`%>%` <- dplyr::`%>%`

source(here::here("R/regressions", "indicator_functions_general.R"))

if(params$stock == "Monkfish"){
  stock <- assessmentdata::stockAssessmentData %>% 
  dplyr::filter(Species == "Goosefish",
                Region == params$region) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(max_year = max(AssessmentYear)) %>%
  dplyr::filter(AssessmentYear == max_year) %>%
  dplyr::rename(Time = Year)
} else {
  stock <- assessmentdata::stockAssessmentData %>% 
  dplyr::filter(Species == params$stock,
                Region == params$region) %>%
  dplyr::ungroup() %>%
  dplyr::mutate(max_year = max(AssessmentYear)) %>%
  dplyr::filter(AssessmentYear == max_year) %>%
  dplyr::rename(Time = Year)
}

if(params$remove_recent){
  max_year <- max(stock$Time)
  stock <- stock %>%
    dplyr::filter(Time <= (max_year - 10))
} 

output <- c()

```

# Regression analysis

All regressions are simple linear correlations assessed at the p < 0.5 level. Please note, due to the large number of indicators tested, a certain amount of statistically significant results are expected even if there are no underlying mechanistic connections. These correlations do not necessarily imply causation.

<!--chapter:end:index.Rmd-->

## Trends with time
```{r, time}
test <- stock %>%
  dplyr::mutate(Val = Time,
                Var = "Time") %>%
  dplyr::select(Time, Val, Var) %>%
  dplyr::distinct()
render_indicator(test)
```

<!--chapter:end:01-time-trends.Rmd-->

## Physical indicators

### Cold pool index
```{r, cold_pool}
test <- ecodata::cold_pool
render_indicator(test)
```

### Warm core rings
```{r, warm}
test <- ecodata::wcr
render_indicator(test)
```

### Marine heatwave index
```{r, heatwave}
test <- ecodata::heatwave
render_indicator(test)
```

### GLORYS bottom temperature
```{r, glorys}
test <- ecodata::bottom_temp_glorys
render_indicator(test)
```

### Long-term sea surface temperature
```{r, sst}
test <- ecodata::long_term_sst
render_indicator(test)
```

### Sea surface temperature anomaly
```{r, sst_anom}
test <- ecodata::seasonal_oisst_anom 
render_indicator(test)
```

### Stratification
```{r, strat}
test <- ecodata::stratification
render_indicator(test)
```

### Winter wind speed
```{r, winter_wind}
test <- ecodata::ne_wind %>%
  dplyr::filter(Var %>% stringr::str_detect("winter"))
render_indicator(test)
```

### Spring wind speed
```{r, spring_wind}
test <- ecodata::ne_wind %>%
  dplyr::filter(Var %>% stringr::str_detect("spring"))
render_indicator(test)
```

### Summer wind speed
```{r, summer_wind}
test <- ecodata::ne_wind %>%
  dplyr::filter(Var %>% stringr::str_detect("summer"))
render_indicator(test)
```

### Fall wind speed
```{r, fall_wind}
test <- ecodata::ne_wind %>%
  dplyr::filter(Var %>% stringr::str_detect("fall"))
render_indicator(test)
```

### Gulf Stream Index
```{r, gsi}
test <- ecodata::gsi %>%
  dplyr::mutate(Time = Time %>%
                  stringr::str_trunc(width = 4, ellipsis = "") %>%
                  as.numeric())
render_indicator(test)
```

### North Atlantic Oscillation
```{r, nao}
test <- ecodata::nao
render_indicator(test)
```

### Chlorophyll
```{r, chl}
test <- ecodata::chl_pp %>%
  dplyr::mutate(Var = Var %>%
                  stringr::str_replace_all("_", "\n") %>%
                  stringr::str_replace_all(" ", "\n")) %>%
  dplyr::filter(stringr::str_detect(Var, "ANNUAL")) %>% # only annual indices
  dplyr::mutate(Time = Time %>%
                  stringr::str_replace("A_", "") %>%
                  as.numeric())
render_indicator(test)

#data_prep(stock, test, lag = params$lag) %>% knitr::kable()
```

<!--chapter:end:02-physical-indicators.Rmd-->

## Trophic indicators

### Spring zooplankton abundance by species
```{r, zoo_oi_spring}
test <- ecodata::zoo_oi %>%
  dplyr::filter(Var %>% stringr::str_detect("SD") == FALSE,
                Var %>% stringr::str_detect("spring")) 
render_indicator(test)
```

### Fall zooplankton abundance by species
```{r, zoo_oi_fall}
test <- ecodata::zoo_oi %>%
  dplyr::filter(Var %>% stringr::str_detect("SD") == FALSE,
                Var %>% stringr::str_detect("fall")) 
render_indicator(test)
```

### Zooplankton abundance by group
```{r, zoo_strat_abun}
test <- ecodata::zoo_strat_abun 
render_indicator(test)
```

### Abundance of Calanus CV and adults
```{r, calanus}
test <- ecodata::CalanusStage %>%
  dplyr::rename(Time = Year) %>%
  dplyr::filter(Var == "Adt" | Var == "CV",
                Units == "No. per 100m^-3") %>%
  dplyr::group_by(Time, EPU, season) %>%
  dplyr::mutate(Val = sum(Value)) %>%
  dplyr::select(Time, Val, EPU, season) %>% 
  dplyr::distinct() %>%
  dplyr::mutate(Var = paste("Calanus CV and adult", season))
render_indicator(test)
```

### Zooplankton abundance anomaly
```{r, zoo_abund}
test <- ecodata::zoo_abund
render_indicator(test)
```

### Zooplankton diversity index
```{r, zoo_diversity}
test <- ecodata::zoo_diversity
render_indicator(test)
```

### Small/large copepod anomaly
```{r, zoo_sli_anom}
test <- ecodata::zoo_sli_anom
render_indicator(test)
```

### Ichthyoplankton diversity
```{r, ich}
test <- ecodata::ichthyo_diversity 
render_indicator(test)
```

```{r, ich_fall, eval = FALSE}
### Fall ichthyoplankton diversity
# seasonal data no longer in ecodata
test <- ecodata::ichthyo_diversity %>%
  dplyr::filter(Season == "Fall") 
render_indicator(test)
```

### Forage fish abundance
```{r, forage}
test <- ecodata::forage_anomaly
render_indicator(test)
```

### Species distribution
```{r, species_dist}
test <- ecodata::species_dist
render_indicator(test)
```

<!--chapter:end:03-trophic-indicators.Rmd-->

## Larvae and YOY indicators

### Recruitment
```{r, recruit}
test <- stock %>%
  dplyr::filter(Metric == "Recruitment") %>%
  dplyr::rename(Var = Metric,
                Val = Value)
if(nrow(test) > params$lag){
  render_indicator(test)
}
```

### Larval growth

<!--chapter:end:04-larvae-YOY.Rmd-->

## Juvenile indicators

### Length-age curves

### Condition

### CPUE

<!--chapter:end:05-juvenile.Rmd-->

## Adult indicators

### Abundance
```{r, ssb}
test <- stock %>%
  dplyr::filter(Metric == "Abundance") %>%
  dplyr::rename(Var = Metric,
                Val = Value)

if(nrow(test) > params$lag){
  render_indicator(test)
} else print("No data")
```

### Mean age of spawning stock

### Age distribution

### Length-age curves

### Condition
```{r, condition}
test <- read.csv("https://raw.githubusercontent.com/Laurels1/Condition/master/data/AnnualRelCond2018_MAB.csv") %>%
  dplyr::mutate(Species = Species %>%
                  stringr::str_to_sentence()) %>%
  dplyr::filter(Species == params$stock) %>%
  dplyr::rename(Val = MeanCond,
                Time = YEAR) %>%
  dplyr::mutate(Var = paste(sex, "condition"))

if(nrow(test) > 0){
  render_indicator(test)
} else print("No data")
```

### Stomach fullness
```{r, stomach}
test <- ecodata::stom_fullness %>% 
  dplyr::mutate(Var = Var %>%
                  stringr::str_to_sentence()) %>%
  dplyr::filter(Var == params$stock)

if(nrow(test) > 0){
  render_indicator(test)
} else print("No data")
```

### Center of gravity and area occupied

<!--chapter:end:06-adult.Rmd-->

## Socioeconomic indicators

### CPUE by catch strategy

### Recreational CPUE
```{r, rec_cpue}
numbers <- ecodata::recdat %>%
  dplyr::filter(Var == "Recreational Seafood") %>%
  dplyr::rename(Number = Value)

days <- ecodata::recdat %>%
  dplyr::filter(Var == "Recreational Effort") %>%
  dplyr::rename(Days = Value)

test <- dplyr::left_join(numbers, days, by = c("Time", "EPU")) %>%
  dplyr::mutate(Val = Number/Days,
                Var = paste("recreational CPUE", "number of fish caught per day fished (all species)", sep = "\n"))
render_indicator(test)
```

<!--chapter:end:07-socioeconomic.Rmd-->

# Summary of statistically significant indicators

```{r, output_summary}
source(here::here("R/full_report_functions", "make_html_table.R"))

summary_table <- function(x){
    DT::datatable(x,
                  rownames = FALSE,
                  filter = list(position = 'top', 
                                clear = FALSE),
                  options = list(search = list(regex = TRUE),
                                 paging = FALSE))
}

colnames(output) <- c("Response_variable", "Indicator", 
                                 "Number of data points", "Slope", 
                                 "P-value", "R2_adj")
```

## Abundance
```{r, eval = TRUE}
output %>% 
  tibble::as_tibble() %>%
  dplyr::filter(Response_variable == "Abundance",
                R2_adj < 1) %>%
  dplyr::select(-Response_variable) %>%
  summary_table()


```

## Recruitment
```{r, eval = TRUE}
output %>% 
  tibble::as_tibble() %>%
  dplyr::filter(Response_variable == "Recruitment",
                R2_adj != 1) %>%
  dplyr::select(-Response_variable) %>%
  summary_table()

```

## Catch
```{r, eval = TRUE}
output %>% 
  tibble::as_tibble() %>%
  dplyr::filter(Response_variable == "Catch",
                R2_adj != 1) %>%
  dplyr::select(-Response_variable) %>%
  summary_table()

```

## Fmort
```{r, eval = TRUE}
output %>% 
  tibble::as_tibble() %>%
  dplyr::filter(Response_variable == "Fmort",
                R2_adj != 1) %>%
  dplyr::select(-Response_variable) %>%
  summary_table()

```


<!--chapter:end:08-summary.Rmd-->

