--- 
title: "`r params$species_ID`"
author: "Abigail Tyrell"
date: "`r format(Sys.time(), '%d %b %Y')`"
site: bookdown::bookdown_site
output: 
  bookdown::gitbook:
    split_by: section
documentclass: book
github-repo: rtabandera/esp_data_aggregation

params: 
  species_ID: "Acadian redfish" # common name of species
  
  latlong_data: data # strata
  shape: shape # shapefile
  
  asmt_sum_data: data # bbmsy, ffmsy, assessment ratings
  asmt_sum_source: "assessmentdata::stockAssessmentSummary"

  diet_data: data # allfh diet data
  
  survey_data: data # survey data
  survey_source: "survdat"

  rec_data: data # recreational catch (MRIP)
  
  asmt_data: data # assessmentdata:stockAssessmentData, for recruitment and catch
  asmt_source: "assessmentdata::stockAssessmentData"
  
  cond_data: data # Laurel's condition data
  
  risk_data: data # relative risk rankings
  
  com_data: data # commercial catch
  
  swept_data: data # swept area estimates
  
  risk_year_hist_data: data # historical risk by year
  
  risk_year_value_data: data # value risk by year
  
  risk_species_data: data # species risk
---

# `r params$species` {-}

```{r, setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE, 
                      warning = FALSE, 
                      fig.width = 8, 
                      fig.align = 'center',
                      fig.cap = params$species
                      )

library(ggplot2)

`%>%` <- dplyr::`%>%`
```

```{r, universal_functions, include = FALSE}
source(here::here("R", "character_to_factor.R"))
source(here::here("R/full_report_functions", "make_html_table.R"))

source(here::here("R/full_report_functions/risk_functions", "plot_risk_by_year.R"))
source(here::here("R/full_report_functions/risk_functions", "plot_risk_within_species.R"))
```

```{r, universal_data, include = FALSE}
# data used in multiple chapters
species <- params$species_ID

risk_year_hist_data <- params$risk_year_hist_data
risk_year_hist_data <- risk_year_hist_data[risk_year_hist_data$Species == species, ]

risk_year_value_data <- params$risk_year_value_data
risk_year_value_data <- risk_year_value_data[risk_year_value_data$Species == species, ]

risk_species_data <- params$risk_species_data
risk_species_data <- risk_species_data[risk_species_data$Species == species, ]
```

This is a preliminary report of previously collected data. This report is pulling information on all Northeast `r species` stocks.

<!--chapter:end:index.Rmd-->

--- 
title: "ESP Methods"
author: "Abigail Tyrell"
date: "`r format(Sys.time(), '%d %b %Y')`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
github-repo: rtabandera/esp_data_aggregation
---

# Methods

## Stock identification
Northeast stocks were identified from NOAA/EDAB ECSA [seasonal species strata](https://github.com/NOAA-EDAB/ECSA/blob/master/data/seasonal_stock_strata.csv).

## Data collection and presentation

Data sources for each analysis are identified in the Results.

All continuous temporal data were plotted against time. If there were 30 or more years of data, a geom_gls regression line was fit (yellow = significant increase; purple = significant decrease; no line = no significant trend). If there were fewer than 30 years of data, no regression was fit.

### `assessmentdata` methods
Stock assessment and data quality information were compiled into a summary table.

B/Bmsy was classified as "DANGER" if it was below 1 and "GOOD" if it was above 1.

F/Fmsy was classified as "DANGER" if it was above 1 and "GOOD" if it was below 1.

### `survdat` methods
`survdat` data with zero abundance were not included in this analysis. Abundance and biomass were summed for each year and season. All other metrics were averaged for each year and season. The tables show summary statistics for the entire time series and for the most recent 5 years in the time series. 

## Risk assessment

### Risk across stocks

#### Suite of indicators
All stocks were ranked in order of increasing risk. The stock with the highest ranking is the stock determined to be at the highest risk. In this case, high risk has two meanings: (1) high importance (e.g., a stock with a high catch would have a high risk ranking for the catch indicator) or high vulnerability (e.g., a stock with low B/Bmsy would have a high risk ranking for the B/Bmsy indicator). The normalized rank was determined by dividing each stock's rank by the total number of stocks considered for that indicator. Stocks that were missing indicator measurements were assigned a normalized rank of 0.5.

#### Individual indicators
Risk was calculated over time for all indicators that were documented for five or more species in a given year. Risk was calculated as the average of the past 5 years, as a percent of the historical average. The normalized risk value was calculated as the normalized rank of this species compared to all other species in that year.

### Risk within stocks
The normalized risk value was calculated as the normalized rank of each yearly measurement compared to all other years.

<!--chapter:end:01-methods.Rmd-->

# Habitat information

<!--chapter:end:02-habitat-title-page.Rmd-->


```{r, geo_functions, include = FALSE}
source(here::here("R/full_report_functions", "get_latlong.R"))
source(here::here("R/full_report_functions", "map_strata_ecsa.R"))
```

```{r, geo_data, include = FALSE}
latlong_data <- params$latlong_data
latlong_data <- latlong_data[latlong_data$Species == species, ]
```

## Map of distribution
Strata maps were pulled and compiled using code from [NOAA/EDAB ECSA](https://github.com/NOAA-EDAB/ECSA). Please note, only fall and spring strata are shown on the map.
```{r, ecsa_map}
map_strata_ecsa(data = latlong_data,
                species_name = species)
```

## Latitude and longitude ranges
Latitude and longitude ranges were calculated from NOAA/EDAB ECSA [seasonal species strata](https://github.com/NOAA-EDAB/ECSA/blob/master/data/seasonal_stock_strata.csv) and [Bottom Trawl Survey (BTS) shapefiles](https://github.com/NOAA-EDAB/ECSA/tree/master/data/strata_shapefiles). The coordinate system is WGS84.
```{r, latlong}
data <- get_latlong(species, latlong_data, shapefile = params$shape) %>%
  character_to_factor()

make_html_table(data, col_names = colnames(data))
```

<!--chapter:end:03-latlong.Rmd-->


```{r, temp_functions, include = FALSE}
source(here::here("R/full_report_functions", "get_survdat_info.R"))
```

```{r, tempe_data, include = FALSE}
survey_data <- params$survey_data
survey_data <- survey_data[survey_data$Species == species, ]
```

## Temperature

Surface and bottom temperature data were pulled from ``r params$survey_source``. 

### Figures

Separate geom_gls() functions were fit for fall and spring measurements; trend lines are only shown when the trend was statistically significant, so some plots may have fewer than two trend lines. Fall has solid trend lines, and spring has dashed trend lines. Please note, sometimes the survey observed a small number of fish outside of the defined stock area.
```{r, temp}
generate_plot(survey_data,
              variable = "SURFTEMP", 
              ytitle = "Surface temperature")

generate_plot(survey_data,
              variable = "BOTTEMP", 
              ytitle = "Bottom temperature")
```

### Summary
```{r, temp_summary}
generate_table(survey_data,
              variable = "SURFTEMP", 
              cap = "Surface temperature")

generate_table(survey_data,
              variable = "BOTTEMP", 
              cap = "Bottom temperature")
```

### Data
```{r, temp_data}
data <- survey_data %>% 
                dplyr::select(YEAR, SEASON, Region, date, 
                              fish_id, SURFTEMP, BOTTEMP) %>%
                dplyr::filter(SURFTEMP > 0 | BOTTEMP > 0) %>%
                dplyr::group_by(YEAR, SEASON, Region, date, fish_id) %>%
                dplyr::distinct() %>% # remove repeated row info
                dplyr::summarise(day_mean_surf = mean(SURFTEMP),
                                 day_mean_bot = mean(BOTTEMP)) %>% # mean by day
                dplyr::ungroup() %>%
                dplyr::group_by(YEAR, SEASON, Region) %>%
                dplyr::summarise(Mean_surface_temperature = mean(day_mean_surf) %>%
                                   round(digits = 2),
                                 Mean_bottom_temperature = mean(day_mean_bot) %>%
                                   round(digits = 2)) %>%
  character_to_factor() # mean by season-year

make_html_table(data, col_names = c("Year", "Season", "Region", 
                       "Mean surface temperature", "Mean bottom temperature"))
```

<!--chapter:end:04-temperature.Rmd-->


```{r, sal_functions, include = FALSE}
#source(here::here("R/full_report_functions", "get_survdat_info.R"))
```

```{r,sal_data, include = FALSE}
#survey_data <- params$survey_data
#survey_data <- survey_data[survey_data$Species == species, ]
```

## Salinity

Surface and bottom salinity data were pulled from ``r params$survey_source``. 

### Figures

Separate geom_gls() functions were fit for fall and spring measurements; trend lines are only shown when the trend was statistically significant, so some plots may have fewer than two trend lines. Fall has solid trend lines, and spring has dashed trend lines. Please note, sometimes the survey observed a small number of fish outside of the defined stock area.
```{r, surfsalin}
generate_plot(survey_data,
              variable = "SURFSALIN", 
              ytitle = "Surface salinity")

generate_plot(survey_data,
              variable = "BOTSALIN", 
              ytitle = "Bottom salinity")
```

### Summary
```{r, salin_summary}
generate_table(survey_data,
              variable = "SURFSALIN", 
              cap = "Surface salinity")

generate_table(survey_data,
              variable = "BOTSALIN", 
              cap = "Bottom salinity")
```

### Data
```{r, salin_data}
data <- survey_data %>% 
                dplyr::select(YEAR, SEASON, Region, date, 
                              fish_id, SURFSALIN, BOTSALIN) %>%
                dplyr::filter(SURFSALIN > 0 | BOTSALIN > 0) %>%
                dplyr::group_by(YEAR, SEASON, Region, date, fish_id) %>%
                dplyr::distinct() %>% # remove repeated row info
                dplyr::summarise(day_mean_surf = mean(SURFSALIN),
                                 day_mean_bot = mean(BOTSALIN)) %>% # mean by day
                dplyr::ungroup() %>%
                dplyr::group_by(YEAR, SEASON, Region) %>%
                dplyr::summarise(Mean_surface_salinity = mean(day_mean_surf) %>%
                                   round(digits = 2),
                                 Mean_bottom_salinity = mean(day_mean_bot) %>%
                                   round(digits = 2)) %>%
  character_to_factor() # mean by season-year

make_html_table(data, col_names = c("Year", "Season", "Region", 
                       "Mean surface salinity", "Mean bottom salinity"))
```

<!--chapter:end:05-salinity.Rmd-->

# Biological information

<!--chapter:end:06-biological-title-page.Rmd-->

```{r, len_functions, include = FALSE}
source(here::here("R/full_report_functions", "get_length.R"))
```

```{r, len_data, include = FALSE}
# length data (survdat) already read in with habitat data
#survey_data <- params$survey_data
#survey_data <- survey_data[survey_data$Species == species, ]
```

## Length

Length data were pulled from ``r params$survey_source``. Only years with more than 10 fish lengths were considered for analysis. 

### Figures

Separate geom_gls() functions were fit for the minimum, mean, and maximum lengths; trend lines are only shown when the trend was statistically significant, so some plots may have fewer than three trend lines. Please note, sometimes the survey observed a small number of fish outside of the defined stock area.
```{r, length}
generate_len_plot(survey_data)
```

#### Risk {-}

See Methods for risk calculation details.
```{r}
indicators <- c("max_length_spring",
               "max_length_fall",
               "avg_length_spring",
               "avg_length_fall")
```

```{r, risk_stock, fig.height = length(indicators) + 2}
plot_risk_by_year(risk_year_hist_data,
                  indicator = indicators,
                  title = "Change compared to historical",
                  include_legend = "yes")
```

```{r, risk_stock2, fig.height = length(indicators) + 1}
plot_risk_by_year(risk_year_value_data,
                  indicator = indicators,
                  title = "Value compared to other stocks",
                  include_legend = "no")

plot_risk_by_stock(risk_species_data,
                  indicator = indicators,
                  include_legend = "no")
```

### Summary
```{r, length_summary}
generate_len_table(survey_data)
```

### Data
```{r, length_data}
data <- get_len_data2(survey_data) %>%
  character_to_factor()

make_html_table(data, col_names = c("Year", "Season", "Region", "Number of fish",
                       "Mean length", "Minimum length", "Maximum length"))
```

<!--chapter:end:07-length.Rmd-->

## von Bertalanffy growth curve

Ricky Tabandera

```{r setup_vonb, include=FALSE}
#install.packages("ggrepel")
#install.packages("ggpubr")
#devtools::install_github("james-thorson/FishLife")

suppressPackageStartupMessages(library(FishLife))
suppressPackageStartupMessages(library(tidyverse))
library(rfishbase)
library(ggpubr)

ricky_stock <- stringr::str_to_lower(species)

```



```{r data_wrangling}

#function to calculate the fitted growth function for von bertalanffy growth curve
vonb<-function(t,Linf,K,tzero){Latthisage<-Linf*(1-exp(-K*(t-tzero)))
  return(Latthisage)
}

#Pauly 1979 based method to estimate a default value for t0 as log (-t0) = -0.3922 - 0.2752 log Linf - 1.038 log K
#consistant way of generating t0  
tzero<-function(Linf,K){logtzero<-(-0.3922 - (0.2752*log(Linf) - 1.038*log(K)))
  t0<-exp(logtzero)
  return(t0)  
}  

stock_list_all_strata<- readr::read_csv("https://raw.githubusercontent.com/NOAA-EDAB/ECSA/master/data/stock_list.csv")
stock_list<-stock_list_all_strata %>% dplyr::distinct( common_name, .keep_all=TRUE)

fishlife_predictions_NE<-readr::read_csv(here::here("life history/fishlife","FishLife_extracted_data_11.16.20_northeast.csv"))
predictions_NE_spaced<-fishlife_predictions_NE %>% dplyr::mutate_all(funs(str_replace(., "\\_", " ")))
predictions_NE_spaced<-predictions_NE_spaced %>% dplyr::rename(sci_name = Species)
predictions_join <-dplyr::inner_join(stock_list, predictions_NE_spaced, by="sci_name" )

predictions_join<-predictions_join %>% 
  dplyr::filter(WORKED==TRUE)
#reformat these into numeric cols
predictions_join<-predictions_join %>% 
  dplyr::mutate(Loo=as.numeric(Loo), K=as.numeric(K))

#return the model output into original scale
predictions_join$Linf<-exp(predictions_join$Loo)
predictions_join$K<-exp(predictions_join$K)

predictions_join<-predictions_join %>% mutate(t0=tzero(Linf, K))
#write.csv(predictions_join, file = "predictions_join.csv")

#select out needed cols
prediction.select<-predictions_join %>% select(common_name, Linf,K,t0)

#catagorize the Linf into rough bins
prediction.select$body_cat<-cut(prediction.select$Linf, breaks=c(-Inf,50, 100, Inf), 
                                labels=c("small","med","large"))

#split into a set of lists for map use
flist<-split(prediction.select, f=prediction.select$common_name)

# vonb wants arguments in the order t,Linf,K,tzero)

spec.age.growth.list<-map(flist, ~vonb(seq(1:75),.x[[2]],.x[[3]],.x[[4]]))

# turn the list of lists into a single object
spec.age.growth<-do.call(cbind,spec.age.growth.list)
#turn this object into a df for manipulation
spec.age.growth.df<-data.frame(spec.age.growth)
# make an index of years int he dataset 

spec.age.growth.df$year<-seq(1:75)
#extract col names for pivoting 
fish.names<-colnames(spec.age.growth.df)
fish.names<-fish.names[fish.names != "year"]

#Pivot from wide format to long to make graphing easy

age.growth.curves<-spec.age.growth.df %>% pivot_longer(cols = fish.names,names_to = "common_name", values_to = "length" )
#remove the period in common name to enable the matching fuction in dplyr join
age.growth.curves$common_name<-str_replace(age.growth.curves$common_name,"\\."," " )

age.growth.join<-right_join(prediction.select,age.growth.curves, by= "common_name")

age.growth.join$common_name<-  str_to_lower(age.growth.join$common_name)
#recode black sea.bass to black sea bass inorder to be consistant with other datasets 

age.growth.join$common_name<- recode(age.growth.join$common_name, "black sea.bass" = "black sea bass")

#########################################################
#logical filtering of params for plotting

is_small<-FALSE
is_med<-FALSE
is_large<-FALSE

sizeclass<-age.growth.join %>% filter(common_name==ricky_stock)

if(nrow(sizeclass) > 0){
  if(sizeclass$body_cat=="small"){
  is_small<-TRUE}

if(sizeclass$body_cat=="med"){
  is_med<-TRUE}

if(sizeclass$body_cat=="large"){
  is_large<-TRUE}
} else missing <- TRUE
 
```

### Length at age growth curve

The predicted von Bertalanffy growth curve for NMFS managed fish species. Growth parameters of Linf (Length infinty), K (growth coefficient)
were estimated using [James Thorson FishLife R package](https://github.com/James-Thorson-NOAA/FishLife). This package uses a mix of previously published life history parameters extracted from fishbase and correlated higher taxon parameters to generate estimates for growth and population parameters. t0 (size at age 0) was estimated using the method described by Pauly 1979 for generating a consistent value for t0 from Linf and K. 

```{r single_growth_curve}
if(nrow(sizeclass) > 0){
  sizeclass %>%
  ggplot(aes(x=year, y=length))+
  geom_line(color= "blue",size=1.4)+
  xlab("Year")+
  ylab(" Total length (cm)")+
  labs(color="Common name of stock")+
  ggtitle(str_to_sentence(ricky_stock), subtitle = "Length at age")+
  theme_minimal()
} else print("NO DATA")

```

### Growth by size class

Fish were categorized into three groups with small bodied species reaching max size < 50 cm, medium 50-100 cm, and large >100 cm as rough divisions for comparisons of growth among similar body sizes.

```{r small_growth_curves, eval=is_small}

if(nrow(sizeclass) > 0){
  age.growth.join %>%
  filter(body_cat=="small")%>%
  ggplot()+
    aes(x=year, y=length, color= common_name,linetype = common_name)+
  geom_line(aes(),size=1.4)+
  xlab("Year")+
  ylab(" Total length (cm)")+
  labs(color="Common name of stock",linetype = "Common name of stock")+
  ggtitle("Small bodied fish growth")+
  theme_minimal()
}

```

```{r med_growth_curves, eval=is_med}
if(nrow(sizeclass) > 0){
  age.growth.join %>%
    filter(body_cat=="med")%>%
    ggplot()+
      aes(x=year, y=length, color= common_name,linetype = common_name)+
    geom_line(aes(),size=1.4)+
    xlab("Year")+
    ylab(" Total length (cm)")+
    labs(color="Common name of stock",linetype = "Common name of stock")+
    ggtitle("Medium bodied fish growth")+
    theme_minimal()
} 

```

```{r large_growth_curves, eval=is_large}
if(nrow(sizeclass) > 0){
  age.growth.join %>%
    filter(body_cat=="large")%>%
    ggplot()+
      aes(x=year, y=length, color= common_name,linetype = common_name)+
    geom_line(aes(),size=1.4)+
    xlab("Year")+
    ylab(" Total length (cm)")+
    labs(color="Common name of stock",linetype = "Common name of stock")+
    ggtitle("large bodied fish growth")+
    theme_minimal()
} 

```

```{r missing_vonb, eval = missing}
 print("NO DATA")
```

<!--chapter:end:08-von_b.Rmd-->

```{r, cond_functions, include = FALSE}
source(here::here("R/full_report_functions", "get_lw.R"))
source(here::here("R/full_report_functions", "get_relative_weight.R"))
```

```{r, cond_data, include = FALSE}
cond_data <- params$cond_data
cond_data <- cond_data[cond_data$Species == species, ]

diet_data <- params$diet_data
diet_data <- diet_data[diet_data$Species == species, ]
```

## Condition

Condition information comes from [diet data](https://github.com/Laurels1/Condition/blob/master/data/allfh.RData); only regions and seasons with more than 10 fish observations were considered. We calculated a rough condition factor as: Weight / Length^3, and relative weight was [previously calculated](https://github.com/Laurels1/Condition/tree/master/data).

### Figures

#### Length vs weight

Please note, no trend lines were fit, points are jittered to reduce overlap, and sometimes the survey observed a small number of fish outside of the defined stock area.
```{r, lw}
plot_lw(diet_data)
```

#### Condition factor: Weight-volume

If there were more than 30 years of data, a geom_gls() regression was fit. In order to fit the geom_gls() regression, we calculated the mean condition factor for each year and plotted the geom_gls() through those points. Please note, points are jittered to reduce overlap, and sometimes the survey observed a small number of fish outside of the defined stock area.
```{r, condition}
plot_cond(diet_data)
```

#### Condition factor: Relative weight

Please note, this data is aggregated by Ecological Protection Unit (EPU), which may differ slightly from the stock assessment regions.
```{r, laurel_cond}
plot_relw(cond_data)
```

### Data

#### Length vs weight with weight-volume condition factor
```{r, lw_data}
data <- diet_data %>%
  dplyr::filter(is.na(pdlen) == FALSE, 
                  is.na(pdwgt) == FALSE) %>%
  dplyr::select(pdlen, pdwgt, season, Region, fish_id, year) %>%
  dplyr::distinct() %>% # remove duplicates
  dplyr::group_by(Region, season) %>%
  dplyr::mutate(n_fish = length(pdlen)) %>%
  dplyr::filter(n_fish > 10) %>% # only region-season with >10 fish)
  dplyr::mutate(condition = pdwgt/(pdlen^3)) %>%
  dplyr::mutate(condition = condition %>%
                  round(digits = 4)) %>%
  dplyr::select(Region, year, season, pdlen, pdwgt, condition) %>%
  character_to_factor()

make_html_table(data, col_names = c("Region", "Year", "Season", "Fish length (cm)", 
                       "Fish weight (g)", "Condition (g/cm^3)"))
```

#### Relative weight condition factor
Please note, this data is aggregated by Ecological Protection Unit (EPU), which may differ slightly from the stock assessment regions.
```{r, laurel_cond_data}
data <- cond_data %>%
  dplyr::select(-Species, -n) %>%
  dplyr::mutate(MeanCond = MeanCond %>%
                  round(digits = 1)) %>%
  character_to_factor()

make_html_table(data, col_names = c("EPU", "Fish sex", "Year", "Mean condition",
                       "Number of fish"))
```

<!--chapter:end:09-condition.Rmd-->


```{r, diet_functions, include = FALSE}
source(here::here("R/full_report_functions", "get_diet.R"))
```

```{r, diet_raw_data, include = FALSE, eval = FALSE}
# already read in
diet_data <- params$diet_data
diet_data <- diet_data[diet_data$Species == species, ]
```

## Diet
Diet data were compiled from [existing data](https://github.com/Laurels1/Condition/blob/master/data/allfh.RData). For analysis, all geographic samples were grouped by season, year, and region, and only year-season-region combinations with more than 20 predators sampled were considered. Prey items that made up more than 5% of the predator's diet in at least one year-season-region were identified to the broad category level; all other prey are grouped into the "other" category.

### Figure
```{r, diet}
get_diet_plot(data = diet_data)
```

### Summary
```{r, diet_summary}
get_diet_table(data = diet_data)
```

### Data
```{r, diet_data}
diet_data2 <- diet_data %>%
    dplyr::filter(pyamtw > 0) %>%
    
    # only look at season/year combinations with >20 predator samples
    dplyr::group_by(year, season, Region) %>%
    dplyr::mutate(n_predators = fish_id %>% unique() %>% length()) %>%
    dplyr::filter(n_predators > 20) %>%
    dplyr::group_by(year, season, Region, n_predators, gensci) %>%
    dplyr::summarise(total_weight = sum(pyamtw)) %>%
    dplyr::mutate(proportion = total_weight/sum(total_weight),
                  proportion = proportion %>%
                    round(digits = 3),
                  total_weight = total_weight %>%
                    round(digits = 2),
                  gensci = gensci %>% 
                    stringr::str_replace("_", " ") %>%
                    stringr::str_to_sentence())  %>%
  character_to_factor()

make_html_table(diet_data2, col_names = c("Year", "Season", "Region", "Number of fish", 
                       "Prey category", "Prey weight", "Prey proportion"))
```

<!--chapter:end:10-diet.Rmd-->

# Population information

<!--chapter:end:11-population-title-page.Rmd-->

```{r, abun_functions, include = FALSE}
source(here::here("R/full_report_functions", "get_assessment.R"))
source(here::here("R/full_report_functions", "get_swept.R"))
```

```{r, abun_data, include = FALSE}
# abun/biomass data (survdat) already read in with habitat data
#survey_data <- params$survey_data
#survey_data <- survey_data[survey_data$Species == species, ]

asmt_data <- params$asmt_data
asmt_data <- asmt_data[asmt_data$Species == species, ]

asmt_sum_data <- params$asmt_sum_data
asmt_sum_data <- asmt_sum_data[asmt_sum_data$Species == species, ]

swept_data <- params$swept_data
swept_data <- swept_data[swept_data$Species == species, ]
```

```{r, legend}
# set colors and linetypes to be constant for all regions
reg <- c(asmt_sum_data$Region, asmt_data$Region) %>% unique()

lines <- 1:length(reg)
names(lines) <- reg

colors <- nmfspalette::nmfs_palette("regional web")(length(reg))
names(colors) <- reg

```

## Abundance

Abundance data were pulled from ``r params$survey_source`` and ``r params$asmt_source``. 

### Figures

Separate geom_gls() functions were fit for fall and spring measurements; trend lines are only shown when the trend was statistically significant, so some plots may have fewer than two trend lines. Fall has solid trend lines, and spring has dashed trend lines. Please note, sometimes the survey observed a small number of fish outside of the defined stock area.

#### Survey abundance (raw measurements)
```{r, abun_survey}
generate_plot(survey_data,
              variable = "ABUNDANCE", 
              ytitle = "Abundance")
```

##### Risk {-}

See Methods for risk calculation details.
```{r}
indicators <- c("abundance_spring", 
                "abundance_fall")
```

```{r, ref.label = "risk_stock", fig.height = length(indicators) + 2}
```

```{r, ref.label = "risk_stock2", fig.height = length(indicators) + 1}
```

#### Survey abundance (swept area estimates)

Please note, these estimates are not parsed by region or season. Swept area estimates are based on spring and fall surveys only. The shaded gray region indicates +/- two standard errors.
```{r, abun_survey_swept}
plot_swept_abun(swept_data)
```

#### Assessment abundance
```{r}
if(nrow(asmt_data %>% dplyr::filter(Metric == "Abundance")) > 0){
ndesc <- asmt_data %>% 
  dplyr::filter(Metric == "Abundance") %>% 
  #dplyr::mutate(options = paste(Description, Units)) %>%
  dplyr::select(Age) %>% 
  unique() %>% 
  nrow()
} else ndesc <- 1
```

```{r, abun_asmt, fig.height = ndesc * 4}
plot_asmt(asmt_data, 
          metric = "Abundance", 
          ytitle = "Abundance",
          lin = lines, 
          col = colors)
```

##### Risk {-}

See Methods for risk calculation details.
```{r}
indicators <- c("asmt_abundance")
```

```{r, ref.label = "risk_stock", fig.height = length(indicators) + 2}
```

```{r, ref.label = "risk_stock2", fig.height = length(indicators) + 1}
```

### Survey summary
```{r, abun_summary}
generate_table(survey_data,
              variable = "ABUNDANCE", 
              cap = "Measured abundance")
```

### Data

#### Survey data (raw measurements)
```{r, surv_abun_data}
data <- survey_data %>%
  get_var_data(variable = "ABUNDANCE") %>%
  dplyr::mutate(variable = variable %>%
                  format(big.mark = ","))  %>%
  character_to_factor()

make_html_table(data, col_names = c("Year", "Season", "Region", "Abundance"))
```

#### Survey data (swept area estimates)
```{r, swept_abun_data}
data <- swept_data %>%
  dplyr::select(YEAR, tot.abundance, tot.abund.SE) %>%
  dplyr::mutate(YEAR = as.numeric(YEAR),
                tot.abundance = tot.abundance %>%
                  format(big.mark = ","),
                tot.abund.SE = tot.abund.SE %>%
                  round(digits = 1) %>%
                  format(big.mark = ","))

make_html_table(data, col_names = c("Year", "Abundance", "Standard error"))
```

#### Assessment data
```{r, asmt_abun_data}
data <- asmt_data %>% 
                dplyr::filter(Metric == "Abundance") %>%
                dplyr::select(-Species, -Metric) %>%
                dplyr::mutate(Value = Value %>%
                                round(digits = 0)) %>%
                dplyr::mutate(Value = Value %>%
                              format(big.mark = ",")) %>%
  character_to_factor()

make_html_table(data, col_names = colnames(data))
```

<!--chapter:end:12-abundance.Rmd-->

```{r, biomass_functions, include = FALSE, eval = FALSE}
# already read in with abundance
source(here::here("R/full_report_functions", "get_assessment.R"))
source(here::here("R/full_report_functions", "get_swept.R"))
```

```{r, biomass_raw_data, include = FALSE, eval = FALSE}
# already read in with abundance

# abun/biomass data (survdat) already read in with habitat data
#survey_data <- params$survey_data
#survey_data <- survey_data[survey_data$Species == species, ]

asmt_data <- params$asmt_data
asmt_data <- asmt_data[asmt_data$Species == species, ]

swept_data <- params$swept_data
swept_data <- swept_data[swept_data$Species == species, ]
```

```{r, eval = FALSE}
# already read in with abundance
# set colors and linetypes to be constant for all regions
reg <- c(asmt_sum_data$Region, asmt_data$Region) %>% unique()

lines <- 1:length(reg)
names(lines) <- reg

colors <- nmfspalette::nmfs_palette("regional web")(length(reg))
names(colors) <- reg

```

## Biomass

Biomass data were pulled from ``r params$survey_source``. 

### Figures

Separate geom_gls() functions were fit for fall and spring measurements; trend lines are only shown when the trend was statistically significant, so some plots may have fewer than two trend lines. Fall has solid trend lines, and spring has dashed trend lines. Please note, sometimes the survey observed a small number of fish outside of the defined stock area.

#### Survey biomass (raw measurements)
```{r, biomass}
generate_plot(survey_data,
              variable = "BIOMASS", 
              ytitle = "Biomass")
```

##### Risk {-}

See Methods for risk calculation details.
```{r}
indicators <- c("biomass_spring", 
                "biomass_fall")
```

```{r, ref.label = "risk_stock", fig.height = length(indicators) + 2}
```

```{r, ref.label = "risk_stock2", fig.height = length(indicators) + 1}
```

#### Survey biomass (swept area estimates)

Please note, these estimates are not parsed by region or season. Swept area estimates are based on spring and fall surveys only. The shaded gray region indicates +/- two standard errors.
```{r, bio_survey_swept}
plot_swept_biomass(swept_data)
```

#### Assessment biomass
```{r}
if(nrow(asmt_data %>% dplyr::filter(Metric == "Biomass")) > 0){
ndesc <- asmt_data %>% 
  dplyr::filter(Metric == "Biomass") %>% 
  #dplyr::mutate(options = paste(Description, Units)) %>%
  dplyr::select(Age) %>% 
  unique() %>% 
  nrow()
} else ndesc <- 1
```

```{r, biomass_asmt,  fig.height = ndesc * 4 + 1}
plot_asmt(asmt_data, 
          metric = "Biomass", 
          ytitle = "Biomass",
          lin = lines, 
          col = colors)
```

##### Risk {-}

See Methods for risk calculation details.
```{r}
indicators <- c("asmt_biomass")
```

```{r, ref.label = "risk_stock", fig.height = length(indicators) + 2}
```

```{r, ref.label = "risk_stock2", fig.height = length(indicators) + 1}
```

### Survey summary
```{r, biomass_summary}
generate_table(survey_data,
              variable = "BIOMASS", 
              cap = "Measured biomass")
```

### Data

#### Survey data (raw measurements)
```{r, biomass_data}
data <- survey_data %>% 
  get_var_data(variable = "BIOMASS") %>%
  dplyr::mutate(variable = variable %>%
                     format(big.mark = ",")) %>%
  character_to_factor()

make_html_table(data, col_names = c("Year", "Season", "Region", "Biomass"))
```

#### Survey data (swept area estimates)
```{r, swept_biomass_data}
data <- swept_data %>%
  dplyr::select(YEAR, tot.biomass, tot.bio.SE) %>%
  dplyr::mutate(YEAR = as.numeric(YEAR),
                tot.biomass = tot.biomass %>%
                  format(big.mark = ","),
                tot.bio.SE = tot.bio.SE %>%
                  round(digits = 1) %>%
                  format(big.mark = ","))

make_html_table(data, col_names = c("Year", "Biomass (kg)", "Standard error"))
```

#### Assessment data
```{r, asmt_biomass_data}
data <- asmt_data %>%
  dplyr::filter(Metric == "Biomass") %>%
  dplyr::select(-Species, -Metric) %>%
  dplyr::mutate(Value = Value %>%
                  round(digits = 0)) %>%
  dplyr::mutate(Value = Value %>%
                  format(big.mark = ",")) %>%
  character_to_factor()

make_html_table(data, col_names = colnames(data))
```

<!--chapter:end:13-biomass.Rmd-->


```{r, b_functions, include = FALSE}
source(here::here("R/full_report_functions", "get_bf.R"))
```

```{r, b_data, include = FALSE}
asmt_sum_data <- params$asmt_sum_data
asmt_sum_data <- asmt_sum_data[asmt_sum_data$Species == species, ]
```

```{r, b_status}
if(length(asmt_sum_data$Species) > 1){
  list_regions <- split(unique(asmt_sum_data$Region), 
                        f = list(unique(asmt_sum_data$Region)))
  bbmsy2 <- purrr::map(list_regions, 
                       ~status(data = asmt_sum_data, 
                               regions = .x, 
                               metric = "bbmsy"))
} else bbmsy2 <- "UNKNOWN"

```

## B/Bmsy 

B/Bmsy data were pulled from ``r params$asmt_sum_source``.

The most recent status of B/Bmsy is: `r bbmsy2`

### Figure
```{r, bbmsy_fig} 
if("GOOD" %in% bbmsy2 | "DANGER" %in% bbmsy2){
  plot_msy(asmt_sum_data[which(is.na(asmt_sum_data$`B/Bmsy`) == FALSE), ], 
           type = "b",
             ytitle = "B / Bmsy",
             lin = lines,
             col = colors)
} else print("NO DATA")
```

##### Risk

See Methods for risk calculation details.
```{r}
indicators <- c("bbmsy")
```

```{r, ref.label = "risk_stock", fig.height = length(indicators) + 2}
```

```{r, ref.label = "risk_stock2", fig.height = length(indicators) + 1}
```

### Data
```{r, bbmsy_data}
data <- asmt_sum_data[which(is.na(asmt_sum_data$`B/Bmsy`) == FALSE), ] %>% 
                dplyr::select(Region, `Assessment Year`, `B/Bmsy`) %>%
                dplyr::mutate(`B/Bmsy` = `B/Bmsy` %>% round(digits = 2)) %>%
  character_to_factor()

make_html_table(data, col_names = c("Region", "Year", "B/Bmsy"))
```

<!--chapter:end:14-bbmsy.Rmd-->


```{r, recruit_functions, include = FALSE, eval = FALSE}
# already read in 
source(here::here("R/full_report_functions", "get_assessment.R"))
```

```{r, recruit_raw_data, include = FALSE, eval = FALSE}
# already read in
asmt_data <- params$asmt_data
asmt_data <- asmt_data[asmt_data$Species == species, ]
```

## Recruitment
Recruitment data were pulled from ``r params$asmt_source``. Separate geom_gls() functions were fit for each region; trend lines are only shown when the trend was statistically significant, so some plots may have fewer trend lines than regions.

### Figure
```{r}
if(nrow(asmt_data %>% dplyr::filter(Metric == "Recruitment")) > 0){
ndesc <- asmt_data %>% 
  dplyr::filter(Metric == "Recruitment") %>% 
  #dplyr::mutate(options = paste(Description, Units)) %>%
  dplyr::select(Age) %>% 
  unique() %>% 
  nrow()
} else ndesc <- 1
```

```{r, recruitment, fig.height = ndesc * 4 + 1}
plot_asmt(asmt_data, 
          metric = "Recruitment", 
          ytitle = "Recruitment",
          lin = lines, 
          col = colors)
```

##### Risk

See Methods for risk calculation details.
```{r}
indicators <- c("recruitment")
```

```{r, ref.label = "risk_stock", fig.height = length(indicators) + 2}
```

```{r, ref.label = "risk_stock2", fig.height = length(indicators) + 1}
```

### Data

```{r, recruit_data}
data <- asmt_data %>% 
  dplyr::filter(Metric == "Recruitment") %>%
  dplyr::select(-Species, -Metric) %>%
  dplyr::mutate(Value = Value %>%
                  round(digits = 0)) %>%
  dplyr::mutate(Value = Value %>%
                  format(big.mark = ",",
                         scientific = FALSE)) %>%
  character_to_factor()

make_html_table(data, col_names = colnames(data))
```

<!--chapter:end:15-recruitment.Rmd-->

# Socio-economic information

<!--chapter:end:16-socio-economic-title-page.Rmd-->


```{r, catch_functions, include = FALSE}
# already run
#source(here::here("R/full_report_functions", "get_assessment.R"))

source(here::here("R/full_report_functions", "get_rec_catch.R"))
source(here::here("R/full_report_functions", "get_com_catch.R"))
source(here::here("R/full_report_functions", "compare_rec_com.R"))
```

```{r, catch_raw_data, include = FALSE}
# already read in
#asmt_data <- params$asmt_data
#asmt_data <- asmt_data[asmt_data$Species == species, ]

rec_data <- params$rec_data
rec_data <- rec_data[rec_data$Species == species, ]

com_data <- params$com_data
com_data <- com_data[com_data$Species == species, ]
```

```{r, eval = FALSE}
# already run in pop section
# set colors and linetypes to be constant for all regions
reg <- c(asmt_sum_data$Region, asmt_data$Region) %>% unique()

lines <- 1:length(reg)
names(lines) <- reg

colors <- nmfspalette::nmfs_palette("regional web")(length(reg))
names(colors) <- reg

```

## Catch

Stock assessment catch data are from ``r params$asmt_source``. Recreational catch data were downloaded from [NOAA MRIP](https://www.st.nmfs.noaa.gov/st1/recreational/MRIP_Estimate_Data/CSV/). Commercial catch data were downloaded from [NOAA FOSS](https://foss.nmfs.noaa.gov/apexfoss/f?p=215:200:4615327020711::NO:::).

### Figures
```{r}
if(nrow(asmt_data %>% dplyr::filter(Metric == "Catch")) > 0){
ndesc <- asmt_data %>% 
  dplyr::filter(Metric == "Catch") %>% 
  dplyr::mutate(options = paste(Description, Units)) %>%
  dplyr::select(options) %>% 
  unique() %>% 
  nrow()
} else ndesc <- 1
```

#### Stock assessment catch
```{r, catch}
#  fig.height = ndesc * 10
plot_asmt(asmt_data, 
          metric = "Catch", 
          ytitle = "Catch",
          lin = lines, 
          col = colors)
```

##### Risk {-}

See Methods for risk calculation details.
```{r}
indicators <- c("asmt_catch")
```

```{r, ref.label = "risk_stock", fig.height = length(indicators) + 2}
```

```{r, ref.label = "risk_stock2", fig.height = length(indicators) + 1}
```

#### Recreational catch
```{r, rec}
get_rec_catch(rec_data)
```

##### Risk {-}

See Methods for risk calculation details.
```{r}
indicators <- c("rec_catch")
```

```{r, ref.label = "risk_stock", fig.height = length(indicators) + 2}
```

```{r, ref.label = "risk_stock2", fig.height = length(indicators) + 1}
```

#### Commercial catch
```{r, com}
plot_com(com_data)
```

##### Risk {-}

See Methods for risk calculation details.
```{r}
indicators <- c("com_catch")
```

```{r, ref.label = "risk_stock", fig.height = length(indicators) + 2}
```

```{r, ref.label = "risk_stock2", fig.height = length(indicators) + 1}
```

#### Commercial vs recreational catch
```{r, comvrec}
prop_catch(com = com_data, rec = rec_data)
```

### Data

#### Stock assessment catch
```{r, catch_data}
data <- asmt_data %>% 
  dplyr::filter(Metric == "Catch") %>%
  dplyr::select(-Species, -Metric) %>%
  dplyr::mutate(Value = Value %>%
                  round(digits = 0)) %>%
  dplyr::mutate(Value = Value %>%
                  format(big.mark = ",")) %>%
  character_to_factor()

make_html_table(data, col_names = colnames(data))
```

#### Recreational catch
```{r, rec_data}
data <- rec_data %>% 
  dplyr::select(year, st_f, mode_fx_f, tot_cat) %>%
  dplyr::mutate(State = st_f %>%
                  stringr::str_to_title(),
                Method = mode_fx_f %>% 
                  stringr::str_to_sentence(),
                Catch_pounds = tot_cat %>%
                  format(big.mark = ",")) %>%
  dplyr::select(year, State, Method, Catch_pounds) %>%
  character_to_factor()

make_html_table(data, col_names = c("Year", "State", "Method", "Catch (lb)"))
```

#### Commercial catch
```{r, com_data}
data <- com_data %>% 
                dplyr::select(Year, State, Dollars_adj, Pounds) %>%
                dplyr::mutate(Dollars_adj = Dollars_adj %>% 
                                round(digits = 0) %>%
                                format(big.mark = ","),
                              State = State %>% 
                                stringr::str_to_title(),
                              Pounds = Pounds %>%
                                format(big.mark = ",")) %>%
                dplyr::rename("Revenue (2019 dollars)" = "Dollars_adj",
                              "Commercial catch (lb)" = "Pounds") %>%
  character_to_factor()

make_html_table(data, col_names = colnames(data))
```

#### Commercial vs recreational catch
```{r, comvrec_data}
data <- prop_catch_data(rec = rec_data, com = com_data) %>%
  character_to_factor()

make_html_table(data, col_names = c("Year", "Recreational catch", 
                                    "Commercial catch", "Total catch", 
                                    "Proportion recreational",
                                    "Proportion commercial"))
```

#### Commercial, recreational, and stock assessment catch
```{r, all_catch}
stock <- asmt_data %>%
  dplyr::filter(Metric == "Catch",
                Region != "Eastern Georges Bank") %>% # double counted
  dplyr::mutate(options = paste(Description, Units, AssessmentYear, sep = " - ")) %>%
  dplyr::group_by(Year, options) %>%
  dplyr::summarise(Value = sum(Value)) %>%
  dplyr::ungroup() %>%
  tidyr::pivot_wider(names_from = options,
                     values_from = Value)

recr <- rec_data %>%
  dplyr::group_by(year) %>%
  dplyr::rename(Year = year) %>%
  dplyr::summarise(`Recreational catch (mt)` = sum(tot_cat)/2204.6)

comm <- com_data %>%
  dplyr::group_by(Year) %>%
  dplyr::summarise(`Commercial catch (mt)` = sum(Pounds)/2204.6)

all_catch <- c()
if(nrow(stock) > 0){
  all_catch <- stock
}
if(nrow(recr) > 0){
  if(is.null(all_catch) == FALSE){
    all_catch <- dplyr::full_join(all_catch, recr, by = "Year")
  } else all_catch <- recr
}
if(nrow(comm) > 0){
  if(is.null(all_catch) == FALSE){
    all_catch <- dplyr::full_join(all_catch, comm, by = "Year")
  } else all_catch <- comm
}

if(is.null(all_catch) == FALSE &
   ncol(all_catch > 2)){
  all_catch <- all_catch %>% 
    dplyr::arrange(Year) %>%
    round(digits = 0) %>%
    #format(big.mark = ",") %>% # messing up formatting
    as.data.frame()
}

# make sure there is data on at least 2 types of catch
test <- c(nrow(stock), nrow(recr), nrow(comm))
sum_test <- sum(test == 0)

if(sum_test >= 2){
  print("NO DATA")
} else make_html_table(all_catch, col_names = colnames(all_catch))
```


<!--chapter:end:17-catch.Rmd-->

```{r, f_functions, include = FALSE, eval = FALSE}
# already run
source(here::here("R/full_report_functions", "get_bf.R")) 
```

```{r, f_data, include = FALSE, eval = FALSE}
# already run
asmt_sum_data <- params$asmt_sum_data
asmt_sum_data <- asmt_sum_data[asmt_sum_data$Species == species, ]
```

```{r, f_status}
if(length(asmt_sum_data$Species) > 1){
  list_regions <- split(unique(asmt_sum_data$Region), 
                        f = list(unique(asmt_sum_data$Region)))
  ffmsy2 <- purrr::map(list_regions, 
                       ~status(data = asmt_sum_data, 
                               regions = .x, 
                               metric = "ffmsy"))
} else ffmsy2 <- "UNKNOWN"

```

```{r, eval = FALSE}
# already run 
# set colors and linetypes to be constant for all regions
reg <- c(asmt_sum_data$Region, asmt_data$Region) %>% unique()

lines <- 1:length(reg)
names(lines) <- reg

colors <- nmfspalette::nmfs_palette("regional web")(length(reg))
names(colors) <- reg

```

## F/Fmsy 

F/Fmsy data were pulled from ``r params$asmt_sum_source``.

The most recent status of F/Fmsy is: `r ffmsy2`

### Figure
```{r, ffmsy}
if("GOOD" %in% ffmsy2 | "DANGER" %in% ffmsy2){
 plot_msy(asmt_sum_data[which(is.na(asmt_sum_data$`F/Fmsy`) == FALSE), ],
          ytitle = "F / Fmsy",
          type = "f",
             lin = lines,
             col = colors)
} else print("NO DATA")
```

##### Risk {-}

See Methods for risk calculation details.
```{r}
indicators <- c("ffmsy")
```

```{r, ref.label = "risk_stock", fig.height = length(indicators) + 2}
```

```{r, ref.label = "risk_stock2", fig.height = length(indicators) + 1}
```

### Data
```{r, ffmsy_data}
data <- asmt_sum_data[which(is.na(asmt_sum_data$`F/Fmsy`) == FALSE), ] %>% 
                dplyr::select(Region, `Assessment Year`, `F/Fmsy`) %>%
                dplyr::mutate(`F/Fmsy` = `F/Fmsy` %>% round(digits = 2)) %>%
  character_to_factor()

make_html_table(data, col_names = c("Region", "Year", "F/Fmsy"))
```

<!--chapter:end:18-ffmsy.Rmd-->


```{r, revenue_functions, include = FALSE}
source(here::here("R/full_report_functions", "get_com_money.R"))
```

```{r, revenue_raw_data, include = FALSE, eval = FALSE}
# already read
com_data <- params$com_data
com_data <- com_data[com_data$Species == species, ]
```


## Revenue 

Commercial catch data were downloaded from [NOAA FOSS](https://foss.nmfs.noaa.gov/apexfoss/f?p=215:200:4615327020711::NO:::).

### Figure
```{r, revenue}
plot_com_money(com_data)
```

##### Risk {-}

See Methods for risk calculation details.
```{r}
indicators <- c("revenue")
```

```{r, ref.label = "risk_stock", fig.height = length(indicators) + 2}
```

```{r, ref.label = "risk_stock2", fig.height = length(indicators) + 1}
```

### Data
```{r, revenue_data}
data <- com_data %>% 
                dplyr::select(Year, State, Dollars_adj, Pounds) %>%
                dplyr::mutate(Dollars_adj = Dollars_adj %>% 
                                round(digits = 0) %>%
                                format(big.mark = ","),
                              State = State %>% 
                                stringr::str_to_title(),
                              Pounds = Pounds %>%
                                format(big.mark = ",")) %>%
                dplyr::rename("Revenue (2019 dollars)" = "Dollars_adj",
                              "Commercial catch (lb)" = "Pounds") %>%
  character_to_factor()

make_html_table(data, col_names = colnames(data))
```

<!--chapter:end:19-revenue.Rmd-->

# Management information

```{r, management_functions, include = FALSE}
# none
```

```{r, management_data, include = FALSE}
# already read in
#asmt_sum_data <- params$asmt_sum_data
#asmt_sum_data <- asmt_sum_data[asmt_sum_data$Species == species, ]

ad_rating_data <- params$asmt_sum_data %>% 
  dplyr::select(Species, Region, `Assessment Year`, `Last Data Year`,
                `Biological Input Data`, `Life History Data`, 
                `Composition Input Data`, `Ecosystem Linkage`, `FSSI Stock?`,
                `Review Result`) %>%
  dplyr::rename(`Biological Data Rating (2019)` = `Biological Input Data`,
                `Biological Data Rating (before 2019)` = `Life History Data`,
                `Size Data Rating` = `Composition Input Data`,
                `Ecosystem Linkage Data Rating` = `Ecosystem Linkage`,
                `FSSI status` = `FSSI Stock?`)
```

## Stock assessment and data quality information
Stock assessment and data quality information were pulled from ``r params$asmt_sum_source``.
```{r, quality}
data <- ad_rating_data %>%
  character_to_factor()

make_html_table(data, col_names = colnames(data))
```

<!--chapter:end:20-management.Rmd-->

# Risk assessment

```{r, risk_functions, include = FALSE}
source(here::here("R/full_report_functions/risk_functions", "plot_risk_within_species.R"))
source(here::here("R/full_report_functions/risk_functions", "plot_risk_by_year.R"))
```

```{r, risk_data, include = FALSE, eval = FALSE}
# don't use?
risk_data <- params$risk_data
risk_data <- risk_data[risk_data$Species == species, ]

# already read in
risk_year_hist_data <- params$risk_year_hist_data
risk_year_hist_data <- risk_year_hist_data[risk_year_hist_data$Species == species, ]

risk_year_value_data <- params$risk_year_value_data
risk_year_value_data <- risk_year_value_data[risk_year_value_data$Species == species, ]

risk_species_data <- params$risk_species_data
risk_species_data <- risk_species_data[risk_species_data$Species == species, ]
```

## Preliminary risk calculation

A preliminary risk analysis was conducted by ranking all species according to their indicator values. A high rank number and a normalized rank near 1 indicates that the species is at risk or of importance based on the measured indicator values. When a species was missing an indicator, it was assigned a normalized rank of 0.5.

### Relative to all other stocks
```{r, risk_all}
data <- risk_year_hist_data %>% 
  dplyr::mutate(Rank = paste(rank, "out of", n_stocks_per_indicator),
                       Value2 = Value %>% 
                         format(big.mark = ",", scientific = FALSE, digits = 2),
                       Normalized_rank = norm_rank %>% round(digits = 2)) %>%
  dplyr::select(Region, category, Indicator, Year, Value2, Rank, Normalized_rank) %>%
  dplyr::rename("Value" = "Value2") %>%
  character_to_factor()

make_html_table(data, col_names = c("Region", "Category", "Indicator", "Year", "Value (% of historical)",
                       "Rank", "Normalized rank"))

data <- risk_year_value_data %>% 
  dplyr::mutate(Rank = paste(rank, "out of", n_stocks_per_indicator),
                       Value2 = Value %>% 
                         format(big.mark = ",", scientific = FALSE, digits = 2),
                       Normalized_rank = norm_rank %>% round(digits = 2)) %>%
  dplyr::select(Region, category, Indicator, Year, Value2, Rank, Normalized_rank) %>%
  dplyr::rename("Value" = "Value2") %>%
  character_to_factor()

make_html_table(data, col_names = c("Region", "Category", "Indicator", "Year", "Value (magnitude)",
                       "Rank", "Normalized rank"))
```

### Within each stock
```{r, risk_within}
data <- risk_species_data %>% dplyr::mutate(Rank = paste(rank, "out of", n_years_per_indicator),
                       Value2 = Value %>% 
                         format(big.mark = ",", scientific = FALSE, digits = 2),
                       Normalized_rank = norm_rank %>% round(digits = 2)) %>%
  dplyr::select(Region, category, Indicator, Year, Value2, Rank, Normalized_rank) %>%
  dplyr::rename("Value" = "Value2") %>%
  character_to_factor()

make_html_table(data, col_names = c("Region", "Category", "Indicator", "Year", "Value",
                       "Rank", "Normalized rank"))
```

## Preliminary risk visualization

### Relative to all other stocks

Risk was calculated over time for all indicators that were documented for five or more species in a given year. Risk was calculated as the average of the past 5 years, as a percent of the historical average. The normalized risk value plotted here reflects the normalized rank of this stock compared to all other stocks in that year.

```{r, year_risk}
plot_risk_by_year(risk_year_hist_data,
                  indicator = "all",
                  title = "Change compared to historical",
                  include_legend = "yes")

plot_risk_by_year(risk_year_hist_data,
                  indicator = "all",
                  title = "Value compared to other stocks",
                  include_legend = "yes")
```

### Within a single stock

For each stock, a five-year running mean was calculated for each indicator. Indicator values were then ranked for all years where a value was present. The normalized risk values plotted here reflects the normalized rank of each year compared to all other years.

```{r, stock_risk}
plot_risk_by_stock(risk_species_data,
                  indicator = "all",
                  include_legend = "yes")
```

<!--chapter:end:21-risk-assessment.Rmd-->

