## Physical indicators

### Cold pool index
```{r, cold-pool}
test <- ecodata::cold_pool
NEesp::render_indicator(test, lab = names[1])
```

### Warm core rings
```{r, warm}
test <- ecodata::wcr
NEesp::render_indicator(test, lab = names[1])
```

### Marine heatwave index
```{r, heatwave}
test <- ecodata::heatwave
NEesp::render_indicator(test, lab = names[1])
```

### GLORYS bottom temperature
```{r, glorys}
test <- ecodata::bottom_temp_glorys
NEesp::render_indicator(test, lab = names[1])
```

### Long-term sea surface temperature
```{r, sst}
test <- ecodata::long_term_sst
NEesp::render_indicator(test, lab = names[1])
```

### Sea surface temperature anomaly in EPU
```{r, sst-anom}
test <- ecodata::seasonal_oisst_anom
NEesp::render_indicator(test, lab = names[1])
```

### Sea surface temperature anomaly in stock region
```{r, sst-anom-stock}
test <- ecodata::ESP_seasonal_oisst_anom %>%
  dplyr::mutate(Pattern_check = ESP %>%
    stringr::str_replace_all("_", " ") %>%
    stringr::str_to_sentence()) %>%
  dplyr::filter(stringr::str_detect(Pattern_check, params$stock)) %>%
  dplyr::distinct()

if (nrow(test) > 0) {
  if (length(unique(test$Pattern_check)) > 1) {
    # add stock region into var for faceting
    test <- test %>%
      dplyr::mutate(Var = paste(Var, Pattern_check, sep = " ") %>%
        stringr::str_replace_all("_", "") %>%
        stringr::str_wrap(width = 8))
    if (nrow(test) > 0) {
      NEesp::render_indicator(test, lab = names[1])
    }
  } else {
    test <- test %>%
      dplyr::mutate(Var = paste0(Var, "_stock"))

    NEesp::render_indicator(test, lab = names[1])
  }
} else {
  "Indicator data has not been parsed over this stock's region"
}
```

### Stratification
```{r, strat}
test <- ecodata::stratification
NEesp::render_indicator(test, lab = names[1])
```

### Winter wind speed
```{r, winter-wind}
test <- ecodata::ne_wind %>%
  dplyr::filter(Var %>% stringr::str_detect("winter"))
NEesp::render_indicator(test, lab = names[1])
```

### Spring wind speed
```{r, spring-wind}
test <- ecodata::ne_wind %>%
  dplyr::filter(Var %>% stringr::str_detect("spring"))
NEesp::render_indicator(test, lab = names[1])
```

### Summer wind speed
```{r, summer-wind}
test <- ecodata::ne_wind %>%
  dplyr::filter(Var %>% stringr::str_detect("summer"))
NEesp::render_indicator(test, lab = names[1])
```

### Fall wind speed
```{r, fall-wind}
test <- ecodata::ne_wind %>%
  dplyr::filter(Var %>% stringr::str_detect("fall"))
NEesp::render_indicator(test, lab = names[1])
```

### Gulf Stream Index
```{r, gsi}
test <- ecodata::gsi %>%
  dplyr::mutate(Time = Time %>%
    stringr::str_trunc(width = 4, ellipsis = "") %>%
    as.numeric())
NEesp::render_indicator(test, lab = names[1])
```

### North Atlantic Oscillation
```{r, nao}
test <- ecodata::nao
NEesp::render_indicator(test, lab = names[1])
```

### Northern range
Northernmost survey observation in each year.
```{r, northern-bound}
test <- NEesp::survey %>%
  dplyr::filter(Species == params$stock,
                ABUNDANCE > 0) %>% # why are there entries with 0 abundance but >0 biomass??
  dplyr::select(YEAR, LAT) %>% 
  dplyr::group_by(YEAR) %>%
  dplyr::mutate(Val = max(LAT, na.rm = TRUE),
                Var = "northern_latitude") %>%
  dplyr::select(-LAT) %>%
  dplyr::rename(Time = YEAR) %>%
  dplyr::distinct()
NEesp::render_indicator(test, lab = names[1])
```

### Southern range
Southernmost survey observation in each year.
```{r, southern-bound}
test <- NEesp::survey %>%
  dplyr::filter(Species == params$stock,
                ABUNDANCE > 0) %>% # why are there entries with 0 abundance but >0 biomass??
  dplyr::select(YEAR, LAT)  %>% 
  dplyr::group_by(YEAR) %>%
  dplyr::mutate(Val = min(LAT, na.rm = TRUE),
                Var = "southern_latitude") %>%
  dplyr::select(-LAT) %>%
  dplyr::rename(Time = YEAR) %>%
  dplyr::distinct()
NEesp::render_indicator(test, lab = names[1])
```
