## Physical indicators

### Cold pool index
```{r, cold-pool}
test <- ecodata::cold_pool
NEesp::render_indicator(test, lab = names[1])
```

### Warm core rings
```{r, warm}
test <- ecodata::wcr
NEesp::render_indicator(test, lab = names[1])
```

### Marine heatwave index
```{r, heatwave}
test <- ecodata::heatwave
NEesp::render_indicator(test, lab = names[1])
```

### GLORYS bottom temperature
```{r, glorys}
test <- ecodata::bottom_temp_glorys
NEesp::render_indicator(test, lab = names[1])
```

### Long-term sea surface temperature
```{r, sst}
test <- ecodata::long_term_sst
NEesp::render_indicator(test, lab = names[1])
```

### Sea surface temperature anomaly in EPU
```{r, sst-anom}
test <- ecodata::seasonal_oisst_anom
NEesp::render_indicator(test, lab = names[1])
```

### Sea surface temperature anomaly in stock region
```{r, sst-anom-stock}
test <- ecodata::ESP_seasonal_oisst_anom %>%
  dplyr::mutate(Pattern_check = ESP %>%
                  stringr::str_replace_all("_", " ") %>%
                  stringr::str_to_sentence()) %>%
  dplyr::filter(stringr::str_detect(Pattern_check, params$stock)) %>%
  dplyr::distinct() 

if(nrow(test) > 0){
  
if(length(unique(test$Pattern_check)) > 1){
  # add stock region into var for faceting
  test <- test %>%
    dplyr::mutate(Var = paste(Var, Pattern_check, sep = " ") %>%
                    stringr::str_replace_all("_", "") %>%
                    stringr::str_wrap(width = 8))
  if(nrow(test) > 0){
    NEesp::render_indicator(test, lab = names[1])
  }
} else {
  test <- test %>%
    dplyr::mutate(Var = paste0(Var, "_stock"))
  
  NEesp::render_indicator(test, lab = names[1])
}
  
}
```

### Stratification
```{r, strat}
test <- ecodata::stratification
NEesp::render_indicator(test, lab = names[1])
```

### Winter wind speed
```{r, winter-wind}
test <- ecodata::ne_wind %>%
  dplyr::filter(Var %>% stringr::str_detect("winter"))
NEesp::render_indicator(test, lab = names[1])
```

### Spring wind speed
```{r, spring-wind}
test <- ecodata::ne_wind %>%
  dplyr::filter(Var %>% stringr::str_detect("spring"))
NEesp::render_indicator(test, lab = names[1])
```

### Summer wind speed
```{r, summer-wind}
test <- ecodata::ne_wind %>%
  dplyr::filter(Var %>% stringr::str_detect("summer"))
NEesp::render_indicator(test, lab = names[1])
```

### Fall wind speed
```{r, fall-wind}
test <- ecodata::ne_wind %>%
  dplyr::filter(Var %>% stringr::str_detect("fall"))
NEesp::render_indicator(test, lab = names[1])
```

### Gulf Stream Index
```{r, gsi}
test <- ecodata::gsi %>%
  dplyr::mutate(Time = Time %>%
    stringr::str_trunc(width = 4, ellipsis = "") %>%
    as.numeric())
NEesp::render_indicator(test, lab = names[1])
```

### North Atlantic Oscillation
```{r, nao}
test <- ecodata::nao
NEesp::render_indicator(test, lab = names[1])
```

### Chlorophyll
```{r, chl}
test <- ecodata::chl_pp %>%
  dplyr::mutate(Var = Var %>%
    stringr::str_replace_all("_", "\n") %>%
    stringr::str_replace_all(" ", "\n")) %>%
  dplyr::filter(stringr::str_detect(Var, "ANNUAL")) %>% # only annual indices
  dplyr::mutate(Time = Time %>%
    stringr::str_replace("A_", "") %>%
    as.numeric())
NEesp::render_indicator(test, lab = names[1])

# data_prep(stock, test, lag = params$lag) %>% knitr::kable()
```

### Chlorophyll in stock region
```{r, chl-stock}
test <- ecodata::ESP_seasonal_chl %>%
  dplyr::mutate(Pattern_check = ESP %>%
                  stringr::str_replace_all("_", " ") %>%
                  stringr::str_to_sentence()) %>%
  dplyr::filter(stringr::str_detect(Pattern_check, params$stock)) %>%
  dplyr::distinct() 

if(nrow(test) > 0){
  
if(length(unique(test$Pattern_check)) > 1){
  # add stock region into var for faceting
  test <- test %>%
    dplyr::mutate(Var = paste(Var, Pattern_check, sep = " ") %>%
                    stringr::str_replace_all("_", "") %>%
                    stringr::str_wrap(width = 8))
  NEesp::render_indicator(test, lab = names[1])
} else {
  test <- test %>%
    dplyr::mutate(Var = paste0(Var, "_stock"))
  
  NEesp::render_indicator(test, lab = names[1])
}
}
```

### Primary production in stock region
```{r, pp-stock}
test <- ecodata::ESP_seasonal_pp %>%
  dplyr::mutate(Pattern_check = ESP %>%
                  stringr::str_replace_all("_", " ") %>%
                  stringr::str_to_sentence()) %>%
  dplyr::filter(stringr::str_detect(Pattern_check, params$stock)) %>%
  dplyr::distinct() 

if(nrow(test) > 0){
  
if(length(unique(test$Pattern_check)) > 1){
  # add stock region into var for faceting
  test <- test %>%
    dplyr::mutate(Var = paste(Var, Pattern_check, sep = " ") %>%
                    stringr::str_replace_all("_", "") %>%
                    stringr::str_wrap(width = 8))
  NEesp::render_indicator(test, lab = names[1])
} else {
  test <- test %>%
    dplyr::mutate(Var = paste0(Var, "_stock"))
  
  NEesp::render_indicator(test, lab = names[1])
}
}
```