# Summary of statistically significant indicators

```{r, output_summary_html, eval = (params$out == "html")}
summary_table <- function(x) {
  DT::datatable(x,
    rownames = FALSE,
    filter = list(
      position = "top",
      clear = FALSE
    ),
    options = list(
      search = list(regex = TRUE),
      paging = FALSE
    )
  )
}
```

```{r, output_summary_word, eval = (params$out == "word")}
summary_table <- function(x) {
  knitr::kable(x)
}
```

```{r}
colnames(output) <- c(
  "Response_variable", "Indicator",
  "Number of data points", "Slope",
  "P-value", "R2_adj"
)
```

## Abundance
```{r}
output %>%
  tibble::as_tibble() %>%
  dplyr::filter(
    Response_variable == "Abundance",
    R2_adj < 1
  ) %>%
  dplyr::select(-Response_variable) %>%
  summary_table()
```

### Generalized linear model

This is an exploratory fit of a poisson GLM. Initial covariates were included based on statistical significance at a Bonferroni-corrected alpha in the linear correlations shown in this report. Final covariates were chosen by forward stepwise AIC selection of additive GLMs.

```{r, bonf}
# bonferroni correction to weed out excess covariates
bonf_p <- 0.05/length(unique(abun_data$Var))
```

```{r, wrangle-abun}
abun_data <- abun_data %>%
  dplyr::mutate(Var = Var %>%
  stringr::str_replace_all(" ", "_") %>%
  stringr::str_replace_all("-", "_") %>%
  stringr::str_replace_all("/", "_") %>%
  stringr::str_replace_all("[:^:]", "_") %>%
  stringr::str_replace_all("\n", "_") %>%
    stringr::str_replace_all("\\(", "") %>%
    stringr::str_replace_all("\\)", "") 
  ) %>%
  dplyr::filter(pval < bonf_p,
                Var != "Time") %>%
  dplyr::select(-pval)

abun_data <- abun_data %>%
  tidyr::pivot_wider(names_from = Var,
                     values_from = Val) %>%
  tidyr::drop_na()

file_name <- paste(params$stock, params$region, params$epu[1],
    params$lag, "lag",
    params$remove_recent, "remove_recent", "abun_model_data",
    ".csv",
    sep = "_"
  ) %>%
    stringr::str_replace_all("/", "-") %>%
    stringr::str_replace_all("Georges Bank", "GB") %>%
    stringr::str_replace_all("Gulf of Maine", "GOM") %>%
    stringr::str_replace_all("Northern", "N") %>%
    stringr::str_replace_all("Southern", "S") %>%
    stringr::str_replace_all("Northwestern", "NW") %>%
    stringr::str_replace_all(" - ", "-") %>%
    stringr::str_replace_all("_.csv", ".csv") %>%
    stringr::str_replace_all(" ", "_")

write.csv(abun_data, paste("data/", file_name, sep = ""))


start_model <- glm(Value ~ 1,
                   data = abun_data,
                   family = poisson())

full_mod <- paste(colnames(abun_data)[6:ncol(abun_data)],
                  collapse = " + ")
full_mod <- paste("~", full_mod)



```

```{r, new-model-abun}
new_mod <- MASS::stepAIC(start_model,
              k = log(nrow(abun_data)),
              trace = FALSE,
              scope = list(lower = ~1,
                           upper = full_mod))
```

```{r, summary-mod}
dat <- summary(new_mod)$coefficients %>% 
  signif(digits = 3)

rnames <- rownames(dat) %>% stringr::str_trunc(width = 30)
dat <- cbind(rnames, dat)

knitr::kable(dat,
             row.names = FALSE,
             format.args = list(scientific = FALSE,
                                  drop0trailing = TRUE)) 
```

```{r}
all_coefs <- colnames(abun_data)[6:ncol(abun_data)] 
mod_coefs <- rownames(summary(new_mod)$coefficients)[-1]

if(length(all_coefs) == length(mod_coefs)){
  dropped_coefs <- "none"
} else {

  dropped_coefs <- all_coefs[(all_coefs %in% mod_coefs) == FALSE]
  
  if(length(dropped_coefs) > 1) {
    dropped_coefs <- paste(collapse = ",")
    print(dropped_coefs)
  }
}

```
Dropped coefficients: `r dropped_coefs`


```{r, predict-abun}
abun_data$Predicted <- predict(new_mod, newdata = abun_data) %>% exp()

fig <- ggplot2::ggplot(abun_data,
       ggplot2::aes(x = Time,
           y = Value))+
  ggplot2::geom_line()+
  ggplot2::geom_point()+
  ggplot2::geom_line(ggplot2::aes(y = Predicted),
            lty = 2,
            color = "red")+
  ggplot2::geom_point(ggplot2::aes(y = Predicted),
               color = "red")+
  ggplot2::ylab("Abundance")+
  ggplot2::xlab("Year")+
  ggplot2::theme_bw()+
  ggplot2::scale_y_continuous(labels = scales::comma,
                     limits = c(0, NA))

print(fig)
```


## Recruitment
```{r}
output %>%
  tibble::as_tibble() %>%
  dplyr::filter(
    Response_variable == "Recruitment",
    R2_adj != 1
  ) %>%
  dplyr::select(-Response_variable) %>%
  summary_table()
```

### Generalized linear model

This is an exploratory fit of a poisson GLM. Initial covariates were included based on statistical significance at a Bonferroni-corrected alpha in the linear correlations shown in this report. Final covariates were chosen by forward stepwise AIC selection of additive GLMs.

```{r, bonf-recruit}
# bonferroni correction to weed out excess covariates
bonf_p <- 0.05/length(unique(recruit_data$Var))
```

```{r, wrangle-recruit}
recruit_data <- recruit_data %>%
  dplyr::mutate(Var = Var %>%
  stringr::str_replace_all(" ", "_") %>%
  stringr::str_replace_all("-", "_") %>%
  stringr::str_replace_all("/", "_") %>%
  stringr::str_replace_all("[:^:]", "_") %>%
  stringr::str_replace("\n", "_") %>%
    stringr::str_replace_all("\\(", "") %>%
    stringr::str_replace_all("\\)", "") 
  ) %>%
  dplyr::filter(pval < bonf_p,
                Var != "Time") %>%
  dplyr::select(-pval)

recruit_data <- recruit_data %>%
  tidyr::pivot_wider(names_from = Var,
                     values_from = Val) %>%
  tidyr::drop_na() 

file_name <- paste(params$stock, params$region, params$epu[1],
    params$lag, "lag",
    params$remove_recent, "remove_recent", "recruit_model_data",
    ".csv",
    sep = "_"
  ) %>%
    stringr::str_replace_all("/", "-") %>%
    stringr::str_replace_all("Georges Bank", "GB") %>%
    stringr::str_replace_all("Gulf of Maine", "GOM") %>%
    stringr::str_replace_all("Northern", "N") %>%
    stringr::str_replace_all("Southern", "S") %>%
    stringr::str_replace_all("Northwestern", "NW") %>%
    stringr::str_replace_all(" - ", "-") %>%
    stringr::str_replace_all("_.csv", ".csv") %>%
    stringr::str_replace_all(" ", "_")

write.csv(recruit_data, paste("data/", file_name, sep = ""))

start_model <- glm(Value ~ 1,
                   data = recruit_data,
                   family = poisson())

full_mod <- paste(colnames(recruit_data)[6:ncol(recruit_data)],
                  collapse = " + ")
full_mod <- paste("~", full_mod)

```

```{r, new-model-recruit}
new_mod <- MASS::stepAIC(start_model,
              k = log(nrow(recruit_data)),
              trace = FALSE,
              scope = list(lower = ~1,
                           upper = full_mod))

dat <- summary(new_mod)$coefficients %>% 
  signif(digits = 3)

rnames <- rownames(dat) %>% stringr::str_trunc(width = 30)
dat <- cbind(rnames, dat)

knitr::kable(dat,
             row.names = FALSE,
             format.args = list(scientific = FALSE,
                                  drop0trailing = TRUE)) 
```

```{r}
all_coefs <- colnames(recruit_data)[6:ncol(recruit_data)] 
mod_coefs <- rownames(summary(new_mod)$coefficients)[-1]

if(length(all_coefs) == length(mod_coefs)){
  dropped_coefs <- "none"
} else {

  dropped_coefs <- all_coefs[(all_coefs %in% mod_coefs) == FALSE]
  
  if(length(dropped_coefs) > 1) {
    dropped_coefs <- paste(collapse = ",")
    print(dropped_coefs)
  }
}

```

Dropped coefficients: `r dropped_coefs`


```{r, predict-recruit}
recruit_data$Predicted <- predict(new_mod, newdata = recruit_data) %>% exp()

fig <- ggplot2::ggplot(recruit_data,
       ggplot2::aes(x = Time,
           y = Value))+
  ggplot2::geom_line()+
  ggplot2::geom_point()+
  ggplot2::geom_line(ggplot2::aes(y = Predicted),
            lty = 2,
            color = "red")+
  ggplot2::geom_point(ggplot2::aes(y = Predicted),
               color = "red")+
  ggplot2::ylab("Abundance")+
  ggplot2::xlab("Year")+
  ggplot2::theme_bw()+
  ggplot2::scale_y_continuous(labels = scales::comma,
                     limits = c(0, NA))

print(fig)
```


## Catch
```{r}
output %>%
  tibble::as_tibble() %>%
  dplyr::filter(
    Response_variable == "Catch",
    R2_adj != 1
  ) %>%
  dplyr::select(-Response_variable) %>%
  summary_table()
```

## Fmort
```{r}
output %>%
  tibble::as_tibble() %>%
  dplyr::filter(
    Response_variable == "Fmort",
    R2_adj != 1
  ) %>%
  dplyr::select(-Response_variable) %>%
  summary_table()
```

