# Summary of statistically significant indicators

```{r, output_summary_html, eval = (params$out == "html")}
summary_table <- function(x) {
  DT::datatable(x,
    rownames = FALSE,
    filter = list(
      position = "top",
      clear = FALSE
    ),
    options = list(
      search = list(regex = TRUE),
      paging = FALSE
    )
  )
}
```

```{r, output_summary_word, eval = (params$out == "word")}
summary_table <- function(x) {
  knitr::kable(x)
}
```

```{r}
colnames(output) <- c(
  "Response_variable", "Indicator",
  "Number of data points", "Slope",
  "P-value", "R2_adj"
)
```

## Abundance
```{r}
output %>%
  tibble::as_tibble() %>%
  dplyr::filter(
    Response_variable == "Abundance",
    R2_adj < 1
  ) %>%
  dplyr::select(-Response_variable) %>%
  summary_table()
```

### Generalized linear model

This is an exploratory fit of a poisson GLM. Initial covariates were included based on statistical significance at a Bonferroni-corrected alpha in the linear correlations shown in this report. Final covariates were chosen by forward stepwise AIC selection of additive GLMs.

```{r, bonf}
# bonferroni correction to weed out excess covariates
bonf_p <- 0.05 / bonf_n
```

```{r, wrangle-abun}
eval_dat <- (class(abun_data) != "NULL")

if (eval_dat) {
  abun_data <- abun_data %>%
    dplyr::mutate(Var = Var %>%
      stringr::str_replace_all(" ", "_") %>%
      stringr::str_replace_all("-", "_") %>%
      stringr::str_replace_all("/", "_") %>%
      stringr::str_replace_all("[:^:]", "_") %>%
      stringr::str_replace_all("\n", "_") %>%
      stringr::str_remove_all("\\(") %>%
      stringr::str_remove_all("\\)") %>%
      stringr::str_remove_all(",")) %>%
    dplyr::filter(
      pval < bonf_p,
      Var != "Time"
    ) %>%
    dplyr::select(-pval)
} else {
  "No statistically significant covariates"
}
```

```{r, eval = eval_dat}
# eval_dat
data <- abun_data
type <- "abundance"

if (nrow(data) > 0) {
  res <- knitr::knit_child(
    text = knitr::knit_expand(
      system.file("correlation_bookdown_template/_summary-child-doc.Rmd", package = "NEesp"),
      label = "abundance-model"
    ),
    quiet = TRUE
  )
  cat(res, sep = "\n")
} else {
  "No statistically significant covariates after Bonferroni correction"
}
```


## Recruitment
```{r}
output %>%
  tibble::as_tibble() %>%
  dplyr::filter(
    Response_variable == "Recruitment",
    R2_adj != 1
  ) %>%
  dplyr::select(-Response_variable) %>%
  summary_table()
```

### Generalized linear model

This is an exploratory fit of a poisson GLM. Initial covariates were included based on statistical significance at a Bonferroni-corrected alpha in the linear correlations shown in this report. Final covariates were chosen by forward stepwise AIC selection of additive GLMs.

```{r, wrangle-recruit}
eval_dat <- (class(recruit_data) != "NULL")

if (eval_dat) {
  recruit_data <- recruit_data %>%
    dplyr::mutate(Var = Var %>%
      stringr::str_replace_all(" ", "_") %>%
      stringr::str_replace_all("-", "_") %>%
      stringr::str_replace_all("/", "_") %>%
      stringr::str_replace_all("[:^:]", "_") %>%
      stringr::str_replace_all("\n", "_") %>%
      stringr::str_remove_all("\\(") %>%
      stringr::str_remove_all("\\)") %>%
      stringr::str_remove_all(",")) %>%
    dplyr::filter(
      pval < bonf_p,
      Var != "Time"
    ) %>%
    dplyr::select(-pval)
} else {
  "No statistically significant covariates"
}
```

```{r, eval = eval_dat}
# eval_dat
data <- recruit_data
type <- "recruit"

if (nrow(data) > 0) {
  res <- knitr::knit_child(
    text = knitr::knit_expand(
      system.file("correlation_bookdown_template/_summary-child-doc.Rmd", package = "NEesp"),
      label = "recruitment-model"
    ),
    quiet = TRUE
  )
  cat(res, sep = "\n")
} else {
  "No statistically significant covariates after Bonferroni correction"
}
```

## Catch
```{r}
output %>%
  tibble::as_tibble() %>%
  dplyr::filter(
    Response_variable == "Catch",
    R2_adj != 1
  ) %>%
  dplyr::select(-Response_variable) %>%
  summary_table()
```

## Fmort
```{r}
output %>%
  tibble::as_tibble() %>%
  dplyr::filter(
    Response_variable == "Fmort",
    R2_adj != 1
  ) %>%
  dplyr::select(-Response_variable) %>%
  summary_table()
```

