---
title: "`r paste(params$species_ID, 'Preliminary ESP regressions')`"
author: "Abigail Tyrell"
date: "`r format(Sys.time(), '%d %b %Y')`"
site: bookdown::bookdown_site
output: 
  bookdown::gitbook:
    split_by: section
documentclass: book
always_allow_html: true
github-repo: NOAA-EDAB/esp_data_aggregation
params:
  lag: 0 # lag for correlations
  remove_recent: FALSE
  cache: FALSE
  stock: 
  epu: MAB # epu options: GB, GOM, MAB, SS, All, OTHER, MA, NE, all 
  region: "" # the stock region
  path: "" # figure path
  save: FALSE
---

```{r setup, include = FALSE}
if (params$stock == "Monkfish") {
  stock <- assessmentdata::stockAssessmentData %>%
    dplyr::filter(
      Species == "Goosefish",
      Region == params$region
    ) %>%
    dplyr::ungroup() %>%
    dplyr::mutate(max_year = max(AssessmentYear)) %>%
    dplyr::filter(AssessmentYear == max_year) %>%
    dplyr::rename(Time = Year)
} else {
  stock <- assessmentdata::stockAssessmentData %>%
    dplyr::filter(
      Species == params$stock,
      Region == params$region
    ) %>%
    dplyr::ungroup() %>%
    dplyr::mutate(max_year = max(AssessmentYear)) %>%
    dplyr::filter(AssessmentYear == max_year) %>%
    dplyr::rename(Time = Year)
}

if (params$remove_recent) {
  max_year <- max(stock$Time)
  stock <- stock %>%
    dplyr::filter(Time <= (max_year - 10))
}

output <- c()
```

```{r, setup2, include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  message = FALSE,
  warning = FALSE,
  eval = nrow(stock) > 0,
  cache = params$cache,
  results = "asis",
  fig.width = 8,
  fig.align = "center",
  fig.path = params$path,
  dev = "png"
)
```

# `r params$stock` {-}
Stock region: `r params$region`

EPU: `r params$epu`

```{r, fail_message, eval = nrow(stock) == 0}
print("Warning! There are no data on this stock under the conditions selected.")
```

```{r, chunks}
names <- knitr::all_labels()
names <- names[-c(1:4)]
```

# Introduction

These are preliminary regressions that compare `r params$region` `r params$stock` catch, abundance, recruitment, and F to various indicators in the `r paste(params$epu, collapse = ", ")` Environmental Protection Units (EPUs) taken from the `ecodata` package. The indicators are lagged by `r params$lag` years.

```{r, fail_message, eval = nrow(stock) == 0}
```

# Regression analysis

All regressions are simple linear correlations assessed at the p < 0.5 level. Please note, due to the large number of indicators tested, a certain amount of statistically significant results are expected even if there are no underlying mechanistic connections. These correlations do not necessarily imply causation.

```{r, fail_message, eval = nrow(stock) == 0}
```
