## Catch

```{r, catch_functions, cache = params$cache}
source(here::here("R/full_report_functions", "get_assessment.R"))
source(here::here("R/full_report_functions", "get_rec_catch.R"))
source(here::here("R/full_report_functions", "get_com_catch.R"))
source(here::here("R/full_report_functions", "compare_rec_com.R"))
```

```{r, catch_raw_data_species}
asmt_data <- params$asmt_data[params$asmt_data$Species == species, ]
rec_data <- params$rec_data[params$rec_data$Species == species, ]
com_data <- params$com_data[params$com_data$Species == species, ]
```

```{r, child = here::here("bookdown", "_legend-child-doc.Rmd")}
```

Stock assessment catch data are from ``r params$asmt_source``. Recreational catch data were downloaded from [NOAA MRIP](https://www.st.nmfs.noaa.gov/st1/recreational/MRIP_Estimate_Data/CSV/). Commercial catch data were downloaded from [NOAA FOSS](https://foss.nmfs.noaa.gov/apexfoss/f?p=215:200:4615327020711::NO:::).

### Figures
```{r}
if(nrow(asmt_data %>% dplyr::filter(Metric == "Catch")) > 0){
ndesc <- asmt_data %>% 
  dplyr::filter(Metric == "Catch") %>% 
  dplyr::mutate(options = paste(Description, Units)) %>%
  dplyr::select(options) %>% 
  unique() %>% 
  nrow()
} else ndesc <- 1
```

#### Stock assessment catch
```{r, catch}
#  fig.height = ndesc * 10
plot_asmt(asmt_data, 
          metric = "Catch", 
          ytitle = "Catch",
          lin = lines, 
          col = colors)
```

##### Risk {-}

See Methods for risk calculation details.
```{r}
indicators <- c("asmt_catch")
```

```{r, child = here::here("bookdown", "_risk-child-doc.Rmd")}
```

#### Recreational catch
```{r, rec}
get_rec_catch(rec_data)
```

##### Risk {-}

See Methods for risk calculation details.
```{r}
indicators <- c("rec_catch")
```

```{r, child = here::here("bookdown", "_risk-child-doc.Rmd")}
```

#### Commercial catch
```{r, com}
plot_com(com_data)
```

##### Risk {-}

See Methods for risk calculation details.
```{r}
indicators <- c("com_catch")
```

```{r, child = here::here("bookdown", "_risk-child-doc.Rmd")}
```

#### Commercial vs recreational catch
```{r, comvrec}
prop_catch(com = com_data, rec = rec_data)
```

### Data

#### Stock assessment catch
```{r, catch_data}
data <- asmt_data %>% 
  dplyr::filter(Metric == "Catch") %>%
  dplyr::select(-Species, -Metric) %>%
  dplyr::mutate(Value = Value %>%
                  round(digits = 0)) %>%
  dplyr::mutate(Value = Value %>%
                  format(big.mark = ",")) %>%
  character_to_factor()

make_html_table(data, col_names = colnames(data))
```

#### Recreational catch
```{r, rec_data}
data <- rec_data %>% 
  dplyr::select(year, st_f, mode_fx_f, lbs_ab1) %>%
  dplyr::mutate(State = st_f %>%
                  stringr::str_to_title(),
                Method = mode_fx_f %>% 
                  stringr::str_to_sentence(),
                Catch_pounds = lbs_ab1 %>%
                  format(big.mark = ",")) %>%
  dplyr::select(year, State, Method, Catch_pounds) %>%
  character_to_factor()

make_html_table(data, col_names = c("Year", "State", "Method", "Catch (lb)"))
```

#### Commercial catch
```{r, com_data}
data <- com_data %>% 
                dplyr::select(Year, State, Dollars_adj, Pounds) %>%
                dplyr::mutate(Dollars_adj = Dollars_adj %>% 
                                round(digits = 0) %>%
                                format(big.mark = ","),
                              State = State %>% 
                                stringr::str_to_title(),
                              Pounds = Pounds %>%
                                format(big.mark = ",")) %>%
                dplyr::rename("Revenue (2019 dollars)" = "Dollars_adj",
                              "Commercial catch (lb)" = "Pounds") %>%
  character_to_factor()

make_html_table(data, col_names = colnames(data))
```

#### Commercial vs recreational catch
```{r, comvrec_data}
data <- prop_catch_data(rec = rec_data, com = com_data) %>%
  character_to_factor()

make_html_table(data, col_names = c("Year", "Recreational catch", 
                                    "Commercial catch", "Total catch", 
                                    "Proportion recreational",
                                    "Proportion commercial"))
```

#### Commercial, recreational, and stock assessment catch
```{r, all_catch}
stock <- asmt_data %>%
  dplyr::filter(Metric == "Catch",
                Region != "Eastern Georges Bank") %>% # double counted
  dplyr::mutate(options = paste(Description, Units, AssessmentYear, sep = " - ")) %>%
  dplyr::group_by(Year, options) %>%
  dplyr::summarise(Value = sum(Value)) %>%
  dplyr::ungroup() %>%
  tidyr::pivot_wider(names_from = options,
                     values_from = Value)

recr <- rec_data %>%
    dplyr::mutate(catch = lbs_ab1 %>%
                  stringr::str_replace_all(",", "") %>%
                  as.numeric()) %>%
  dplyr::group_by(year) %>%
  dplyr::rename(Year = year) %>%
  dplyr::summarise(`Recreational catch (mt)` = sum(catch)/2204.6)

comm <- com_data %>%
  dplyr::group_by(Year) %>%
  dplyr::summarise(`Commercial catch (mt)` = sum(Pounds)/2204.6)

all_catch <- c()
if(nrow(stock) > 0){
  all_catch <- stock
}
if(nrow(recr) > 0){
  if(is.null(all_catch) == FALSE){
    all_catch <- dplyr::full_join(all_catch, recr, by = "Year")
  } else all_catch <- recr
}
if(nrow(comm) > 0){
  if(is.null(all_catch) == FALSE){
    all_catch <- dplyr::full_join(all_catch, comm, by = "Year")
  } else all_catch <- comm
}

if(is.null(all_catch) == FALSE &
   ncol(all_catch > 2)){
  all_catch <- all_catch %>% 
    dplyr::arrange(Year) %>%
    round(digits = 0) %>%
    #format(big.mark = ",") %>% # messing up formatting
    as.data.frame()
}

# make sure there is data on at least 2 types of catch
test <- c(nrow(stock), nrow(recr), nrow(comm))
sum_test <- sum(test == 0)

if(sum_test >= 2){
  print("NO DATA")
} else make_html_table(all_catch, col_names = colnames(all_catch))
```

