---
title: "`r params$species_ID`"
author: "Abigail Tyrell"
date: "`r format(Sys.time(), '%d %b %Y')`"
site: bookdown::bookdown_site
output: 
  bookdown::gitbook:
    split_by: section
documentclass: book
github-repo: NOAA-EDAB/esp_data_aggregation

language:
  ui:
    chapter_name: "Section "

params: 
  species_ID: "Acadian redfish" # common name of species
  
  latlong_data: data # strata
  shape: shape # shapefile
  
  asmt_sum_data: data # bbmsy, ffmsy, assessment ratings
  asmt_sum_source: "assessmentdata::stockAssessmentSummary"

  diet_data: data # allfh diet data
  
  survey_data: data # survey data
  survey_source: "survdat"
  
  ricky_survey_data: data # ricky survey data

  rec_data: data # recreational catch (MRIP)
  
  asmt_data: data # assessmentdata:stockAssessmentData, for recruitment and catch
  asmt_source: "assessmentdata::stockAssessmentData"
  
  cond_data: data # Laurel's condition data
  
  risk_data: data # relative risk rankings
  
  com_data: data # commercial catch
  
  swept_data: data # swept area estimates
  
  risk_year_hist_data: data # historical risk by year
  
  risk_year_value_data: data # value risk by year
  
  risk_species_data: data # species risk
  
  cache: FALSE
---

# `r params$species` {-}

```{r, setup, include = FALSE}
knitr::knit_hooks$set(time_it = local({
  now <- NULL
  function(before, options) {
    if (before) {
      # record the current time before each chunk
      now <<- Sys.time()
    } else {
      # calculate the time difference after a chunk
      res <- difftime(Sys.time(), now) %>% round(digits = 2)
      # return a character string to show the time
      paste("Time for this code chunk to run:", res)
    }
  }
}))

knitr::opts_chunk$set(echo = FALSE,
                      message = FALSE, 
                      warning = FALSE, 
                      #time_it = TRUE,
                      fig.width = 8, 
                      fig.align = 'center',
                      fig.cap = params$species
                     # cache.path = here::here("bookdown/cache/", "")
                      )
```

```{r, setup2, cache = params$cache}
library(ggplot2)

`%>%` <- dplyr::`%>%`
```

```{r, universal_functions, cache = params$cache}
source(here::here("R", "character_to_factor.R"))
source(here::here("R/full_report_functions", "make_html_table.R"))

source(here::here("R/full_report_functions/risk_functions", "plot_risk_by_year.R"))
source(here::here("R/full_report_functions/risk_functions", "plot_risk_within_species.R"))
```

```{r, species_data}
# data used in multiple chapters
species <- params$species_ID

risk_year_hist_data <- params$risk_year_hist_data[params$risk_year_hist_data$Species == species, ]

risk_year_value_data <- params$risk_year_value_data[params$risk_year_value_data$Species == species, ]

risk_species_data <- params$risk_species_data[params$risk_species_data$Species == species, ]
```

This is a preliminary report of previously collected data. This report is pulling information on all Northeast `r species` stocks.

<!--chapter:end:index.Rmd-->

# Methods

## Stock identification
Northeast stocks were identified from NOAA/EDAB ECSA [seasonal species strata](https://github.com/NOAA-EDAB/ECSA/blob/master/data/seasonal_stock_strata.csv).

## Data collection and presentation

Data sources for each analysis are identified in the Results.

All continuous temporal data were plotted against time. If there were 30 or more years of data, a geom_gls regression line was fit (yellow = significant increase; purple = significant decrease; no line = no significant trend). If there were fewer than 30 years of data, no regression was fit.

### `assessmentdata` methods
Stock assessment and data quality information were compiled into a summary table.

B/Bmsy was classified as "DANGER" if it was below 1 and "GOOD" if it was above 1.

F/Fmsy was classified as "DANGER" if it was above 1 and "GOOD" if it was below 1.

### `survdat` methods
`survdat` data with zero abundance were not included in this analysis. Abundance and biomass were summed for each year and season. All other metrics were averaged for each year and season. The tables show summary statistics for the entire time series and for the most recent 5 years in the time series. 

## Risk assessment

### Risk across stocks

#### Suite of indicators
All stocks were ranked in order of increasing risk. The stock with the highest ranking is the stock determined to be at the highest risk. In this case, high risk has two meanings: (1) high importance (e.g., a stock with a high catch would have a high risk ranking for the catch indicator) or high vulnerability (e.g., a stock with low B/Bmsy would have a high risk ranking for the B/Bmsy indicator). The normalized rank was determined by dividing each stock's rank by the total number of stocks considered for that indicator. Stocks that were missing indicator measurements were assigned a normalized rank of 0.5.

#### Individual indicators
Risk was calculated over time for all indicators that were documented for five or more species in a given year. Risk was calculated as the average of the past 5 years, as a percent of the historical average. The normalized risk value was calculated as the normalized rank of this species compared to all other species in that year.

### Risk within stocks
The normalized risk value was calculated as the normalized rank of each yearly measurement compared to all other years.

<!--chapter:end:01-methods.Rmd-->

# Habitat information

<!--chapter:end:02-habitat-title-page.Rmd-->

## Distribution

```{r, geo_functions, cache = params$cache}
source(here::here("R/full_report_functions", "get_latlong.R"))
source(here::here("R/full_report_functions", "map_strata_ecsa.R"))
```

```{r, geo_data_species}
latlong_data <- params$latlong_data[params$latlong_data$Species == species, ]
```

### Map of distribution
Strata maps were pulled and compiled using code from [NOAA/EDAB ECSA](https://github.com/NOAA-EDAB/ECSA). Please note, only fall and spring strata are shown on the map.
```{r, ecsa_map}
map_strata_ecsa(data = latlong_data,
                species_name = species)
```

### Latitude and longitude ranges
Latitude and longitude ranges were calculated from NOAA/EDAB ECSA [seasonal species strata](https://github.com/NOAA-EDAB/ECSA/blob/master/data/seasonal_stock_strata.csv) and [Bottom Trawl Survey (BTS) shapefiles](https://github.com/NOAA-EDAB/ECSA/tree/master/data/strata_shapefiles). The coordinate system is WGS84.
```{r, latlong}
data <- get_latlong(species, latlong_data, shapefile = params$shape) %>%
  character_to_factor()

make_html_table(data, col_names = colnames(data))
```

<!--chapter:end:03-latlong.Rmd-->

## Temperature

```{r, temp_functions, cache = params$cache}
source(here::here("R/full_report_functions", "get_survdat_info.R"))
```

```{r, tempe_data_species}
survey_data <- params$survey_data[params$survey_data$Species == species, ]
```

Surface and bottom temperature data were pulled from ``r params$survey_source``. 

### Figures

Separate geom_gls() functions were fit for fall and spring measurements; trend lines are only shown when the trend was statistically significant, so some plots may have fewer than two trend lines. Fall has solid trend lines, and spring has dashed trend lines. Please note, sometimes the survey observed a small number of fish outside of the defined stock area.
```{r, temp}
generate_plot(survey_data,
              variable = "SURFTEMP", 
              ytitle = "Surface temperature")

generate_plot(survey_data,
              variable = "BOTTEMP", 
              ytitle = "Bottom temperature")
```

### Summary
```{r, temp_summary}
generate_table(survey_data,
              variable = "SURFTEMP", 
              cap = "Surface temperature")

generate_table(survey_data,
              variable = "BOTTEMP", 
              cap = "Bottom temperature")
```

### Data
```{r, temp_data}
data <- survey_data %>% 
                dplyr::select(YEAR, SEASON, Region, date, 
                              fish_id, SURFTEMP, BOTTEMP) %>%
                dplyr::filter(SURFTEMP > 0 | BOTTEMP > 0) %>%
                dplyr::group_by(YEAR, SEASON, Region, date, fish_id) %>%
                dplyr::distinct() %>% # remove repeated row info
                dplyr::summarise(day_mean_surf = mean(SURFTEMP),
                                 day_mean_bot = mean(BOTTEMP)) %>% # mean by day
                dplyr::ungroup() %>%
                dplyr::group_by(YEAR, SEASON, Region) %>%
                dplyr::summarise(Mean_surface_temperature = mean(day_mean_surf) %>%
                                   round(digits = 2),
                                 Mean_bottom_temperature = mean(day_mean_bot) %>%
                                   round(digits = 2)) %>%
  character_to_factor() # mean by season-year

make_html_table(data, col_names = c("Year", "Season", "Region", 
                       "Mean surface temperature", "Mean bottom temperature"))
```

<!--chapter:end:04-temperature.Rmd-->

## Salinity

```{r, sal_functions}
source(here::here("R/full_report_functions", "get_survdat_info.R"))
```

```{r,sal_data}
survey_data <- params$survey_data[params$survey_data$Species == species, ]
```

Surface and bottom salinity data were pulled from ``r params$survey_source``. 

### Figures

Separate geom_gls() functions were fit for fall and spring measurements; trend lines are only shown when the trend was statistically significant, so some plots may have fewer than two trend lines. Fall has solid trend lines, and spring has dashed trend lines. Please note, sometimes the survey observed a small number of fish outside of the defined stock area.
```{r, surfsalin}
generate_plot(survey_data,
              variable = "SURFSALIN", 
              ytitle = "Surface salinity")

generate_plot(survey_data,
              variable = "BOTSALIN", 
              ytitle = "Bottom salinity")
```

### Summary
```{r, salin_summary}
generate_table(survey_data,
              variable = "SURFSALIN", 
              cap = "Surface salinity")

generate_table(survey_data,
              variable = "BOTSALIN", 
              cap = "Bottom salinity")
```

### Data
```{r, salin_data}
data <- survey_data %>% 
                dplyr::select(YEAR, SEASON, Region, date, 
                              fish_id, SURFSALIN, BOTSALIN) %>%
                dplyr::filter(SURFSALIN > 0 | BOTSALIN > 0) %>%
                dplyr::group_by(YEAR, SEASON, Region, date, fish_id) %>%
                dplyr::distinct() %>% # remove repeated row info
                dplyr::summarise(day_mean_surf = mean(SURFSALIN),
                                 day_mean_bot = mean(BOTSALIN)) %>% # mean by day
                dplyr::ungroup() %>%
                dplyr::group_by(YEAR, SEASON, Region) %>%
                dplyr::summarise(Mean_surface_salinity = mean(day_mean_surf) %>%
                                   round(digits = 2),
                                 Mean_bottom_salinity = mean(day_mean_bot) %>%
                                   round(digits = 2)) %>%
  character_to_factor() # mean by season-year

make_html_table(data, col_names = c("Year", "Season", "Region", 
                       "Mean surface salinity", "Mean bottom salinity"))
```

<!--chapter:end:05-salinity.Rmd-->

## Depth
Ricky Tabandera

```{r, cache = params$cache}
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(here))
suppressPackageStartupMessages(library(survdat))
suppressPackageStartupMessages(library(dbutils))
```

The range of depths that a species occupies is linked to many other habitat characteristics such as benthic structure, food availability, or temperature. Thus, observed depth can signal changes in habitat suitability. Changes in this metric can indicate the required resources are changing their distribution on the landscape. Seasonal differences in occurrence can also help identify essential habitat and the timing of migration to acquire seasonal resources 

```{r, depth_data_cache, cache = params$cache}
#survdata<-readRDS(here::here("survdat_pull_bio.rds"))
#survdata <- ricky_survey 

survdata <- params$ricky_survey_data

stock_list_all_strata <- read.csv("https://raw.githubusercontent.com/NOAA-EDAB/ECSA/master/data/stock_list.csv")
stock_list_all_strata<-stock_list_all_strata %>% dplyr::rename(SVSPP=svspp)
stock_list<-stock_list_all_strata %>% dplyr::distinct(SVSPP,.keep_all =T)
```

```{r, depth_data}
survdata<-survdata %>% dplyr::mutate(SVSPP=as.numeric(SVSPP))

survdata.w.codes <- dplyr::inner_join(survdata, stock_list, by= "SVSPP" ) %>%
  dplyr::mutate(common_name = stringr::str_to_sentence(common_name))

# # abundance is adjusted but not length for bigalo data
# #groupdiscription =svspp
# # filter by group giv list of svspp codes
# get_fish_info()
# surv.species<-survdat::get_species(channel)
# 
# surv.species.1<-surv.species$data
# surv.code<-surv.species.1 %>% dplyr::distinct(SVSPP,.keep_all =T)

# black sea bass is svspp= 	141 bluefish =135

test.sp.code<-c(141,135)

selected.spp<-survdata.w.codes %>% dplyr::filter(common_name == species)

selected.spp.sum<- selected.spp %>% 
  dplyr::group_by(YEAR, SEASON) %>% 
  dplyr::summarise(ave.d=mean(DEPTH), 
            sd.d= sd(DEPTH,na.rm = TRUE),
            ave.t=mean(BOTTEMP), 
            sd.t=sd(BOTTEMP,na.rm = TRUE))

has_data <- nrow(selected.spp.sum) > 0

```

### Figures

```{r, depth_fig, eval = has_data}

selected.spp.sum %>%
  ggplot(aes(x=YEAR %>% as.numeric(), 
             y=-ave.d, 
             color=SEASON, 
             group=SEASON))+
  geom_line()+
  geom_point()+
  geom_pointrange(aes(ymin=-ave.d-sd.d, 
                      ymax=-ave.d+sd.d))+
  xlab("Year")+
  ylab("depth of trawl (ft)")+
  ggtitle("Seasonal depth profile of", 
          subtitle=species)+
  theme(axis.text.x = element_text(angle = 90))

```

```{r, depth_fig2, eval = has_data}

selected.spp.sum %>%
  ggplot(aes(x=YEAR %>% as.numeric(), 
             y=ave.t, 
             color=SEASON, 
             group=SEASON))+
  geom_line()+
  geom_point()+
  geom_pointrange(aes(ymin=ave.t-sd.t, 
                      ymax=ave.t+sd.t))+
  xlab("Year")+
  ylab("Bottom temperature (°C)")+
  ggtitle("Seasonal bottom temperature of tows that contain", 
          subtitle=species)+
  theme(axis.text.x = element_text(angle = 90))

```

```{r, no_data, eval = has_data == FALSE}
print("NO DATA")
```


<!--chapter:end:06-depth.Rmd-->

# Biological information

<!--chapter:end:07-biological-title-page.Rmd-->

## Length

```{r, len_functions, cache = params$cache}
source(here::here("R/full_report_functions", "get_length.R"))
```

```{r, len_data}
# length data (survdat) already read in with habitat data
survey_data <- params$survey_data[params$survey_data$Species == species, ]
```

Length data were pulled from ``r params$survey_source``. Only years with more than 10 fish lengths were considered for analysis. 

### Figures

Separate geom_gls() functions were fit for the minimum, mean, and maximum lengths; trend lines are only shown when the trend was statistically significant, so some plots may have fewer than three trend lines. Please note, sometimes the survey observed a small number of fish outside of the defined stock area.
```{r, length}
generate_len_plot(survey_data)
```

#### Risk {-}

See Methods for risk calculation details.
```{r}
indicators <- c("max_length_spring",
               "max_length_fall",
               "avg_length_spring",
               "avg_length_fall")
```

```{r, child = here::here("bookdown", "_risk-child-doc.Rmd")}
```

### Summary
```{r, length_summary}
generate_len_table(survey_data)
```

### Data
```{r, length_data}
data <- get_len_data2(survey_data) %>%
  character_to_factor()

make_html_table(data, col_names = c("Year", "Season", "Region", "Number of fish",
                       "Mean length", "Minimum length", "Maximum length"))
```

<!--chapter:end:08-length.Rmd-->

## von Bertalanffy growth curve

Ricky Tabandera

```{r setup_vonb, cache = params$cache}
#install.packages("ggrepel")
#install.packages("ggpubr")
#devtools::install_github("james-thorson/FishLife")

suppressPackageStartupMessages(library(FishLife))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(nlstools))
library(rfishbase)
library(ggpubr)
library(FSA)

```



```{r data_wrangling, cache = params$cache, cache.lazy = FALSE}

survdata<-readRDS(here::here("data", "survdat_pull_bio.rds"))

survdata<-survdata %>% mutate(SVSPP=as.numeric(SVSPP))
stock_list_all_strata <- read.csv("https://raw.githubusercontent.com/NOAA-EDAB/ECSA/master/data/stock_list.csv")
stock_list_all_strata<-stock_list_all_strata %>% rename(SVSPP=svspp)
stock_list<-stock_list_all_strata %>% dplyr::distinct(SVSPP,.keep_all =T)

survdata.w.codes<-inner_join(survdata, stock_list, by= "SVSPP" )

selected.surv<-survdata.w.codes %>% 
  dplyr::mutate(common_name = common_name %>%
                  stringr::str_to_sentence()) %>%
  dplyr::filter(common_name == species)


#only take records with age associated
selected.surv.clean<-selected.surv %>% drop_na(AGE)
#define the type of von b model
vb <- vbFuns(param="Typical")
#define starting parameters based on the avalible data 
f.starts<-FSA::vbStarts(LENGTH~AGE ,data = selected.surv.clean,methLinf="Walford" )

#fit a non-linear least squares model based on the data and starting values 
f.fit <- nls(LENGTH~vb(AGE,Linf,K,t0),data=selected.surv.clean, start=f.starts)
#store the fit parameters for later investigation
f.fit.summary<-summary(f.fit,correlation=TRUE)

#define the range of age values that will be used to generate points from the fitted model
#roughly by 0.2 year steps
newages<- data.frame(AGE=seq(0, 50, length = 250))
#predict(f.fit,newdata=newages) this funtion uses the model from f.fit to generate new lengths 
#make a dataset with the values from the model
selected.surv.vonb<-data.frame(AGE=seq(1, 50, length = 250), LENGTH=predict(f.fit,newdata=newages))




#########################################################




```

```{r, vonb_species}

ricky_stock <- stringr::str_to_lower(species)
#########################################################
#logical filtering of params for plotting

```

### Length at age growth curve

The predicted von Bertalanffy growth curve for NMFS managed fish species. Growth parameters of Linf (Length infinty), K (growth coefficient), and t0 (size at time 0) were estimated using non-linear least square model. The starting point for model building is accomplished using FSA::vbStarts. Age and length data sourced from Survdat and spans all years and survey areas. 

`r if(params$species == "Monkfish"){"The age determination method for monkfish has not been validated, and the anatomic structure used has changed through time. In addition these stocks display a high degree of sexual dimorphisms making fitting of single growth curves unreliable.  This results a high degree of uncertainty in assessing the age structure of the stock and the effects of fishing pressure."}`



```{r single_growth_curve}
if(nrow(selected.surv.vonb) > 0){
  selected.surv.vonb %>%
  ggplot(aes(x=AGE, y=LENGTH))+
  geom_line(color= "blue",size=1.4)+
  xlab("Age")+
  ylab(" Total length (cm)")+
  labs(color="Common name of stock")+
  ggtitle(str_to_sentence(ricky_stock), subtitle = "Length at age")+
  theme_minimal()+
  geom_point(data=selected.surv,aes(x=AGE, y=LENGTH), color="red")+
  xlim(0,(1.5*max(selected.surv$AGE, na.rm = TRUE)))+
  ylim(0,(1.5*max(selected.surv$LENGTH, na.rm = TRUE)))


} else print("NO DATA")

```

<!--chapter:end:09-von_b.Rmd-->

## Condition

```{r, cond_functions, cache = params$cache}
source(here::here("R/full_report_functions", "get_lw.R"))
source(here::here("R/full_report_functions", "get_relative_weight.R"))
```

```{r, cond_data_species}
cond_data <- params$cond_data[params$cond_data$Species == species, ]

diet_data <- params$diet_data[params$diet_data$Species == species, ]
```

Condition information comes from [diet data](https://github.com/Laurels1/Condition/blob/master/data/allfh.RData); only regions and seasons with more than 10 fish observations were considered. We calculated a rough condition factor as: Weight / Length^3, and relative weight was [previously calculated](https://github.com/Laurels1/Condition/tree/master/data).

### Figures

#### Length vs weight

Please note, no trend lines were fit, points are jittered to reduce overlap, and sometimes the survey observed a small number of fish outside of the defined stock area.
```{r, lw}
plot_lw(diet_data)
```

#### Condition factor: Weight-volume

If there were more than 30 years of data, a geom_gls() regression was fit. In order to fit the geom_gls() regression, we calculated the mean condition factor for each year and plotted the geom_gls() through those points. Please note, points are jittered to reduce overlap, and sometimes the survey observed a small number of fish outside of the defined stock area.
```{r, condition}
plot_cond(diet_data)
```

#### Condition factor: Relative weight

Please note, this data is aggregated by Ecological Protection Unit (EPU), which may differ slightly from the stock assessment regions.
```{r, laurel_cond}
plot_relw(cond_data)
```

### Data

#### Length vs weight with weight-volume condition factor
```{r, lw_data}
data <- diet_data %>%
  dplyr::filter(is.na(pdlen) == FALSE, 
                  is.na(pdwgt) == FALSE) %>%
  dplyr::select(pdlen, pdwgt, season, Region, fish_id, year) %>%
  dplyr::distinct() %>% # remove duplicates
  dplyr::group_by(Region, season) %>%
  dplyr::mutate(n_fish = length(pdlen)) %>%
  dplyr::filter(n_fish > 10) %>% # only region-season with >10 fish)
  dplyr::mutate(condition = pdwgt/(pdlen^3)) %>%
  dplyr::mutate(condition = condition %>%
                  round(digits = 4)) %>%
  dplyr::select(Region, year, season, pdlen, pdwgt, condition) %>%
  character_to_factor()

make_html_table(data, col_names = c("Region", "Year", "Season", "Fish length (cm)", 
                       "Fish weight (g)", "Condition (g/cm^3)"))
```

#### Relative weight condition factor
Please note, this data is aggregated by Ecological Protection Unit (EPU), which may differ slightly from the stock assessment regions.
```{r, laurel_cond_data}
data <- cond_data %>%
  dplyr::select(-Species, -n) %>%
  dplyr::mutate(MeanCond = MeanCond %>%
                  round(digits = 1)) %>%
  character_to_factor()

make_html_table(data, col_names = c("EPU", "Fish sex", "Year", "Mean condition",
                       "Number of fish"))
```

<!--chapter:end:10-condition.Rmd-->

## Diet

```{r, diet_functions, cache = params$cache}
source(here::here("R/full_report_functions", "get_diet.R"))
```

```{r, diet_raw_data}
diet_data <- params$diet_data[params$diet_data$Species == species, ]
```

Diet data were compiled from [existing data](https://github.com/Laurels1/Condition/blob/master/data/allfh.RData). For analysis, all geographic samples were grouped by season, year, and region, and only year-season-region combinations with more than 20 predators sampled were considered. Prey items that made up more than 5% of the predator's diet in at least one year-season-region were identified to the broad category level; all other prey are grouped into the "other" category.

### Figure
```{r, diet}
get_diet_plot(data = diet_data)
```

### Summary
```{r, diet_summary}
get_diet_table(data = diet_data)
```

### Data
```{r, diet_data}
diet_data2 <- diet_data %>%
    dplyr::filter(pyamtw > 0) %>%
    
    # only look at season/year combinations with >20 predator samples
    dplyr::group_by(year, season, Region) %>%
    dplyr::mutate(n_predators = fish_id %>% unique() %>% length()) %>%
    dplyr::filter(n_predators > 20) %>%
    dplyr::group_by(year, season, Region, n_predators, gensci) %>%
    dplyr::summarise(total_weight = sum(pyamtw)) %>%
    dplyr::mutate(proportion = total_weight/sum(total_weight),
                  proportion = proportion %>%
                    round(digits = 3),
                  total_weight = total_weight %>%
                    round(digits = 2),
                  gensci = gensci %>% 
                    stringr::str_replace("_", " ") %>%
                    stringr::str_to_sentence())  %>%
  character_to_factor()

make_html_table(diet_data2, col_names = c("Year", "Season", "Region", "Number of fish", 
                       "Prey category", "Prey weight", "Prey proportion"))
```

<!--chapter:end:11-diet.Rmd-->

# Population information

<!--chapter:end:12-population-title-page.Rmd-->

## Abundance

```{r, abun_functions, cache = params$cache}
source(here::here("R/full_report_functions", "get_assessment.R"))
source(here::here("R/full_report_functions", "get_survdat_info.R"))
source(here::here("R/full_report_functions", "get_swept.R"))
```

```{r, abun_data_species}
survey_data <- params$survey_data[params$survey_data$Species == species, ]

asmt_data <- params$asmt_data[params$asmt_data$Species == species, ]

asmt_sum_data <- params$asmt_sum_data[params$asmt_sum_data$Species == species, ]

swept_data <- params$swept_data[params$swept_data$Species == species, ]
```

```{r, child = here::here("bookdown", "_legend-child-doc.Rmd")}
```

Abundance data were pulled from ``r params$survey_source`` and ``r params$asmt_source``. 

### Figures

Separate geom_gls() functions were fit for fall and spring measurements; trend lines are only shown when the trend was statistically significant, so some plots may have fewer than two trend lines. Fall has solid trend lines, and spring has dashed trend lines. Please note, sometimes the survey observed a small number of fish outside of the defined stock area.

#### Survey abundance (raw measurements)
```{r, abun_survey}
generate_plot(survey_data,
              variable = "ABUNDANCE", 
              ytitle = "Abundance")
```

##### Risk {-}

See Methods for risk calculation details.
```{r}
indicators <- c("abundance_spring", 
                "abundance_fall")
```

```{r, child = here::here("bookdown", "_risk-child-doc.Rmd")}
```

#### Survey abundance (swept area estimates)

Please note, these estimates are not parsed by region or season. Swept area estimates are based on spring and fall surveys only. The shaded gray region indicates +/- two standard errors.
```{r, abun_survey_swept}
plot_swept(swept_data, var = "abundance")
```

#### Assessment abundance
```{r}
if(nrow(asmt_data %>% dplyr::filter(Metric == "Abundance")) > 0){
ndesc <- asmt_data %>% 
  dplyr::filter(Metric == "Abundance") %>% 
  #dplyr::mutate(options = paste(Description, Units)) %>%
  dplyr::select(Age) %>% 
  unique() %>% 
  nrow()
} else ndesc <- 1
```

```{r, abun_asmt, fig.height = ndesc * 4}
plot_asmt(asmt_data, 
          metric = "Abundance", 
          ytitle = "Abundance",
          lin = lines, 
          col = colors)
```

##### Risk {-}

See Methods for risk calculation details.
```{r}
indicators <- c("asmt_abundance")
```

```{r, child = here::here("bookdown", "_risk-child-doc.Rmd")}
```

### Survey summary
```{r, abun_summary}
generate_table(survey_data,
              variable = "ABUNDANCE", 
              cap = "Measured abundance")
```

### Data

#### Survey data (raw measurements)
```{r, surv_abun_data}
data <- survey_data %>%
  get_var_data(variable = "ABUNDANCE") %>%
  dplyr::mutate(variable = variable %>%
                  format(big.mark = ","))  %>%
  character_to_factor()

make_html_table(data, col_names = c("Year", "Season", "Region", "Abundance"))
```

#### Survey data (swept area estimates)
```{r, swept_abun_data}
data <- swept_data %>%
  dplyr::select(YEAR, tot.abundance, tot.abund.SE) %>%
  dplyr::mutate(YEAR = as.numeric(YEAR),
                tot.abundance = tot.abundance %>%
                  format(big.mark = ","),
                tot.abund.SE = tot.abund.SE %>%
                  round(digits = 1) %>%
                  format(big.mark = ","))

make_html_table(data, col_names = c("Year", "Abundance", "Standard error"))
```

#### Assessment data
```{r, asmt_abun_data}
data <- asmt_data %>% 
                dplyr::filter(Metric == "Abundance") %>%
                dplyr::select(-Species, -Metric) %>%
                dplyr::mutate(Value = Value %>%
                                round(digits = 0)) %>%
                dplyr::mutate(Value = Value %>%
                              format(big.mark = ",")) %>%
  character_to_factor()

make_html_table(data, col_names = colnames(data))
```

<!--chapter:end:13-abundance.Rmd-->

```{r, biomass_functions, cache = params$cache}
source(here::here("R/full_report_functions", "get_assessment.R"))
source(here::here("R/full_report_functions", "get_survdat_info.R"))
source(here::here("R/full_report_functions", "get_swept.R"))
```

```{r, biomass_raw_data}
survey_data <- params$survey_data[params$survey_data$Species == species, ]

asmt_data <- params$asmt_data[params$asmt_data$Species == species, ]

swept_data <-  params$swept_data[ params$swept_data$Species == species, ]
```

```{r, child = here::here("bookdown", "_legend-child-doc.Rmd")}
```

## Biomass

Biomass data were pulled from ``r params$survey_source``. 

### Figures

Separate geom_gls() functions were fit for fall and spring measurements; trend lines are only shown when the trend was statistically significant, so some plots may have fewer than two trend lines. Fall has solid trend lines, and spring has dashed trend lines. Please note, sometimes the survey observed a small number of fish outside of the defined stock area.

#### Survey biomass (raw measurements)
```{r, biomass}
generate_plot(survey_data,
              variable = "BIOMASS", 
              ytitle = "Biomass")
```

##### Risk {-}

See Methods for risk calculation details.
```{r}
indicators <- c("biomass_spring", 
                "biomass_fall")
```

```{r, child = here::here("bookdown", "_risk-child-doc.Rmd")}
```

#### Survey biomass (swept area estimates)

Please note, these estimates are not parsed by region or season. Swept area estimates are based on spring and fall surveys only. The shaded gray region indicates +/- two standard errors.
```{r, bio_survey_swept}
plot_swept(swept_data, var = "biomass")
```

#### Assessment biomass
```{r}
if(nrow(asmt_data %>% dplyr::filter(Metric == "Biomass")) > 0){
ndesc <- asmt_data %>% 
  dplyr::filter(Metric == "Biomass") %>% 
  #dplyr::mutate(options = paste(Description, Units)) %>%
  dplyr::select(Age) %>% 
  unique() %>% 
  nrow()
} else ndesc <- 1
```

```{r, biomass_asmt,  fig.height = ndesc * 4 + 1}
plot_asmt(asmt_data, 
          metric = "Biomass", 
          ytitle = "Biomass",
          lin = lines, 
          col = colors)
```

##### Risk {-}

See Methods for risk calculation details.
```{r}
indicators <- c("asmt_biomass")
```

```{r, child = here::here("bookdown", "_risk-child-doc.Rmd")}
```

### Survey summary
```{r, biomass_summary}
generate_table(survey_data,
              variable = "BIOMASS", 
              cap = "Measured biomass")
```

### Data

#### Survey data (raw measurements)
```{r, biomass_data}
data <- survey_data %>% 
  get_var_data(variable = "BIOMASS") %>%
  dplyr::mutate(variable = variable %>%
                     format(big.mark = ",")) %>%
  character_to_factor()

make_html_table(data, col_names = c("Year", "Season", "Region", "Biomass"))
```

#### Survey data (swept area estimates)
```{r, swept_biomass_data}
data <- swept_data %>%
  dplyr::select(YEAR, tot.biomass, tot.bio.SE) %>%
  dplyr::mutate(YEAR = as.numeric(YEAR),
                tot.biomass = tot.biomass %>%
                  format(big.mark = ","),
                tot.bio.SE = tot.bio.SE %>%
                  round(digits = 1) %>%
                  format(big.mark = ","))

make_html_table(data, col_names = c("Year", "Biomass (kg)", "Standard error"))
```

#### Assessment data
```{r, asmt_biomass_data}
data <- asmt_data %>%
  dplyr::filter(Metric == "Biomass") %>%
  dplyr::select(-Species, -Metric) %>%
  dplyr::mutate(Value = Value %>%
                  round(digits = 0)) %>%
  dplyr::mutate(Value = Value %>%
                  format(big.mark = ",")) %>%
  character_to_factor()

make_html_table(data, col_names = colnames(data))
```

<!--chapter:end:14-biomass.Rmd-->


```{r, b_functions, cache = params$cache}
source(here::here("R/full_report_functions", "get_bf.R"))
```

```{r, b_data_species}
asmt_sum_data <- params$asmt_sum_data[params$asmt_sum_data$Species == species, ]
```

```{r, b_status}
if(length(asmt_sum_data$Species) > 1){
  list_regions <- split(unique(asmt_sum_data$Region), 
                        f = list(unique(asmt_sum_data$Region)))
  bbmsy2 <- purrr::map(list_regions, 
                       ~status(data = asmt_sum_data, 
                               regions = .x, 
                               metric = "bbmsy"))
} else bbmsy2 <- "UNKNOWN"

```

```{r, child = here::here("bookdown", "_legend-child-doc.Rmd")}
```

## B/Bmsy 

B/Bmsy data were pulled from ``r params$asmt_sum_source``.

The most recent status of B/Bmsy is: `r bbmsy2`

### Figure
```{r, bbmsy_fig} 
if("GOOD" %in% bbmsy2 | "DANGER" %in% bbmsy2){
  plot_msy(asmt_sum_data[which(is.na(asmt_sum_data$`B/Bmsy`) == FALSE), ], 
           type = "b",
             ytitle = "B / Bmsy",
             lin = lines,
             col = colors)
} else print("NO DATA")
```

##### Risk {-}

See Methods for risk calculation details.
```{r}
indicators <- c("bbmsy")
```

```{r, child = here::here("bookdown", "_risk-child-doc.Rmd")}
```

### Data
```{r, bbmsy_data}
data <- asmt_sum_data[which(is.na(asmt_sum_data$`B/Bmsy`) == FALSE), ] %>% 
                dplyr::select(Region, `Assessment Year`, `B/Bmsy`) %>%
                dplyr::mutate(`B/Bmsy` = `B/Bmsy` %>% round(digits = 2)) %>%
  character_to_factor()

make_html_table(data, col_names = c("Region", "Year", "B/Bmsy"))
```

<!--chapter:end:15-bbmsy.Rmd-->

## Recruitment

```{r, recruit_functions, cache = params$cache}
source(here::here("R/full_report_functions", "get_assessment.R"))
```

```{r, recruit_raw_data}
asmt_data <- params$asmt_data[params$asmt_data$Species == species, ]
```

```{r, child = here::here("bookdown", "_legend-child-doc.Rmd")}
```

Recruitment data were pulled from ``r params$asmt_source``. Separate geom_gls() functions were fit for each region; trend lines are only shown when the trend was statistically significant, so some plots may have fewer trend lines than regions.

### Figure
```{r}
if(nrow(asmt_data %>% dplyr::filter(Metric == "Recruitment")) > 0){
ndesc <- asmt_data %>% 
  dplyr::filter(Metric == "Recruitment") %>% 
  #dplyr::mutate(options = paste(Description, Units)) %>%
  dplyr::select(Age) %>% 
  unique() %>% 
  nrow()
} else ndesc <- 1
```

```{r, recruitment, fig.height = ndesc * 4 + 1}
plot_asmt(asmt_data, 
          metric = "Recruitment", 
          ytitle = "Recruitment",
          lin = lines, 
          col = colors)
```

##### Risk

See Methods for risk calculation details.
```{r}
indicators <- c("recruitment")
```

```{r, child = here::here("bookdown", "_risk-child-doc.Rmd")}
```

### Data

```{r, recruit_data}
data <- asmt_data %>% 
  dplyr::filter(Metric == "Recruitment") %>%
  dplyr::select(-Species, -Metric) %>%
  dplyr::mutate(Value = Value %>%
                  round(digits = 0)) %>%
  dplyr::mutate(Value = Value %>%
                  format(big.mark = ",",
                         scientific = FALSE)) %>%
  character_to_factor()

make_html_table(data, col_names = colnames(data))
```

<!--chapter:end:16-recruitment.Rmd-->

# Socio-economic information

<!--chapter:end:17-socio-economic-title-page.Rmd-->

## Catch

```{r, catch_functions, cache = params$cache}
source(here::here("R/full_report_functions", "get_assessment.R"))
source(here::here("R/full_report_functions", "get_rec_catch.R"))
source(here::here("R/full_report_functions", "get_com_catch.R"))
source(here::here("R/full_report_functions", "compare_rec_com.R"))
```

```{r, catch_raw_data_species}
asmt_data <- params$asmt_data[params$asmt_data$Species == species, ]
rec_data <- params$rec_data[params$rec_data$Species == species, ]
com_data <- params$com_data[params$com_data$Species == species, ]
```

```{r, child = here::here("bookdown", "_legend-child-doc.Rmd")}
```

Stock assessment catch data are from ``r params$asmt_source``. Recreational catch data were downloaded from [NOAA MRIP](https://www.st.nmfs.noaa.gov/st1/recreational/MRIP_Estimate_Data/CSV/). Commercial catch data were downloaded from [NOAA FOSS](https://foss.nmfs.noaa.gov/apexfoss/f?p=215:200:4615327020711::NO:::).

### Figures
```{r}
if(nrow(asmt_data %>% dplyr::filter(Metric == "Catch")) > 0){
ndesc <- asmt_data %>% 
  dplyr::filter(Metric == "Catch") %>% 
  dplyr::mutate(options = paste(Description, Units)) %>%
  dplyr::select(options) %>% 
  unique() %>% 
  nrow()
} else ndesc <- 1
```

#### Stock assessment catch
```{r, catch}
#  fig.height = ndesc * 10
plot_asmt(asmt_data, 
          metric = "Catch", 
          ytitle = "Catch",
          lin = lines, 
          col = colors)
```

##### Risk {-}

See Methods for risk calculation details.
```{r}
indicators <- c("asmt_catch")
```

```{r, child = here::here("bookdown", "_risk-child-doc.Rmd")}
```

#### Recreational catch
```{r, rec}
get_rec_catch(rec_data)
```

##### Risk {-}

See Methods for risk calculation details.
```{r}
indicators <- c("rec_catch")
```

```{r, child = here::here("bookdown", "_risk-child-doc.Rmd")}
```

#### Commercial catch
```{r, com}
plot_com(com_data)
```

##### Risk {-}

See Methods for risk calculation details.
```{r}
indicators <- c("com_catch")
```

```{r, child = here::here("bookdown", "_risk-child-doc.Rmd")}
```

#### Commercial vs recreational catch
```{r, comvrec}
prop_catch(com = com_data, rec = rec_data)
```

### Data

#### Stock assessment catch
```{r, catch_data}
data <- asmt_data %>% 
  dplyr::filter(Metric == "Catch") %>%
  dplyr::select(-Species, -Metric) %>%
  dplyr::mutate(Value = Value %>%
                  round(digits = 0)) %>%
  dplyr::mutate(Value = Value %>%
                  format(big.mark = ",")) %>%
  character_to_factor()

make_html_table(data, col_names = colnames(data))
```

#### Recreational catch
```{r, rec_data}
data <- rec_data %>% 
  dplyr::select(year, st_f, mode_fx_f, lbs_ab1) %>%
  dplyr::mutate(State = st_f %>%
                  stringr::str_to_title(),
                Method = mode_fx_f %>% 
                  stringr::str_to_sentence(),
                Catch_pounds = lbs_ab1 %>%
                  format(big.mark = ",")) %>%
  dplyr::select(year, State, Method, Catch_pounds) %>%
  character_to_factor()

make_html_table(data, col_names = c("Year", "State", "Method", "Catch (lb)"))
```

#### Commercial catch
```{r, com_data}
data <- com_data %>% 
                dplyr::select(Year, State, Dollars_adj, Pounds) %>%
                dplyr::mutate(Dollars_adj = Dollars_adj %>% 
                                round(digits = 0) %>%
                                format(big.mark = ","),
                              State = State %>% 
                                stringr::str_to_title(),
                              Pounds = Pounds %>%
                                format(big.mark = ",")) %>%
                dplyr::rename("Revenue (2019 dollars)" = "Dollars_adj",
                              "Commercial catch (lb)" = "Pounds") %>%
  character_to_factor()

make_html_table(data, col_names = colnames(data))
```

#### Commercial vs recreational catch
```{r, comvrec_data}
data <- prop_catch_data(rec = rec_data, com = com_data) %>%
  character_to_factor()

make_html_table(data, col_names = c("Year", "Recreational catch", 
                                    "Commercial catch", "Total catch", 
                                    "Proportion recreational",
                                    "Proportion commercial"))
```

#### Commercial, recreational, and stock assessment catch
```{r, all_catch}
stock <- asmt_data %>%
  dplyr::filter(Metric == "Catch",
                Region != "Eastern Georges Bank") %>% # double counted
  dplyr::mutate(options = paste(Description, Units, AssessmentYear, sep = " - ")) %>%
  dplyr::group_by(Year, options) %>%
  dplyr::summarise(Value = sum(Value)) %>%
  dplyr::ungroup() %>%
  tidyr::pivot_wider(names_from = options,
                     values_from = Value)

recr <- rec_data %>%
    dplyr::mutate(catch = lbs_ab1 %>%
                  stringr::str_replace_all(",", "") %>%
                  as.numeric()) %>%
  dplyr::group_by(year) %>%
  dplyr::rename(Year = year) %>%
  dplyr::summarise(`Recreational catch (mt)` = sum(catch)/2204.6)

comm <- com_data %>%
  dplyr::group_by(Year) %>%
  dplyr::summarise(`Commercial catch (mt)` = sum(Pounds)/2204.6)

all_catch <- c()
if(nrow(stock) > 0){
  all_catch <- stock
}
if(nrow(recr) > 0){
  if(is.null(all_catch) == FALSE){
    all_catch <- dplyr::full_join(all_catch, recr, by = "Year")
  } else all_catch <- recr
}
if(nrow(comm) > 0){
  if(is.null(all_catch) == FALSE){
    all_catch <- dplyr::full_join(all_catch, comm, by = "Year")
  } else all_catch <- comm
}

if(is.null(all_catch) == FALSE &
   ncol(all_catch > 2)){
  all_catch <- all_catch %>% 
    dplyr::arrange(Year) %>%
    round(digits = 0) %>%
    #format(big.mark = ",") %>% # messing up formatting
    as.data.frame()
}

# make sure there is data on at least 2 types of catch
test <- c(nrow(stock), nrow(recr), nrow(comm))
sum_test <- sum(test == 0)

if(sum_test >= 2){
  print("NO DATA")
} else make_html_table(all_catch, col_names = colnames(all_catch))
```


<!--chapter:end:18-catch.Rmd-->

## F/Fmsy 

```{r, f_functions, cache = params$cache}
source(here::here("R/full_report_functions", "get_bf.R")) 
```

```{r, f_data}
asmt_sum_data <- params$asmt_sum_data[params$asmt_sum_data$Species == species, ]
```

```{r, f_status}
if(length(asmt_sum_data$Species) > 1){
  list_regions <- split(unique(asmt_sum_data$Region), 
                        f = list(unique(asmt_sum_data$Region)))
  ffmsy2 <- purrr::map(list_regions, 
                       ~status(data = asmt_sum_data, 
                               regions = .x, 
                               metric = "ffmsy"))
} else ffmsy2 <- "UNKNOWN"

```

```{r, child = here::here("bookdown", "_legend-child-doc.Rmd")}
```

F/Fmsy data were pulled from ``r params$asmt_sum_source``.

The most recent status of F/Fmsy is: `r ffmsy2`

### Figure
```{r, ffmsy}
if("GOOD" %in% ffmsy2 | "DANGER" %in% ffmsy2){
 plot_msy(asmt_sum_data[which(is.na(asmt_sum_data$`F/Fmsy`) == FALSE), ],
          ytitle = "F / Fmsy",
          type = "f",
             lin = lines,
             col = colors)
} else print("NO DATA")
```

##### Risk {-}

See Methods for risk calculation details.
```{r}
indicators <- c("ffmsy")
```

```{r, child = here::here("bookdown", "_risk-child-doc.Rmd")}
```

### Data
```{r, ffmsy_data}
data <- asmt_sum_data[which(is.na(asmt_sum_data$`F/Fmsy`) == FALSE), ] %>% 
                dplyr::select(Region, `Assessment Year`, `F/Fmsy`) %>%
                dplyr::mutate(`F/Fmsy` = `F/Fmsy` %>% round(digits = 2)) %>%
  character_to_factor()

make_html_table(data, col_names = c("Region", "Year", "F/Fmsy"))
```

<!--chapter:end:19-ffmsy.Rmd-->

## Revenue 

```{r, revenue_functions, cache = params$cache}
source(here::here("R/full_report_functions", "get_com_money.R"))
```

```{r, revenue_raw_data}
com_data <- params$com_data[params$com_data$Species == species, ]
```

Commercial catch data were downloaded from [NOAA FOSS](https://foss.nmfs.noaa.gov/apexfoss/f?p=215:200:4615327020711::NO:::).

### Figure
```{r, revenue}
plot_com_money(com_data)
```

##### Risk {-}

See Methods for risk calculation details.
```{r}
indicators <- c("revenue")
```

```{r, child = here::here("bookdown", "_risk-child-doc.Rmd")}
```

### Data
```{r, revenue_data}
data <- com_data %>% 
                dplyr::select(Year, State, Dollars_adj, Pounds) %>%
                dplyr::mutate(Dollars_adj = Dollars_adj %>% 
                                round(digits = 0) %>%
                                format(big.mark = ","),
                              State = State %>% 
                                stringr::str_to_title(),
                              Pounds = Pounds %>%
                                format(big.mark = ",")) %>%
                dplyr::rename("Revenue (2019 dollars)" = "Dollars_adj",
                              "Commercial catch (lb)" = "Pounds") %>%
  character_to_factor()

make_html_table(data, col_names = colnames(data))
```

<!--chapter:end:20-revenue.Rmd-->

# Management information

```{r, management_functions}
# none
```

```{r, management_data, cache = params$cache}
asmt_sum_data <- params$asmt_sum_data[params$asmt_sum_data$Species == species, ]

ad_rating_data <- params$asmt_sum_data %>% 
  dplyr::select(Species, Region, `Assessment Year`, `Last Data Year`,
                `Biological Input Data`, `Life History Data`, 
                `Composition Input Data`, `Ecosystem Linkage`, `FSSI Stock?`,
                `Review Result`) %>%
  dplyr::rename(`Biological Data Rating (2019)` = `Biological Input Data`,
                `Biological Data Rating (before 2019)` = `Life History Data`,
                `Size Data Rating` = `Composition Input Data`,
                `Ecosystem Linkage Data Rating` = `Ecosystem Linkage`,
                `FSSI status` = `FSSI Stock?`)
```

## Stock assessment and data quality information
Stock assessment and data quality information were pulled from ``r params$asmt_sum_source``.
```{r, quality}
data <- ad_rating_data %>%
  dplyr::filter(Species == species) %>%
  character_to_factor()

make_html_table(data, col_names = colnames(data))
```

<!--chapter:end:21-management.Rmd-->

# Vulnerability

```{r, vul_functions}
source(here::here("R/full_report_functions", "plot_vulnerability.R"))
```

## Climate vulnerability

Climate vulnerability is sourced from Hare et al. (2016).

```{r, clim_vul}
data <- read.csv(url("https://doi.org/10.1371/journal.pone.0146756.s001")) %>%
  dplyr::mutate(Species = Species %>%
                  stringr::str_to_sentence()) %>%
  dplyr::filter(Species == species) %>%
  dplyr::select(-Species)
```

### Figures
```{r, fig.height = 10}
plot_climate_vulnerability(data)
```

### Data
```{r}
make_html_table(data,
                col_names = c("Functional group", "Attribute", "Attribute category", 
                             "Expert score: low",
                             "Expert score: moderate",
                             "Expert score: high",
                             "Expert score: very high"))
```

## Habitat vulnerability

Habitat information is sourced from the `ecodata` package. 

```{r, hab_vul}
data <- ecodata::habitat_vulnerability %>%
  dplyr::filter(Species == species) %>%
  dplyr::select(-Species)

data <- data[,c(8, 1:7)]

make_html_table(data,
                col_names = c("EPU", "Habitat name", "Habitat vulnerability", 
                             "Importance of habitat to eggs/larvae",
                             "Importance of habitat to juveniles/YOY",
                             "Importance of habitat to adults",
                             "Importance of habitat to spawning adults",
                             "Overall species vulnerability"))

```

<!--chapter:end:22-vulnerability.Rmd-->

# Risk assessment

```{r, risk_functions, cache = params$cache}
source(here::here("R/full_report_functions/risk_functions", "plot_risk_within_species.R"))
source(here::here("R/full_report_functions/risk_functions", "plot_risk_by_year.R"))
```

```{r, risk_data}
# don't use?
risk_data <- params$risk_data[params$risk_data$Species == species, ]

risk_year_hist_data <- params$risk_year_hist_data[params$risk_year_hist_data$Species == species, ]

risk_year_value_data <- params$risk_year_value_data[params$risk_year_value_data$Species == species, ]

risk_species_data <- params$risk_species_data[params$risk_species_data$Species == species, ]
```

## Preliminary risk calculation

A preliminary risk analysis was conducted by ranking all species according to their indicator values. A high rank number and a normalized rank near 1 indicates that the species is at risk or of importance based on the measured indicator values. When a species was missing an indicator, it was assigned a normalized rank of 0.5.

### Relative to all other stocks
```{r, risk_all}
data <- risk_year_hist_data %>% 
  dplyr::mutate(Rank = paste(rank, "out of", n_stocks_per_indicator),
                       Value2 = Value %>% 
                         format(big.mark = ",", scientific = FALSE, digits = 2),
                       Normalized_rank = norm_rank %>% round(digits = 2)) %>%
  dplyr::select(Region, category, Indicator, Year, Value2, Rank, Normalized_rank) %>%
  dplyr::rename("Value" = "Value2") %>%
  character_to_factor()

make_html_table(data, col_names = c("Region", "Category", "Indicator", "Year", "Value (% of historical)",
                       "Rank", "Normalized rank"))

data <- risk_year_value_data %>% 
  dplyr::mutate(Rank = paste(rank, "out of", n_stocks_per_indicator),
                       Value2 = Value %>% 
                         format(big.mark = ",", scientific = FALSE, digits = 2),
                       Normalized_rank = norm_rank %>% round(digits = 2)) %>%
  dplyr::select(Region, category, Indicator, Year, Value2, Rank, Normalized_rank) %>%
  dplyr::rename("Value" = "Value2") %>%
  character_to_factor()

make_html_table(data, col_names = c("Region", "Category", "Indicator", "Year", "Value (magnitude)",
                       "Rank", "Normalized rank"))
```

### Within each stock
```{r, risk_within}
data <- risk_species_data %>% dplyr::mutate(Rank = paste(rank, "out of", n_years_per_indicator),
                       Value2 = Value %>% 
                         format(big.mark = ",", scientific = FALSE, digits = 2),
                       Normalized_rank = norm_rank %>% round(digits = 2)) %>%
  dplyr::select(Region, category, Indicator, Year, Value2, Rank, Normalized_rank) %>%
  dplyr::rename("Value" = "Value2") %>%
  character_to_factor()

make_html_table(data, col_names = c("Region", "Category", "Indicator", "Year", "Value",
                       "Rank", "Normalized rank"))
```

## Preliminary risk visualization

### Relative to all other stocks

Risk was calculated over time for all indicators that were documented for five or more species in a given year. Risk was calculated as the average of the past 5 years, as a percent of the historical average. The normalized risk value plotted here reflects the normalized rank of this stock compared to all other stocks in that year.

```{r, year_risk}
plot_risk_by_year(risk_year_hist_data,
                  indicator = "all",
                  title = "Change compared to historical",
                  include_legend = "yes")

plot_risk_by_year(risk_year_value_data,
                  indicator = "all",
                  title = "Value compared to other stocks",
                  include_legend = "yes")
```

### Within a single stock

For each stock, a five-year running mean was calculated for each indicator. Indicator values were then ranked for all years where a value was present. The normalized risk values plotted here reflects the normalized rank of each year compared to all other years.

```{r, stock_risk}
plot_risk_by_stock(risk_species_data,
                  indicator = "all",
                  include_legend = "yes")
```

<!--chapter:end:23-risk-assessment.Rmd-->

