---
author: "Ricky Tabandera"
date: "`r format(Sys.time(), '%d %b %Y')`"
output: html_document
params:
  stock:
    label: "Stock:"
    value: Alewife
    input: select
    choices: [atlantic hagfish, smooth dogfish, spiny dogfish, barndoor skate, winter skate, clearnose skate, rosette skate, little skate, smooth skate, thorny skate, atlantic herring, alewife, blueback herring, offshore hake, silver hake, atlantic cod, haddock, pollock, white hake, red hake, cusk, atlantic halibut, american plaice, summer flounder, yellowtail flounder, winter flounder, witch flounder, atlantic mackerel, butterfish, bluefish, black sea bass, scup, acadian redfish, atlantic wolffish, ocean pout, monkfish, american lobster, northern shrimp, northern shortfin squid, longfin inshore squid, atlantic menhaden, windowpane flounder ]
  
title: "Commercial landings of `r stringr::str_to_sentence(params$stock)`"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

```{r data processing, message= FALSE, error=FALSE,warning = FALSE, echo=FALSE}
#list.files(here::here("commercial"))


#load in the stocks managed in NEFSC
stock_list_all_strata <- read.csv("https://raw.githubusercontent.com/NOAA-EDAB/ECSA/master/data/stock_list.csv")
#lowercase all of the names
stock_list_all_strata$common_name <- stringr::str_to_lower(stock_list_all_strata$common_name)
#extract out the unique names and rename column  
stock_list<-stock_list_all_strata %>% distinct(common_name, .keep_all=TRUE)


#load commercial landings data
com_landings<-readr::read_csv(here::here("commercial","foss_ commercial_landings_NE_MA_1950_2019.csv"))
#renaming cols to aling with stocklist
com_landings<-com_landings %>% rename(common_name=`NMFS Name`)
#simplifying the names to be lowercase 
com_landings$common_name<-stringr::str_to_lower(com_landings$common_name)

#removing empty rows due to confidentiality and counting the number of records removed
confid<-com_landings %>% 
  filter(common_name =="withheld for confidentiality" | Confidentiality == "Confidential" )%>%
  group_by(Year) %>%
  summarise(count=n())
com_landings<-com_landings %>% filter(!common_name =="withheld for confidentiality" , !Confidentiality == "Confidential", !is.na(Pounds), !is.na(Dollars), Dollars >1)


# removing aggrigated landings records and counting their occurance
grouped_landings<-com_landings %>% filter(stringr::str_detect(common_name, "\\*+"))

# rearranging the species names to align with NEFSC conventions
com_landings<-com_landings %>% filter(!stringr::str_detect(common_name, "\\*+"))

#splitting off the records without a comma


single_name<-com_landings %>% filter(!stringr::str_detect(common_name, "\\,\\s"))
multipe_name<-com_landings %>% filter(stringr::str_detect(common_name, "\\,\\s"))
split_names<-multipe_name%>% separate (col=common_name,sep ="\\,\\s", into = c("first" ,"second", "third"))
#fixing plaice to remove flounder
split_names<- split_names %>% dplyr::mutate(first=replace(first, which(second=="american plaice"), NA))
#fixing the shark tag for dogfish
split_names<- split_names %>% dplyr::mutate(first=replace(first, which(second=="dogfish"), NA))
#longfin loligo into inshore
split_names<- split_names %>% dplyr::mutate(second=replace(second, which(second=="longfin loligo"), "longfin inshore"))
#shortfin illex into northern shortfin
split_names<- split_names %>% dplyr::mutate(second=replace(second, which(second=="shortfin illex"), "northern shortfin"))
#goosefish into monkfish
single_name<- single_name %>% dplyr::mutate(common_name=replace(common_name, which(common_name=="goosefish"), "monkfish"))
#adding flounder to windowpane 
single_name<- single_name %>% dplyr::mutate(common_name=replace(common_name, which(common_name=="windowpane"), "windowpane flounder"))
split_names<-split_names %>% unite(col = "last", c(third, second),sep = " ", remove = TRUE ,na.rm = TRUE)
split_names<-split_names %>% unite(col = "common_name", c(last, first),sep = " ", remove = TRUE ,na.rm = TRUE)

joined_names<- dplyr::bind_rows(split_names,single_name)
#used to check that there are no missing stocks in the data
#missing_stocks<-dplyr::anti_join(stock_list,joined_names, by= "common_name")

#final clean dataset
com_landings_clean<-dplyr::full_join(stock_list,joined_names, by= "common_name")

#write.csv(com_landings_clean, "com_landings_clean_20201222.csv")
selected_com_sp<-com_landings_clean %>%
  dplyr::group_by(Year) %>%
  filter(common_name==params$stock) %>%
  summarise(total_pounds=sum(Pounds,na.rm = FALSE),total_dollars=sum(Dollars,na.rm = FALSE))

selected_com_state<-com_landings_clean %>%
  dplyr::group_by(Year,State) %>%
  filter(common_name==params$stock) %>%
  summarise(total_pounds=sum(Pounds,na.rm = FALSE),total_dollars=sum(Dollars,na.rm = FALSE))

##########################

#red hake
data_qual<-TRUE

num_states<-selected_com_state %>% group_by( State) %>% count()

if(min(num_states$n)<5){data_qual<-FALSE
} else{data_qual<-TRUE}





```


## Commercial Landings of `r params$stock` 

This data was sourced from [FOSS - Fisheries One Stop Shop](https://foss.nmfs.noaa.gov/apexfoss/f?p=215:200:4615327020711::NO:::) and is inclusive of `r min(com_landings$Year) ` to `r max(com_landings$Year)`. The entire data set contains `r length(unique(com_landings$common_name))` species. A total of `r length(grouped_landings$common_name) ` aggrigated records were removed. These landing reports were aggrigations of mutiple species and can span several taxanomic groups were removied due to their lack of specificity. `r sum(confid$count) ` records were withheld due to confidentiality of thoese specific landings.
Figures produced reflect the coverage of the data. With stocks that have high coverage, a running average is calculated. In low coverage stocks, missing values are excluded and a simple time series is produced. 


```{r landings of single stock, warning= FALSE, message= FALSE, error= FALSE, echo=FALSE}


selected_com_sp %>% 
  ggplot2::ggplot(aes(x=Year, y=(total_dollars/10^6)))+
  ggplot2::geom_point()+
  tidyquant::geom_ma( n=5, lwd=1.4)+
  ggtitle(label = params$stock ,subtitle = "5 year rolling average")+
  xlab(label= "Year")+
  ylab(label = "Dollar value of landings in millions ($) ")+
  theme_minimal()

```

## Commercial landings of `r params$stock` across states in the mid-Atlantic and north-east regions

```{r landings across states, for good data, echo=FALSE, eval=data_qual}


selected_com_state %>% 
  ggplot2::ggplot(aes(x=Year, y=(total_dollars/10^6)))+
  tidyquant::geom_ma(aes(color= State, lty= State) ,n=3, lwd=1.4)+
  ggtitle(label = params$stock ,subtitle = "5 year rolling average")+
  xlab(label= "Year")+
  ylab(label = "Dollar value of landings in millions ($)  ")+
  theme_minimal()




```


```{r, landings across states for bad data, echo=FALSE, eval=!data_qual}


selected_com_state %>% 
  ggplot2::ggplot(aes(x=Year, y=(total_dollars/10^6)))+
  geom_point(aes(color=State))+
  geom_line(aes(color=State))+
  ggtitle(label = params$stock ,subtitle = "Data coverage for some states is poor with >5 years reported")+
  xlab(label= "Year")+
  ylab(label = "Dollar value of landings in millions ($)  ")+
  theme_minimal()

```




## Non-inflation adjusted price history of `r params$stock`
The calculated price is dollars per pound of reported landings. 



```{r Non-inflation adjusted price history , warning= FALSE, message= FALSE, error= FALSE, echo=FALSE }

selected_com_sp %>%
  dplyr::mutate(price_per_lb=(total_dollars/total_pounds))%>%
  ggplot2::ggplot(aes(x=Year, y=price_per_lb))+
  tidyquant::geom_ma(n=5, lwd=1.4)+
  ggtitle(label = params$stock ,subtitle = "5 year rolling average ")+
  xlab(label= "Year")+
  ylab(label = "Price per pound (lb/$) ")+
  theme_minimal()




```

## Non-inflation adjusted price history of `r params$stock` across states in the mid-Atlantic and north-east regions
The calculated price is dollars per pound of reported landings. 


```{r Non-inflation adjusted price history across states for good data , warning= FALSE, message= FALSE, error= FALSE, echo=FALSE, eval=data_qual}

selected_com_state %>%
  dplyr::mutate(price_per_lb=(total_dollars/total_pounds))%>%
  ggplot2::ggplot(aes(x=Year, y=price_per_lb))+
  tidyquant::geom_ma(aes(color= State, lty= State),n=5, lwd=1.4,show.legend = FALSE)+
  facet_wrap(vars(State))+
  ggtitle(label = params$stock ,subtitle = "5 year rolling average ")+
  xlab(label= "Year")+
  ylab(label = "Price per pound (lb/$) ")+
  theme_minimal()





```

```{r Non-inflation adjusted price history across states for poor data , warning= FALSE, message= FALSE, error= FALSE, echo=FALSE, eval=!data_qual}


selected_com_state %>%
  dplyr::mutate(price_per_lb=(total_dollars/total_pounds))%>%
  ggplot2::ggplot(aes(x=Year, y=price_per_lb))+
  geom_point(aes(color=State),show.legend = FALSE)+
  geom_line(aes(color=State),show.legend = FALSE)+
  facet_wrap(vars(State))+
  ggtitle(label = params$stock ,subtitle = "Data coverage for some states is poor with >5 years reported")+
  xlab(label= "Year")+
  ylab(label = "Price per pound (lb/$) ")+
  theme_minimal()


```





## inflation adjusted price history of `r params$stock`
The price history of `r params$stock` adjusted to 2019 values using data from the World Bank inflation data as informed by Inflation adjustment calculations in Principles of Macroeconomics by Gregory Mankiw et al (2014). Adjustements perfomed using the packages [priceR](https://github.com/stevecondylios/priceR).

```{r, inflation adjusted price history , warning= FALSE, message= FALSE, error= FALSE, echo=FALSE}

library(priceR)

stock_price<-selected_com_state %>%
  group_by(Year) %>% 
  dplyr::mutate(price_per_lb=(total_dollars/total_pounds))
country <- "US"
stock_price$adj_price<-adjust_for_inflation(stock_price$price_per_lb, stock_price$Year,country, to_date = 2019)



stock_price%>%
  group_by(Year)%>%
  summarise(adj_ave_price=mean(adj_price)) %>%
  ggplot2::ggplot(aes(x=Year, y=adj_ave_price))+
  tidyquant::geom_ma(n=5, lwd=1.4)+
  ggtitle(label = params$stock ,subtitle = "5 year rolling average ")+
  xlab(label= "Year")+
  ylab(label = "Price per pound (lb/$) in 2019 value ")+
  theme_minimal()

```


## inflation adjusted price history of `r params$stock` across states in the mid-Atlantic and north-east regions 



```{r, inflation adjusted price across states for good data, warning= FALSE, message= FALSE, error= FALSE, echo=FALSE, eval=data_qual}

stock_price%>%
  ggplot2::ggplot(aes(x=Year, y=adj_price))+
  tidyquant::geom_ma(aes( color=State),n=5, lwd=1.4,show.legend = FALSE)+
  facet_wrap(vars(State))+
  ggtitle(label = params$stock ,subtitle = "5 year rolling average ")+
  xlab(label= "Year")+
  ylab(label = "Price per pound (lb/$) in 2019 value ")+
  theme_minimal()
```





```{r, inflation adjusted price across states for poor data, warning= FALSE, message= FALSE, error= FALSE, echo=FALSE, eval=!data_qual}

stock_price %>%
  ggplot2::ggplot(aes(x=Year, y=adj_price))+
  geom_point(aes(color=State),show.legend = FALSE)+
  geom_line(aes(color=State),show.legend = FALSE)+
  facet_wrap(vars(State))+
  ggtitle(label = params$stock ,subtitle = "Data coverage for some states is poor with >5 years reported ")+
  xlab(label= "Year")+
  ylab(label = "Price per pound (lb/$) in 2019 value")+
  theme_minimal()
  

```

