---
title: "`r stringr::str_to_sentence(params$species)` indicator analysis"
author: "Abigail Tyrell"
date: "`r format(Sys.time(), '%d %b %Y')`"
output: word_document
params: 
  species: "black sea bass" # species name (for titles, etc.)
  fig_size: 6.5 # figure size (inches. square shape)
---

```{r, setup, include = FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.height = params$fig_size,
  fig.width = params$fig_size,
  dpi = 200 # not sure why default 72 dpi looks worse in word output compared to saved figs
)

`%>%` <- magrittr::`%>%`

source(here::here("R-scripts", "package-dev", "tailored_rpt_functions.R"))
```

## Recruitment 

### Winter SST anomaly
```{r, winter-sst}
wrap_analysis(file_path = here::here(
  "black-sea-bass", "Black sea bass_regression_report_0lag", "data",
  "Black_sea_bass_Mid_MAB_0_lag_FALSE_remove_recent_sst-anom-stock.csv"
  ),
  pattern = c("north", "south", "winter"),
  remove = c(TRUE, TRUE, FALSE),
  species = params$species
)
```

### Fall SST anomaly (prior year)
```{r, fall-sst}
wrap_analysis(file_path = here::here(
  "black-sea-bass", "Black sea bass_regression_report_1lag", "data",
  "Black_sea_bass_Mid_MAB_1_lag_FALSE_remove_recent_sst-anom-stock.csv"
  ),
  pattern = c("north", "south", "fall"),
  remove = c(TRUE, TRUE, FALSE),
  lag = 1,
  species = params$species
)
```

### Winter primary production
```{r, winter-pp}
wrap_analysis(file_path = here::here(
  "black-sea-bass", "Black sea bass_regression_report_0lag", "data",
  "Black_sea_bass_Mid_MAB_0_lag_FALSE_remove_recent_pp-stock.csv"
  ),
  pattern = c("north", "south", "winter"),
  remove = c(TRUE, TRUE, FALSE),
  species = params$species
)
```

### Fall primary production (prior year)
```{r, fall-pp}
wrap_analysis(file_path = here::here(
  "black-sea-bass", "Black sea bass_regression_report_1lag", "data",
  "Black_sea_bass_Mid_MAB_1_lag_FALSE_remove_recent_pp-stock.csv"
  ),
  pattern = c("north", "south", "fall"),
  remove = c(TRUE, TRUE, FALSE),
  lag = 1,
  species = params$species
)
```

### Cold pool index (prior year)
```{r, cold-pool}
wrap_analysis(file_path = here::here(
  "black-sea-bass", "Black sea bass_regression_report_1lag", "data",
  "Black_sea_bass_Mid_MAB_1_lag_FALSE_remove_recent_cold-pool.csv"
  ),
  lag = 1,
  species = params$species
)
```

### Predator abundance


## Ecosystem

### Sea surface temperature

#### Recruitment
```{r, sst-rec}
wrap_analysis(file_path = here::here(
  "black-sea-bass", "Black sea bass_regression_report_0lag", "data",
  "Black_sea_bass_Mid_MAB_0_lag_FALSE_remove_recent_sst.csv"
  ),
  species = params$species
)
```

#### Abundance
```{r, sst-abun}
wrap_analysis(file_path = here::here(
  "black-sea-bass", "Black sea bass_regression_report_0lag", "data",
  "Black_sea_bass_Mid_MAB_0_lag_FALSE_remove_recent_sst.csv"
  ),
  metric = "Abundance",
  species = params$species
)
```

### Marine heatwaves

#### Recruitment
```{r, heat-rec}
wrap_analysis(file_path = here::here(
  "black-sea-bass", "Black sea bass_regression_report_0lag", "data",
  "Black_sea_bass_Mid_MAB_0_lag_FALSE_remove_recent_heatwave.csv"
  ),
  species = params$species
)
```

#### Abundance
```{r, heat-abun}
wrap_analysis(file_path = here::here(
  "black-sea-bass", "Black sea bass_regression_report_0lag", "data",
  "Black_sea_bass_Mid_MAB_0_lag_FALSE_remove_recent_heatwave.csv"
  ),
  metric = "Abundance",
  species = params$species
)
```

## Management

### Total catch
```{r, catch, fig.height = 4.25, fig.width = 3.75}
data <- here::here(
  "black-sea-bass", "Black sea bass_regression_report_0lag", "data",
  "Black_sea_bass_Mid_MAB_0_lag_FALSE_remove_recent_time.csv"
) %>%
  read.csv() %>%
  dplyr::filter(Metric == "Catch") # pull catch only

my_colors <- c("black", "#B2292E")
names(my_colors) <- c("FALSE", "TRUE")

data$sig <- factor(data$sig, levels = c("TRUE", "FALSE"))

fig <- ggplot2::ggplot(data, ggplot2::aes(
  x = Val,
  y = Value
)) +
  ggplot2::geom_path(ggplot2::aes(color = Time)) +
  ggplot2::geom_point(ggplot2::aes(color = Time)) +
  ggplot2::scale_color_gradient(
    low = "#7F7FFF",
    high = "#575195",
    breaks = scales::breaks_extended(n = 4),
    name = "Year",
    guide = ggplot2::guide_colorbar(order = 2)
  ) +
  ggnewscale::new_scale_color() +
  ggplot2::stat_smooth(ggplot2::aes(color = sig),
    method = "lm"
  ) +
  ggplot2::scale_color_manual(
    values = my_colors,
    name = "Statistically significant\n(p < 0.05)",
    drop = FALSE,
    guide = ggplot2::guide_legend(order = 1)
  ) +
  ggplot2::scale_y_continuous(labels = scales::comma) +
  ggplot2::theme_bw() +
  ggplot2::labs(title = "Catch over time") +
  ggplot2::ylab(paste(unique(data$Description), unique(data$Units), sep = ", ")) +
  ggplot2::xlab("Year") +
  ggplot2::theme(
    legend.position = "top",
    legend.box = "vertical",
    legend.spacing = ggplot2::unit(0.1, units = "cm"),
    legend.box.spacing = ggplot2::unit(0.1, units = "cm")
  )

ggplot2::ggsave(
  filename = paste("figures/",
    unique(data$Var) %>%
      stringr::str_replace_all(" ", "_") %>%
      stringr::str_replace_all("\n", "_"),
    ".png",
    sep = ""
  ),
  width = 3.75,
  height = 4.25,
  units = "in",
  device = "png"
)

return(fig)
```

### CPUE

### Catch vs TAC

### Stock range

#### Northern range

Northernmost survey observation in each year.


```{r, north}
wrap_analysis(file_path = here::here(
  "black-sea-bass", "Black sea bass_regression_report_0lag", "data",
  "Black_sea_bass_Mid_MAB_0_lag_FALSE_remove_recent_northern-bound.csv"
  ),
  metric = "Catch",
  species = params$species
)
```

#### Southern range

Southernmost survey observation in each year.


```{r, south}
wrap_analysis(file_path = here::here(
  "black-sea-bass", "Black sea bass_regression_report_0lag", "data",
  "Black_sea_bass_Mid_MAB_0_lag_FALSE_remove_recent_southern-bound.csv"
  ),
  metric = "Catch",
  species = params$species
)
```

### Center of mass

## Report card
```{r, rpt-card}
knitr::kable(rpt_card_time)
```
