---
title: "Black sea bass indicator analysis"
author: "Abigail Tyrell"
date: "`r format(Sys.time(), '%d %b %Y')`"
output: word_document
---

```{r, setup, include = FALSE}
# set width & height
n <- 6.5

knitr::opts_chunk$set(
  echo = FALSE,
  warning = FALSE,
  message = FALSE,
  fig.height = n,
  fig.width = n,
  dpi = 200 # not sure why default 72 dpi looks worse in word output compared to saved figs
)

`%>%` <- magrittr::`%>%`
dir.create("figures")

# report cards to fill
rpt_card_ind <- c() # all indicators (not used yet/ever)
rpt_card_time <- c() # only statistically sig indicators

# functions
plot_corr_only <- function(data, title = "",
                           tab_data = NULL, lag = 0) { # option to correct for lagged data
  if (nrow(data) > 0) {
    my_colors <- c("black", "#B2292E")
    names(my_colors) <- c("FALSE", "TRUE")

    data$sig <- factor(data$sig, levels = c("TRUE", "FALSE"))

    fig <- ggplot2::ggplot(data, ggplot2::aes(
      x = Val,
      y = Value
    )) +
      ggplot2::geom_path(ggplot2::aes(color = Time)) +
      ggplot2::geom_point(ggplot2::aes(color = Time)) +
      ggplot2::scale_color_gradient(
        low = "#7F7FFF",
        high = "#575195",
        breaks = scales::breaks_extended(n = 4),
        name = "Year",
        guide = ggplot2::guide_colorbar(order = 2)
      ) +
      ggnewscale::new_scale_color() +
      ggplot2::stat_smooth(ggplot2::aes(color = sig),
        method = "lm"
      ) +
      ggplot2::scale_color_manual(
        values = my_colors,
        name = "Statistically significant\n(p < 0.05)",
        drop = FALSE,
        guide = ggplot2::guide_legend(order = 1)
      ) +
      ggplot2::scale_y_continuous(labels = scales::comma) +
      ggplot2::scale_x_continuous(labels = scales::comma) +
      ggplot2::theme_bw() +
      ggplot2::labs(title = "Correlation between black sea bass and indicator") +
      ggplot2::ylab(unique(data$Metric[!is.na(data$Metric)])) +
      ggplot2::xlab(unique(data$Var))

    # test if bsb is sig over time - overwrite sig
    dat <- data %>%
      dplyr::select(Value, Time) %>%
      dplyr::distinct() %>%
      dplyr::mutate(Time = Time + lag)
    model <- lm(Value ~ Time, data = dat)
    pval <- summary(model)$coefficients[2, 4]
    data$sig <- (pval < 0.05)

    bsb_fig <- ggplot2::ggplot(
      data,
      ggplot2::aes(
        x = Time + lag,
        y = Value
      )
    ) +
      ggplot2::geom_path(ggplot2::aes(color = Time)) +
      ggplot2::geom_point(ggplot2::aes(color = Time)) +
      ggplot2::scale_color_gradient(
        low = "#7F7FFF",
        high = "#575195",
        breaks = scales::breaks_extended(n = 4),
        name = "Year"
      ) +
      ggnewscale::new_scale_color() +
      ggplot2::stat_smooth(ggplot2::aes(color = sig),
        method = "lm"
      ) +
      ggplot2::scale_color_manual(
        values = my_colors,
        name = "Statistically significant\n(p < 0.05)",
        drop = FALSE
      ) +
      ggplot2::scale_y_continuous(labels = scales::comma) +
      ggplot2::theme_bw() +
      ggplot2::labs(title = "Black sea bass") +
      ggplot2::xlab("Year") +
      ggplot2::ylab(unique(data$Metric[!is.na(data$Metric)]))

    # test if indicator is sig over time - overwrite sig
    dat <- data %>%
      dplyr::select(Val, Time) %>%
      dplyr::distinct()
    model <- lm(Val ~ Time, data = dat)
    pval <- summary(model)$coefficients[2, 4]
    data$sig <- (pval < 0.05)

    # reformat Var for y-label
    data$Var <- data$Var %>%
      stringr::str_replace("\n", " ") %>%
      stringr::str_wrap(width = 30)

    ind_fig <- ggplot2::ggplot(
      data,
      ggplot2::aes(
        x = Time,
        y = Val
      )
    ) +
      ggplot2::geom_path(ggplot2::aes(color = Time)) +
      ggplot2::geom_point(ggplot2::aes(color = Time)) +
      ggplot2::scale_color_gradient(
        low = "#7F7FFF",
        high = "#575195",
        breaks = scales::breaks_extended(n = 4),
        name = "Year"
      ) +
      ggnewscale::new_scale_color() +
      ggplot2::stat_smooth(ggplot2::aes(color = sig),
        method = "lm"
      ) +
      ggplot2::scale_color_manual(
        values = my_colors,
        name = "Statistically significant\n(p < 0.05)",
        drop = FALSE
      ) +
      ggplot2::scale_y_continuous(labels = scales::comma) +
      ggplot2::theme_bw() +
      ggplot2::labs(title = "Indicator") +
      ggplot2::xlab("Year") +
      ggplot2::ylab(unique(data$Var))

    # version with summary table
    # tbl <- ggpubr::ggtexttable(tab_data, theme = ggpubr::ttheme("blank"), rows = NULL)  %>%
    #  ggpubr::tbody_add_border() %>%
    #  ggpubr::thead_add_border()

    # big_fig <- ggpubr::ggarrange(
    #  ggpubr::ggarrange(bsb_fig, ind_fig, tbl,
    #    ncol = 3,
    #    legend = "none",
    #    labels = c("A", "B", "C")
    #  ),
    #  fig,
    #  labels = c(NA, "D"),
    #  common.legend = TRUE,
    #  legend = "top",
    #  nrow = 2
    # )

    big_fig <- ggpubr::ggarrange(
      ggpubr::ggarrange(bsb_fig, ind_fig,
        ncol = 2,
        legend = "none",
        labels = c("A", "B")
      ),
      fig,
      labels = c(NA, "C"),
      common.legend = TRUE,
      legend = "top",
      nrow = 2
    )

    big_fig <- big_fig +
      ggplot2::theme(
        plot.background = ggplot2::element_rect(color = "black", size = 1),
        plot.margin = ggplot2::unit(c(0.1, 0.1, 0.1, 0.1), "cm")
      )

    ggpubr::annotate_figure(big_fig, top = title)

    # save
    file <- paste("figures/",
      unique(data$Metric)[!is.na(unique(data$Metric))],
      "_",
      unique(data$Var),
      ".png",
      sep = ""
    ) %>%
      stringr::str_replace_all(" ", "_") %>%
      stringr::str_replace_all("\n", "_")

    ggplot2::ggsave(
      filename = file,
      width = 6.5,
      height = 6.5,
      units = "in",
      device = "png"
    )

    return(big_fig)
  }
  else {
    print("No data under conditions selected")
  }
}

prep_data <- function(data, metric = "Recruitment") {
  data <- data %>%
    read.csv() %>%
    dplyr::filter(
      Metric == metric | is.na(Metric),
      stringr::str_detect(Var, "north", negate = TRUE),
      stringr::str_detect(Var, "south", negate = TRUE)
    ) %>%
    dplyr::mutate(Var = Var %>%
      stringr::str_replace_all("\n", " ") %>%
      stringr::str_wrap(40)) %>%
    dplyr::arrange(Time)

  return(data)
}

# for northern/southern bounds
prep_data2 <- function(data, metric = "Recruitment") {
  data <- data %>%
    read.csv() %>%
    dplyr::filter(
      Metric == metric | is.na(Metric),
    ) %>%
    dplyr::mutate(Var = Var %>%
      stringr::str_replace_all("\n", " ") %>%
      stringr::str_wrap(40)) %>%
    dplyr::arrange(Time)

  return(data)
}

time_rpt <- function(data, col_name = "Val", out_name = "unnamed") {
  data <- data %>%
    dplyr::rename(Test = col_name) %>%
    dplyr::select(Time, Test) %>%
    dplyr::distinct()

  data <- data %>%
    dplyr::group_by(Time) %>%
    dplyr::mutate(avg_value = mean(Test, na.rm = TRUE)) %>% # average by year
    dplyr::select(-Test) %>%
    dplyr::distinct()

  analysis <- data %>%
    dplyr::ungroup() %>%
    dplyr::mutate(
      long_avg = mean(avg_value) %>%
        round(digits = 2),
      long_sd = sd(avg_value) %>%
        round(digits = 2)
    ) %>%
    dplyr::filter(Time >= 2016) %>%
    dplyr::mutate(
      yr5_avg = mean(avg_value) %>%
        round(digits = 2),
      yr5_sd = sd(avg_value) %>%
        round(digits = 2),
      avg_value = round(avg_value, digits = 2)
    )

  status <- c()
  for (i in 1:nrow(analysis)) {
    if (analysis$avg_value[i] > (analysis$long_avg[1] + analysis$long_sd[1])) {
      status[i] <- "high"
    } else if (analysis$avg_value[i] < (analysis$long_avg[1] - analysis$long_sd[1])) {
      status[i] <- "low"
    } else {
      status[i] <- "neutral"
    }
  }

  analysis$avg_value <- paste(analysis$avg_value,
    status,
    sep = ", "
  )

  output <- rbind(
    cbind(analysis$Time, analysis$avg_value),
    c("5-year mean", paste(analysis$yr5_avg, "±", analysis$yr5_sd)),
    c("long-term mean", paste(analysis$long_avg, "±", analysis$long_sd))
  ) %>%
    tibble::as_tibble()

  colnames(output) <- c("Time", out_name)

  return(output)
}
```

## Recruitment 

### Winter SST anomaly
```{r, winter-sst}
data <- here::here(
  "black-sea-bass", "Black sea bass_regression_report_0lag", "data",
  "Black_sea_bass_Mid_MAB_0_lag_FALSE_remove_recent_sst-anom-stock.csv"
) %>%
  prep_data() %>%
  dplyr::filter(stringr::str_detect(Var, "winter"))

data %>%
  plot_corr_only()

rpt_card_time <- time_rpt(data, out_name = "SST_winter")
```

### Fall SST anomaly (prior year)
```{r, fall-sst}
data <- here::here(
  "black-sea-bass", "Black sea bass_regression_report_1lag", "data",
  "Black_sea_bass_Mid_MAB_1_lag_FALSE_remove_recent_sst-anom-stock.csv"
) %>%
  prep_data() %>%
  dplyr::filter(stringr::str_detect(Var, "fall"))

data %>%
  plot_corr_only(lag = 1)

rpt_card_time <- dplyr::full_join(rpt_card_time,
  time_rpt(data, out_name = "SST_fall"),
  by = "Time"
)
```

### Winter primary production
```{r, winter-pp}
data <- here::here(
  "black-sea-bass", "Black sea bass_regression_report_0lag", "data",
  "Black_sea_bass_Mid_MAB_0_lag_FALSE_remove_recent_pp-stock.csv"
) %>%
  prep_data() %>%
  dplyr::filter(stringr::str_detect(Var, "winter"))

data %>%
  plot_corr_only()

rpt_card_time <- dplyr::full_join(rpt_card_time,
  time_rpt(data, out_name = "PP_winter"),
  by = "Time"
)
```

### Fall primary production (prior year)
```{r, fall-pp}
data <- here::here(
  "black-sea-bass", "Black sea bass_regression_report_1lag", "data",
  "Black_sea_bass_Mid_MAB_1_lag_FALSE_remove_recent_pp-stock.csv"
) %>%
  prep_data() %>%
  dplyr::filter(stringr::str_detect(Var, "fall"))

data %>%
  plot_corr_only(lag = 1)

rpt_card_time <- dplyr::full_join(rpt_card_time,
  time_rpt(data, out_name = "PP_fall"),
  by = "Time"
)
```

### Cold pool index (prior year)
```{r, cold-pool}
data <- here::here(
  "black-sea-bass", "Black sea bass_regression_report_1lag", "data",
  "Black_sea_bass_Mid_MAB_1_lag_FALSE_remove_recent_cold-pool.csv"
) %>%
  prep_data()

for (i in unique(data$Var)) {
  this_data <- data %>%
    dplyr::filter(Var == i)

  plt <- plot_corr_only(this_data, title = i, lag = 1)
  print(plt)
  cat("\n\n")

  rpt_card_time <- dplyr::full_join(rpt_card_time,
    time_rpt(this_data, out_name = i),
    by = "Time"
  )
}
```

### Predator abundance


## Ecosystem

### Sea surface temperature

#### Recruitment
```{r, sst-rec}
data <- here::here(
  "black-sea-bass", "Black sea bass_regression_report_0lag", "data",
  "Black_sea_bass_Mid_MAB_0_lag_FALSE_remove_recent_sst.csv"
) %>%
  prep_data()

data %>%
  plot_corr_only()

rpt_card_time <- dplyr::full_join(rpt_card_time,
  time_rpt(data, out_name = "SST_annual"),
  by = "Time"
)
```

#### Abundance
```{r, sst-abun}
here::here(
  "black-sea-bass", "Black sea bass_regression_report_0lag", "data",
  "Black_sea_bass_Mid_MAB_0_lag_FALSE_remove_recent_sst.csv"
) %>%
  prep_data(metric = "Abundance") %>%
  plot_corr_only()
```

### Marine heatwaves

#### Recruitment
```{r, heat-rec}
data <- here::here(
  "black-sea-bass", "Black sea bass_regression_report_0lag", "data",
  "Black_sea_bass_Mid_MAB_0_lag_FALSE_remove_recent_heatwave.csv"
) %>%
  prep_data()

for (i in unique(data$Var)) {
  this_data <- data %>%
    dplyr::filter(Var == i)

  plt <- plot_corr_only(this_data, title = i)
  print(plt)
  cat("\n\n")

  rpt_card_time <- dplyr::full_join(rpt_card_time,
    time_rpt(this_data, out_name = i),
    by = "Time"
  )
}
```

#### Abundance
```{r, heat-abun}
data <- here::here(
  "black-sea-bass", "Black sea bass_regression_report_0lag", "data",
  "Black_sea_bass_Mid_MAB_0_lag_FALSE_remove_recent_heatwave.csv"
) %>%
  prep_data(metric = "Abundance")

for (i in unique(data$Var)) {
  this_data <- data %>%
    dplyr::filter(Var == i)

  plt <- plot_corr_only(this_data, title = i)
  print(plt)
  cat("\n\n")
}
```

## Management

### Total catch
```{r, catch, fig.height = 4.25, fig.width = 3.75}
data <- here::here(
  "black-sea-bass", "Black sea bass_regression_report_0lag", "data",
  "Black_sea_bass_Mid_MAB_0_lag_FALSE_remove_recent_time.csv"
) %>%
  read.csv() %>%
  dplyr::filter(Metric == "Catch") # pull catch only

my_colors <- c("black", "#B2292E")
names(my_colors) <- c("FALSE", "TRUE")

data$sig <- factor(data$sig, levels = c("TRUE", "FALSE"))

fig <- ggplot2::ggplot(data, ggplot2::aes(
  x = Val,
  y = Value
)) +
  ggplot2::geom_path(ggplot2::aes(color = Time)) +
  ggplot2::geom_point(ggplot2::aes(color = Time)) +
  ggplot2::scale_color_gradient(
    low = "#7F7FFF",
    high = "#575195",
    breaks = scales::breaks_extended(n = 4),
    name = "Year",
    guide = ggplot2::guide_colorbar(order = 2)
  ) +
  ggnewscale::new_scale_color() +
  ggplot2::stat_smooth(ggplot2::aes(color = sig),
    method = "lm"
  ) +
  ggplot2::scale_color_manual(
    values = my_colors,
    name = "Statistically significant\n(p < 0.05)",
    drop = FALSE,
    guide = ggplot2::guide_legend(order = 1)
  ) +
  ggplot2::scale_y_continuous(labels = scales::comma) +
  ggplot2::theme_bw() +
  ggplot2::labs(title = "Catch over time") +
  ggplot2::ylab(paste(unique(data$Description), unique(data$Units), sep = ", ")) +
  ggplot2::xlab("Year") +
  ggplot2::theme(
    legend.position = "top",
    legend.box = "vertical",
    legend.spacing = ggplot2::unit(0.1, units = "cm"),
    legend.box.spacing = ggplot2::unit(0.1, units = "cm")
  )

ggplot2::ggsave(
  filename = paste("figures/",
    unique(data$Var) %>%
      stringr::str_replace_all(" ", "_") %>%
      stringr::str_replace_all("\n", "_"),
    ".png",
    sep = ""
  ),
  width = 3.75,
  height = 4.25,
  units = "in",
  device = "png"
)

return(fig)
```

### CPUE

### Catch vs TAC

### Stock range

#### Northern range

Northernmost survey observation in each year.


```{r, north}
data <- here::here(
  "black-sea-bass", "Black sea bass_regression_report_0lag", "data",
  "Black_sea_bass_Mid_MAB_0_lag_FALSE_remove_recent_northern-bound.csv"
) %>%
  prep_data2(metric = "Catch")

data %>%
  plot_corr_only()

rpt_card_time <- dplyr::full_join(rpt_card_time,
  time_rpt(data, out_name = "northern_range"),
  by = "Time"
)
```

#### Southern range

Southernmost survey observation in each year.


```{r, south}
data <- here::here(
  "black-sea-bass", "Black sea bass_regression_report_0lag", "data",
  "Black_sea_bass_Mid_MAB_0_lag_FALSE_remove_recent_southern-bound.csv"
) %>%
  prep_data2(metric = "Catch")

data %>%
  plot_corr_only()

rpt_card_time <- dplyr::full_join(rpt_card_time,
  time_rpt(data, out_name = "southern_range"),
  by = "Time"
)
```

### Center of mass

## Report card
```{r, rpt-card}
knitr::kable(rpt_card_time)
```
