---
title: "risk_plot"
author: "Abigail Tyrell"
date: "`r format(Sys.time(), '%d %b %Y')`"
output: 
  html_document:
    toc: FALSE
    toc_float: 
      collapsed: FALSE
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, 
                      message = FALSE,
                      warning = FALSE, 
                      fig.width = 8, 
                      fig.align = 'center')

library(ggplot2)
`%>%` <- dplyr::`%>%`

source(here::here("R", "character_to_factor.R"))
source(here::here("R/full_report_functions", "make_html_table.R"))

data <- read.csv(here::here("data/risk_ranking", "full_risk_data.csv"))
guild_data <- read.csv(here::here("data/risk_ranking", "guild_data.csv"))
guild_info <- read.csv(here::here("data/risk_ranking", "guild_info.csv"))
indicator_info <- read.csv(here::here("data/risk_ranking", "indicator_key.csv"))

data$category <- factor(data$category, 
                        c("Socioeconomic", "Biological", "Population", "Habitat", "Management"))

guild_data$category <- factor(guild_data$category, 
                        c("Socioeconomic", "Biological", "Population", "Habitat", "Management"))

# distinguish between 0.5 due to missing data and true 0.5 ranking
for(i in 1:nrow(data)){
  if(data$Value[i] %>% is.na()){
    data$norm_rank[i] <- NA
  }
}
```

# Risk ranking of Northeast stocks {.tabset .tabset-fade}

## Methods {.tabset .tabset-fade}

### Risk ranking

All stocks were ranked in order of increasing risk. The stock with the highest ranking is the stock determined to be at the highest risk. In this case, high risk has two meanings: (1) high importance (e.g., a stock with a high catch would have a high risk ranking for the catch indicator) or high vulnerability (e.g., a stock with low B/Bmsy would have a high risk ranking for the B/Bmsy indicator). The normalized rank was determined by dividing each stock's rank by the total number of stocks considered for that indicator. Stocks that were missing indicator measurements were assigned a normalized rank of 0.5.

```{r, indicators}
dat <- indicator_info %>%
  dplyr::arrange(Category)
make_html_table_thin(dat, col_names = colnames(dat))
```

### Guilds

Guilds were determined from [previous reporting](https://repository.library.noaa.gov/view/noaa/5277). Guilds were further broken into small (<40cm) and large (>=40cm) size classes. To determine guild risk, the risk values for each species were averaged over all stock areas. Then risk values were averaged by guild. Guilds were ranked in each indicator category.

```{r, guild_info, results = "asis"}
guild_info <- guild_info %>% 
  dplyr::mutate(Scientific_name = paste(Scientific_name, sep = "")) %>% # figure out how to italicize in html table
  dplyr::rename("Size" = "size") %>%
  dplyr::select(Guild, Size, Species, Scientific_name) %>%
  dplyr::arrange(Guild, Size) 

make_html_table(guild_info, col_names = colnames(guild_info))
```

## Results {.tabset .tabset-fade}

### Stock risk {.tabset .tabset-fade .tabset-pills}
Ranks for all stocks. 

#### Figures {.tabset .tabset-fade}

The number following the species name is the total risk (sum of all normalized risk values).

```{r, regions}
dat <- data.frame(Abbreviation = c("CC", "GOM", "GB", "NE", "Mid", 
                                   "E", "N", "S"),
                  Region = c("Cape Cod", "Gulf of Maine", "Georges Bank", 
                             "New England", "Mid-Atlantic", "East", "North", 
                             "South"))

dat %>% DT::datatable(rownames = FALSE,
                      extensions = 'Scroller',
                      caption = "Region abbreviations",
                      options = list(deferRender = TRUE,
                                     scrollY = 100,
                                     scrollX = TRUE,
                                     scroller = TRUE))
```

##### By region
```{r plot_region, fig.height = 15}
data$Region2 <- data$Region %>% 
    stringr::str_replace("Georges Bank", "GB") %>%
    stringr::str_replace("Gulf of Maine", "GOM") %>%
    stringr::str_replace("Northwestern", "NW") %>%
    stringr::str_replace("Northern", "N") %>%
    stringr::str_replace("Southern", "S") %>%
    stringr::str_replace("Eastern", "E") %>%
    stringr::str_replace("Cape Cod", "CC") %>%
    stringr::str_replace("New England", "NE") %>%
    stringr::str_replace(" / ", "/\n") %>%
    stringr::str_replace("Mid-Atlantic", "Mid") %>%
    stringr::str_replace("Cape Hatteras", "Cape\nHatteras")

fig <- ggplot(data,
                   aes(y = reorder(label, -total_risk),
                       x = Indicator %>% 
                              stringr::str_replace_all("_", " "),
                       fill = norm_rank ))+
  geom_raster(stat = "identity")+
  theme_bw()+
  viridis::scale_fill_viridis(limits = c(0, 1),
                              breaks = c(0, 0.5, 1),
                              direction = -1,
                              na.value = "gray90",
                              name = "Normalized rank")+
  theme(axis.text.x = element_text(angle = 90),
        
        panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        strip.text.y = element_text(angle = 0),
        strip.text.x = element_text(angle = 75),
        legend.position = "top")+
  facet_grid(rows = vars(Region2),
             cols = vars(category),
             scales = "free",
             space = "free")+
  ylab("") +
  xlab("Indicator")

print(fig)

```

##### In order of risk
```{r, all_stocks, fig.height = 15}
new_label <- paste(data$Region, " ", data$Species, ", ", 
                              data$total_risk %>% round(digits = 2), 
                              sep = "")

# shorten labels
data$new_label <- new_label %>%
  stringr::str_replace("Georges Bank", "GB") %>%
  stringr::str_replace("Gulf of Maine", "GOM") %>%
  stringr::str_replace("Northwestern", "NW") %>%
  stringr::str_replace("Northern", "N") %>%
  stringr::str_replace("Southern", "S") %>%
  stringr::str_replace("Cape Cod", "CC") %>%
  stringr::str_replace("New England", "NE") %>%
  stringr::str_replace("Cape Hatteras", "C Hatteras") %>%
  stringr::str_replace("Atlantic mackerel", "Atl mackerel") %>%
  stringr::str_replace(" / ", "/")

fig <- ggplot(data,
              aes(y = reorder(new_label, -total_risk),
                  x = Indicator %>% 
                              stringr::str_replace_all("_", " "),
                  fill = norm_rank ))+
  geom_raster(stat = "identity")+
  facet_grid(cols = vars(category),
             scales = "free_x",
             space = "free_x")+
  theme_bw()+
  viridis::scale_fill_viridis(limits = c(0, 1),
                              breaks = c(0, 0.5, 1),
                              direction = -1,
                              na.value = "gray90",
                              name = "Normalized rank")+
  theme(axis.title.y = element_blank(),
        axis.text.x = element_text(angle = 90),
        strip.text.x = element_text(angle = 75),
        legend.position = "top")+
  xlab("Indicator")

print(fig)
```

#### Data  {.tabset .tabset-fade}

##### All data
```{r, data}
dat <- data %>% 
  dplyr::mutate(Value = Value %>%
                  round(digits = 2) %>%
                  format(scientific = FALSE,
                         big.mark = ","),
                norm_rank = norm_rank %>%
                  round(digits = 3),
                total_risk = total_risk %>%
                  round(digits = 2)) %>%
  dplyr::select(-X, -stock, -label, -Region2, -new_label, -n_stocks_per_region) %>%
  character_to_factor()

make_html_table(dat, col_names = colnames(dat))
```

##### Total risk only
```{r, data_totrisk}
dat <- data %>%
  dplyr::select(Species, Region, total_risk, overall_rank) %>%
  dplyr::mutate(total_risk = total_risk %>%
                  round(digits = 2)) %>%
  dplyr::distinct() %>%
  character_to_factor()

make_html_table(dat, col_names = colnames(dat))
```

### Guild risk{.tabset .tabset-fade .tabset-pills}

#### Figure

The number following the guild name is the total risk (sum of all average normalized risk values).

```{r guild, fig.height = 8}
ggplot(guild_data,
       aes(y = reorder(paste(Guild, "\n", 
                             size, "\n", 
                             total_guild_risk %>% round(digits = 2)) %>%
                         stringr::str_replace("_", "\n"), 
                       total_guild_risk),
           x = Indicator %>% 
             stringr::str_replace_all("_", " "), 
           fill = avg_guild_risk)) +
  geom_raster(stat = "identity")+
  facet_grid(cols = vars(category),
             scales = "free",
             space = "free")+
  viridis::scale_fill_viridis(limits = c(0, 1),
                              breaks = c(0, 0.5, 1),
                              direction = -1,
                              na.value = "gray90",
                              name = "Average normalized rank")+
  theme_bw()+
  ylab("Guild")+
  xlab("")+
  theme(legend.position = "top",
        strip.text.x = element_text(angle = 75),
        axis.text.x = element_text(angle = 90)
        )
```

#### Data
```{r, guild_data}
dat <- guild_data %>% 
  dplyr::mutate(Guild = paste(Guild, size)) %>%
  dplyr::select(Guild, n_guild_species, category, Indicator, avg_guild_risk, 
                total_guild_risk) %>%
  dplyr::mutate(avg_guild_risk = avg_guild_risk %>%
                  round(digits = 2) %>%
                  format(scientific = FALSE,
                         big.mark = ","),
                total_guild_risk = total_guild_risk %>%
                  round(digits = 2)) %>%
  character_to_factor()

make_html_table(dat, col_names = colnames(dat))
```