## Physical indicators

### Cold pool index
```{r, cold_pool}
test <- ecodata::cold_pool
render_indicator(test)
```

### Warm core rings
```{r, warm}
test <- ecodata::wcr %>%
  dplyr::rename(Val = Value) %>%
  dplyr::mutate(Var = paste(Var, Units, sep = "\n"))
render_indicator(test)
```

### Marine heatwave index
```{r, heatwave}
test <- ecodata::heatwave %>%
  dplyr::rename(Val = Value) %>%
  dplyr::mutate(Var = paste(Var, Units, sep = "\n"))
render_indicator(test)
```

### GLORYS bottom temperature
```{r, glorys}
test <- ecodata::bottom_temp_glorys %>%
  dplyr::rename(Val = Value) %>%
  dplyr::filter(EPU == "MAB") %>%
  dplyr::mutate(Var = paste(Var, Units, sep = "\n"))
render_indicator(test)
```

### Long-term sea surface temperature
```{r, sst}
test <- ecodata::long_term_sst  %>%
  dplyr::rename(Val = Value) %>%
  dplyr::mutate(Var = paste(Var, Units, sep = "\n"))
render_indicator(test)
```

### Sea surface temperature anomaly
```{r, sst_anom}
test <- ecodata::seasonal_oisst_anom %>%
  dplyr::rename(Val = Value) %>%
  dplyr::filter(EPU == "MAB") %>%
  dplyr::mutate(Var = paste(Var, Units, sep = "\n"))
render_indicator(test)
```

### Spring offshore bottom temperature
Bluefish are typically found offshore in the spring.

### Fall onshore bottom temperature
Bluefish are typically found onshore in the fall.

### Stratification
```{r, strat}
test <- ecodata::stratification %>%
  dplyr::rename(Val = Value) %>%
  dplyr::filter(EPU == "MAB") %>%
  dplyr::mutate(Var = paste(Var, Units, sep = "\n"))
render_indicator(test)
```

### Winter wind speed
```{r, winter_wind}
test <- ecodata::ne_wind %>%
  dplyr::rename(Val = Value) %>%
  dplyr::filter(EPU == "MAB",
                Var %>% stringr::str_detect("winter")) %>%
  dplyr::mutate(Var = paste(Var, Units, sep = "\n"))
render_indicator(test)
```

### Spring wind speed
```{r, spring_wind}
test <- ecodata::ne_wind %>%
  dplyr::rename(Val = Value) %>%
  dplyr::filter(EPU == "MAB",
                Var %>% stringr::str_detect("spring")) %>%
  dplyr::mutate(Var = paste(Var, Units, sep = "\n"))
render_indicator(test)
```

### Summer wind speed
```{r, summer_wind}
test <- ecodata::ne_wind %>%
  dplyr::rename(Val = Value) %>%
  dplyr::filter(EPU == "MAB",
                Var %>% stringr::str_detect("summer")) %>%
  dplyr::mutate(Var = paste(Var, Units, sep = "\n"))
render_indicator(test)
```

### Fall wind speed
```{r, fall_wind}
test <- ecodata::ne_wind %>%
  dplyr::rename(Val = Value) %>%
  dplyr::filter(EPU == "MAB",
                Var %>% stringr::str_detect("fall")) %>%
  dplyr::mutate(Var = paste(Var, Units, sep = "\n"))
render_indicator(test)
```

### Gulf Stream Index
```{r, gsi}
test <- ecodata::gsi %>%
  dplyr::rename(Val = Value) %>%
  dplyr::mutate(Var = paste(Var, Units, sep = "\n"),
                Time = Time %>%
                  stringr::str_trunc(width = 4, ellipsis = "") %>%
                  as.numeric())
render_indicator(test)
```

### North Atlantic Oscillation
```{r, nao}
test <- ecodata::nao %>%
  dplyr::rename(Val = Value) %>%
  dplyr::mutate(Var = paste(Var, Units, sep = "\n"))
render_indicator(test)
```

### Chlorophyll
```{r, chl}
test <- ecodata::chl_pp %>%
  dplyr::rename(Val = Value) %>%
  dplyr::filter(EPU == "MAB") %>%
  dplyr::mutate(Var = Var %>%
                  stringr::str_replace_all("_", "\n") %>%
                  stringr::str_replace_all(" ", "\n")) %>%
  dplyr::mutate(Var = paste(Var, Units, sep = "\n"))
render_indicator(test)
```
