## von Bertalanffy growth curve

Ricky Tabandera

```{r setup_vonb, cache = params$cache}
#install.packages("ggrepel")
#install.packages("ggpubr")
#devtools::install_github("james-thorson/FishLife")

suppressPackageStartupMessages(library(FishLife))
suppressPackageStartupMessages(library(tidyverse))
library(rfishbase)
library(ggpubr)
```



```{r data_wrangling, cache = params$cache, cache.lazy = FALSE}

#function to calculate the fitted growth function for von bertalanffy growth curve
vonb<-function(t,Linf,K,tzero){Latthisage<-Linf*(1-exp(-K*(t-tzero)))
  return(Latthisage)
}

#Pauly 1979 based method to estimate a default value for t0 as log (-t0) = -0.3922 - 0.2752 log Linf - 1.038 log K
#consistant way of generating t0  
tzero<-function(Linf,K){logtzero<-(-0.3922 - (0.2752*log(Linf) - 1.038*log(K)))
  t0<-exp(logtzero)
  return(t0)  
}  

stock_list_all_strata<- readr::read_csv("https://raw.githubusercontent.com/NOAA-EDAB/ECSA/master/data/stock_list.csv")
stock_list<-stock_list_all_strata %>% dplyr::distinct( common_name, .keep_all=TRUE)

fishlife_predictions_NE<-readr::read_csv(here::here("life history/fishlife","FishLife_extracted_data_11.16.20_northeast.csv"))
predictions_NE_spaced<-fishlife_predictions_NE %>% dplyr::mutate_all(funs(str_replace(., "\\_", " ")))
predictions_NE_spaced<-predictions_NE_spaced %>% dplyr::rename(sci_name = Species)
predictions_join <-dplyr::inner_join(stock_list, predictions_NE_spaced, by="sci_name" )

predictions_join<-predictions_join %>% 
  dplyr::filter(WORKED==TRUE)
#reformat these into numeric cols
predictions_join<-predictions_join %>% 
  dplyr::mutate(Loo=as.numeric(Loo), K=as.numeric(K))

#return the model output into original scale
predictions_join$Linf<-exp(predictions_join$Loo)
predictions_join$K<-exp(predictions_join$K)

predictions_join<-predictions_join %>% mutate(t0=tzero(Linf, K))
#write.csv(predictions_join, file = "predictions_join.csv")

#select out needed cols
prediction.select<-predictions_join %>% select(common_name, Linf,K,t0)

#catagorize the Linf into rough bins
prediction.select$body_cat<-cut(prediction.select$Linf, breaks=c(-Inf,50, 100, Inf), 
                                labels=c("small","med","large"))

#split into a set of lists for map use
flist<-split(prediction.select, f=prediction.select$common_name)

# vonb wants arguments in the order t,Linf,K,tzero)

spec.age.growth.list<-map(flist, ~vonb(seq(1:75),.x[[2]],.x[[3]],.x[[4]]))

# turn the list of lists into a single object
spec.age.growth<-do.call(cbind,spec.age.growth.list)
#turn this object into a df for manipulation
spec.age.growth.df<-data.frame(spec.age.growth)
# make an index of years int he dataset 

spec.age.growth.df$year<-seq(1:75)
#extract col names for pivoting 
fish.names<-colnames(spec.age.growth.df)
fish.names<-fish.names[fish.names != "year"]

#Pivot from wide format to long to make graphing easy

age.growth.curves<-spec.age.growth.df %>% pivot_longer(cols = fish.names,names_to = "common_name", values_to = "length" )
#remove the period in common name to enable the matching fuction in dplyr join
age.growth.curves$common_name<-str_replace(age.growth.curves$common_name,"\\."," " )

age.growth.join<-right_join(prediction.select,age.growth.curves, by= "common_name")

age.growth.join$common_name<-  str_to_lower(age.growth.join$common_name)
#recode black sea.bass to black sea bass inorder to be consistant with other datasets 

age.growth.join$common_name<- recode(age.growth.join$common_name, "black sea.bass" = "black sea bass")
```

```{r, vonb_species}
ricky_stock <- stringr::str_to_lower(species)
#########################################################
#logical filtering of params for plotting

is_small<-FALSE
is_med<-FALSE
is_large<-FALSE

sizeclass<-age.growth.join %>% filter(common_name==ricky_stock)

if(nrow(sizeclass) > 0 &
   species != "Black sea bass"){ # formatting problem with Black sea bass, omit for now
  
  if(sizeclass$body_cat=="small"){
  is_small<-TRUE}

  if(sizeclass$body_cat=="med"){
    is_med<-TRUE}
  
  if(sizeclass$body_cat=="large"){
    is_large<-TRUE}
} else missing <- TRUE
 
```

### Length at age growth curve

The predicted von Bertalanffy growth curve for NMFS managed fish species. Growth parameters of Linf (Length infinty), K (growth coefficient)
were estimated using [James Thorson FishLife R package](https://github.com/James-Thorson-NOAA/FishLife). This package uses a mix of previously published life history parameters extracted from fishbase and correlated higher taxon parameters to generate estimates for growth and population parameters. t0 (size at age 0) was estimated using the method described by Pauly 1979 for generating a consistent value for t0 from Linf and K. 

```{r single_growth_curve}
if(nrow(sizeclass) > 0){
  sizeclass %>%
  ggplot(aes(x=year, y=length))+
  geom_line(color= "blue",size=1.4)+
  xlab("Year")+
  ylab(" Total length (cm)")+
  labs(color="Common name of stock")+
  ggtitle(str_to_sentence(ricky_stock), subtitle = "Length at age")+
  theme_minimal()
} else print("NO DATA")

```

### Growth by size class

Fish were categorized into three groups with small bodied species reaching max size < 50 cm, medium 50-100 cm, and large >100 cm as rough divisions for comparisons of growth among similar body sizes.

```{r small_growth_curves, eval=is_small, cache = params$cache}

if(nrow(sizeclass) > 0){
  age.growth.join %>%
  filter(body_cat=="small")%>%
  ggplot()+
    aes(x=year, y=length, color= common_name,linetype = common_name)+
  geom_line(aes(),size=1.4)+
  xlab("Year")+
  ylab(" Total length (cm)")+
  labs(color="Common name of stock",linetype = "Common name of stock")+
  ggtitle("Small bodied fish growth")+
  theme_minimal()
}

```

```{r med_growth_curves, eval=is_med, cache = params$cache}
if(nrow(sizeclass) > 0){
  age.growth.join %>%
    filter(body_cat=="med")%>%
    ggplot()+
      aes(x=year, y=length, color= common_name,linetype = common_name)+
    geom_line(aes(),size=1.4)+
    xlab("Year")+
    ylab(" Total length (cm)")+
    labs(color="Common name of stock",linetype = "Common name of stock")+
    ggtitle("Medium bodied fish growth")+
    theme_minimal()
} 

```

```{r large_growth_curves, eval=is_large, cache = params$cache}
if(nrow(sizeclass) > 0){
  age.growth.join %>%
    filter(body_cat=="large")%>%
    ggplot()+
      aes(x=year, y=length, color= common_name,linetype = common_name)+
    geom_line(aes(),size=1.4)+
    xlab("Year")+
    ylab(" Total length (cm)")+
    labs(color="Common name of stock",linetype = "Common name of stock")+
    ggtitle("large bodied fish growth")+
    theme_minimal()
} 

```

```{r missing_vonb, eval = missing}
 print("NO DATA")
```