## Condition

```{r, cond_functions, cache = params$cache}
source(here::here("R/full_report_functions", "get_lw.R"))
source(here::here("R/full_report_functions", "get_relative_weight.R"))
```

```{r, cond_data_species}
cond_data <- params$cond_data[params$cond_data$Species == species, ]

diet_data <- params$diet_data[params$diet_data$Species == species, ]
```

Condition information comes from [diet data](https://github.com/Laurels1/Condition/blob/master/data/allfh.RData); only regions and seasons with more than 10 fish observations were considered. We calculated a rough condition factor as: Weight / Length^3, and relative weight was [previously calculated](https://github.com/Laurels1/Condition/tree/master/data).

### Figures

#### Length vs weight

Please note, no trend lines were fit, points are jittered to reduce overlap, and sometimes the survey observed a small number of fish outside of the defined stock area.
```{r, lw}
plot_lw(diet_data)
```

#### Condition factor: Weight-volume

If there were more than 30 years of data, a geom_gls() regression was fit. In order to fit the geom_gls() regression, we calculated the mean condition factor for each year and plotted the geom_gls() through those points. Please note, points are jittered to reduce overlap, and sometimes the survey observed a small number of fish outside of the defined stock area.
```{r, condition}
plot_cond(diet_data)
```

#### Condition factor: Relative weight

Please note, this data is aggregated by Ecological Protection Unit (EPU), which may differ slightly from the stock assessment regions.
```{r, laurel_cond}
plot_relw(cond_data)
```

### Data

#### Length vs weight with weight-volume condition factor
```{r, lw_data}
data <- diet_data %>%
  dplyr::filter(is.na(pdlen) == FALSE, 
                  is.na(pdwgt) == FALSE) %>%
  dplyr::select(pdlen, pdwgt, season, Region, fish_id, year) %>%
  dplyr::distinct() %>% # remove duplicates
  dplyr::group_by(Region, season) %>%
  dplyr::mutate(n_fish = length(pdlen)) %>%
  dplyr::filter(n_fish > 10) %>% # only region-season with >10 fish)
  dplyr::mutate(condition = pdwgt/(pdlen^3)) %>%
  dplyr::mutate(condition = condition %>%
                  round(digits = 4)) %>%
  dplyr::select(Region, year, season, pdlen, pdwgt, condition) %>%
  character_to_factor()

make_html_table(data, col_names = c("Region", "Year", "Season", "Fish length (cm)", 
                       "Fish weight (g)", "Condition (g/cm^3)"))
```

#### Relative weight condition factor
Please note, this data is aggregated by Ecological Protection Unit (EPU), which may differ slightly from the stock assessment regions.
```{r, laurel_cond_data}
data <- cond_data %>%
  dplyr::select(-Species, -n) %>%
  dplyr::mutate(MeanCond = MeanCond %>%
                  round(digits = 1)) %>%
  character_to_factor()

make_html_table(data, col_names = c("EPU", "Fish sex", "Year", "Mean condition",
                       "Number of fish"))
```
