## Catch

```{r, child = system.file("indicator_bookdown_template/_legend-child-doc.Rmd", package = "NEesp")}
```

Stock assessment catch data are from ``r params$asmt_source``. Recreational catch data were downloaded from [NOAA MRIP](https://www.st.nmfs.noaa.gov/st1/recreational/MRIP_Estimate_Data/CSV/). Commercial catch data were downloaded from [NOAA FOSS](https://foss.nmfs.noaa.gov/apexfoss/f?p=215:200:4615327020711::NO:::).


```{r}
if (nrow(asmt_data %>% dplyr::filter(Metric == "Catch")) > 0) {
  ndesc <- asmt_data %>%
    dplyr::filter(Metric == "Catch") %>%
    dplyr::mutate(options = paste(Description, Units)) %>%
    dplyr::select(options) %>%
    unique() %>%
    nrow()
} else {
  ndesc <- 1
}
```

### Stock assessment catch


```{r, catch, fig.cap = paste(species, "assessment catch")}
#  fig.height = ndesc * 10
NEesp::plot_asmt(asmt_data,
  metric = "Catch",
  ytitle = "Catch",
  lin = lines,
  col = colors
)
```


#### Risk {-}

See Methods for risk calculation details.
```{r}
indicators <- c("asmt_catch")
```

```{r, child = system.file("indicator_bookdown_template/_risk-child-doc.Rmd", package = "NEesp")}
```

### Recreational catch


```{r, rec, fig.cap = paste(species, "recreational catch")}
NEesp::plot_rec_catch(rec_data)
```


#### Risk {-}

See Methods for risk calculation details.
```{r}
indicators <- c("rec_catch")
```

```{r, child = system.file("indicator_bookdown_template/_risk-child-doc.Rmd", package = "NEesp")}
```

### Commercial catch


```{r, com, fig.cap = paste(species, "commercial catch")}
NEesp::plot_com(com_data)
```


#### Risk {-}

See Methods for risk calculation details.
```{r}
indicators <- c("com_catch")
```

```{r, child = system.file("indicator_bookdown_template/_risk-child-doc.Rmd", package = "NEesp")}
```

### Commercial vs recreational catch


```{r, comvrec, fig.cap = paste(species, "proportional commercial and recreational catch")}
NEesp::plot_prop_catch(com = com_data, rec = rec_data)
```


### Data

#### Stock assessment catch
```{r, catch_data}
data <- asmt_data %>%
  dplyr::filter(Metric == "Catch") %>%
  dplyr::select(-Species, -Metric) %>%
  dplyr::mutate(Value = Value %>%
    round(digits = 0)) %>%
  dplyr::mutate(Value = Value %>%
    format(big.mark = ",")) %>%
  NEesp::character_to_factor()

NEesp::make_html_table(data, 
                       type = params$file)
```

#### Recreational catch
```{r, rec_data}
data <- rec_data %>%
  dplyr::select(year, st_f, mode_fx_f, lbs_ab1) %>%
  dplyr::mutate(
    State = st_f %>%
      stringr::str_to_title(),
    Method = mode_fx_f %>%
      stringr::str_to_sentence(),
    Catch_pounds = lbs_ab1 %>%
      format(big.mark = ",")
  ) %>%
  dplyr::select(year, State, Method, Catch_pounds) %>%
  NEesp::character_to_factor()

NEesp::make_html_table(data, 
                       type = params$file,
                       col_names = c("Year", "State", "Method", "Catch (lb)"))
```

#### Commercial catch
```{r, com_data}
data <- com_data %>%
  dplyr::select(Year, State, Dollars_adj, Pounds) %>%
  dplyr::mutate(
    Dollars_adj = Dollars_adj %>%
      round(digits = 0) %>%
      format(big.mark = ","),
    State = State %>%
      stringr::str_to_title(),
    Pounds = Pounds %>%
      format(big.mark = ",")
  ) %>%
  dplyr::rename(
    "Revenue (2019 dollars)" = "Dollars_adj",
    "Commercial catch (lb)" = "Pounds"
  ) %>%
  NEesp::character_to_factor()


NEesp::make_html_table(data, 
                       type = params$file)
```

#### Commercial vs recreational catch
```{r, comvrec_data}
data <- NEesp::prop_catch_data(rec = rec_data, com = com_data) %>%
  NEesp::character_to_factor()

NEesp::make_html_table(data, 
                       type = params$file,
                       col_names = c(
  "Year", "Recreational catch",
  "Commercial catch", "Total catch",
  "Proportion recreational",
  "Proportion commercial"
))
```

#### Commercial, recreational, and stock assessment catch
```{r, all_catch}
stock <- asmt_data %>%
  dplyr::filter(
    Metric == "Catch",
    Region != "Eastern Georges Bank"
  ) %>% # double counted
  dplyr::mutate(options = paste(Description, Units, AssessmentYear, sep = " - ")) %>%
  dplyr::group_by(Year, options) %>%
  dplyr::summarise(Value = sum(Value)) %>%
  dplyr::ungroup() %>%
  tidyr::pivot_wider(
    names_from = options,
    values_from = Value
  )

recr <- rec_data %>%
  dplyr::mutate(catch = lbs_ab1 %>%
    stringr::str_replace_all(",", "") %>%
    as.numeric()) %>%
  dplyr::group_by(year) %>%
  dplyr::rename(Year = year) %>%
  dplyr::summarise(`Recreational catch (mt)` = sum(catch) / 2204.6)

comm <- com_data %>%
  dplyr::group_by(Year) %>%
  dplyr::summarise(`Commercial catch (mt)` = sum(Pounds) / 2204.6)

all_catch <- c()
if (nrow(stock) > 0) {
  all_catch <- stock
}
if (nrow(recr) > 0) {
  if (is.null(all_catch) == FALSE) {
    all_catch <- dplyr::full_join(all_catch, recr, by = "Year")
  } else {
    all_catch <- recr
  }
}
if (nrow(comm) > 0) {
  if (is.null(all_catch) == FALSE) {
    all_catch <- dplyr::full_join(all_catch, comm, by = "Year")
  } else {
    all_catch <- comm
  }
}

if (is.null(all_catch) == FALSE &
  ncol(all_catch > 2)) {
  all_catch <- all_catch %>%
    dplyr::arrange(Year) %>%
    round(digits = 0) %>%
    # format(big.mark = ",") %>% # messing up formatting
    as.data.frame()
}

# make sure there is data on at least 2 types of catch
test <- c(nrow(stock), nrow(recr), nrow(comm))
sum_test <- sum(test == 0)

if (sum_test >= 2) {
  print("NO DATA")
} else {
  NEesp::make_html_table(data, 
                       type = params$file)
}
```

