# Risk assessment

```{r, risk_functions, cache = params$cache}
source(here::here("R/full_report_functions/risk_functions", "plot_risk_within_species.R"))
source(here::here("R/full_report_functions/risk_functions", "plot_risk_by_year.R"))

# risk data read in in index
```

A preliminary risk analysis was conducted by ranking all species according to their indicator values. A high rank number and a normalized rank near 1 indicates that the species is at risk or of importance based on the measured indicator values. When a species was missing an indicator, it was assigned a normalized rank of 0.5.

## Figures

### Relative to all other stocks

Risk was calculated over time for all indicators that were documented for five or more species in a given year. Risk was calculated as the average of the past 5 years, as a percent of the historical average. The normalized risk value plotted here reflects the normalized rank of this stock compared to all other stocks in that year.

#### Comprehensive risk assessment {-}
```{r, comp_risk_plot}
ggplot(risk_data,
       aes(x = reorder(Region, -total_risk),
           y = Indicator %>% 
             stringr::str_replace_all("_", " "),
           fill = norm_rank))+
  geom_raster(stat = "identity")+
  facet_grid(rows = vars(category),
             scales = "free",
             space = "free")+
  theme_bw()+
  viridis::scale_fill_viridis(limits = c(0, 1),
                              breaks = c(0, 0.5, 1),
                              direction = -1,
                              na.value = "gray90",
                              name = "Normalized rank")+
  theme(legend.position = "top",
        strip.text.y = element_text(angle = 0))+
  xlab("Stock region")+
  ylab("Indicator")
```

#### Normalized rank of magnitude of change compared to historical value by year {-}
```{r, year_risk_hist}
plot_risk_by_year(risk_year_hist_data,
                  indicator = "all",
                  title = "Change compared to historical",
                  include_legend = "yes")
```

#### Normalized rank of value in each year {-}
```{r,year_risk_value}
plot_risk_by_year(risk_year_value_data,
                  indicator = "all",
                  title = "Value compared to other stocks",
                  include_legend = "yes")
```

### Within a single stock

For each stock, a five-year running mean was calculated for each indicator. Indicator values were then ranked for all years where a value was present. The normalized risk values plotted here reflects the normalized rank of each year compared to all other years.

```{r, stock_risk}
plot_risk_by_stock(risk_species_data,
                  indicator = "all",
                  include_legend = "yes")
```

## Data

### Relative to all other stocks

#### Comprehensive risk assessment {-}
```{r, risk_comp}
data <- risk_data %>%
  dplyr::select(Species, Region, Indicator, category, Year, Value, rank, 
                n_stocks_per_indicator, norm_rank, total_risk, overall_rank, 
                overall_stocks) %>%
  dplyr::mutate(Value = Value %>%
                  format(big.mark = ",", scientific = FALSE, digits = 2),
                norm_rank = norm_rank %>%
                  format(scientific = FALSE, digits = 3),
                total_risk = total_risk %>%
                  round(digits = 2))

make_html_table(data, col_names = colnames(data))
```

#### Normalized rank of magnitude of change compared to historical value by year {-}
```{r, risk_hist}
data <- risk_year_hist_data %>% 
  dplyr::mutate(Rank = paste(rank, "out of", n_stocks_per_indicator),
                       Value2 = Value %>% 
                         format(big.mark = ",", scientific = FALSE, digits = 2),
                       Normalized_rank = norm_rank %>% round(digits = 2)) %>%
  dplyr::select(Region, category, Indicator, Year, Value2, Rank, Normalized_rank) %>%
  dplyr::rename("Value" = "Value2") %>%
  character_to_factor()

make_html_table(data, col_names = c("Region", "Category", "Indicator", "Year", "Value (% of historical)",
                       "Rank", "Normalized rank"))
```

#### Normalized rank of value in each year {-}
```{r, risk_year}
data <- risk_year_value_data %>% 
  dplyr::mutate(Rank = paste(rank, "out of", n_stocks_per_indicator),
                       Value2 = Value %>% 
                         format(big.mark = ",", scientific = FALSE, digits = 2),
                       Normalized_rank = norm_rank %>% round(digits = 2)) %>%
  dplyr::select(Region, category, Indicator, Year, Value2, Rank, Normalized_rank) %>%
  dplyr::rename("Value" = "Value2") %>%
  character_to_factor()

make_html_table(data, col_names = c("Region", "Category", "Indicator", "Year", "Value (magnitude)",
                       "Rank", "Normalized rank"))
```

### Value within each stock, ranked by year 
```{r, risk_within}
data <- risk_species_data %>% dplyr::mutate(Rank = paste(rank, "out of", n_years_per_indicator),
                       Value2 = Value %>% 
                         format(big.mark = ",", scientific = FALSE, digits = 2),
                       Normalized_rank = norm_rank %>% round(digits = 2)) %>%
  dplyr::select(Region, category, Indicator, Year, Value2, Rank, Normalized_rank) %>%
  dplyr::rename("Value" = "Value2") %>%
  character_to_factor()

make_html_table(data, col_names = c("Region", "Category", "Indicator", "Year", "Value",
                       "Rank", "Normalized rank"))
```

