## Geographic distribution

Using fisheries independent data from bottom trawls in `survdat`. Several metrics of distribution in observed catches are calculated. Changes in distribution can indicate changes in the geographic range of a species due to a variety of constraints such as food availability or physiologic limitations.The unweighted centroid is one metric that describes the geometric center of the observed range of `r params$species` in a given year. Calculated as the mean of latitude and longitude of all tows that captured the species of interest.The density of observations for `r params$species` was visualized using two-dimensional kernel density estimation on a grid as documented in `MASS::kde2d`. This density estimation is then visualized to indicate areas of greater or lesser probability of occurrence 


```{r geo_conc data wrangling}
source(here::here("R", "common_names_survdat.r"))

survdata.w.codes <- Common_names_survdat(here::here("data", "survdat_03022021_B.RDS"))


selected.surv <- survdata.w.codes %>%
  dplyr::filter(common_name == params$species)


if (nrow(selected.surv > 3)) {
  selected.surv.centroid <- selected.surv %>%
    dplyr::group_by(YEAR) %>%
    dplyr::summarise(lon = mean(LON), lat = mean(LAT))
} else {
  ("Not enough data")
}
```
### Figures

#### Density and distribution estimation


```{r, fig.cap = paste(species, "density and distribution by decade"), fig.height = 12}
if (nrow(selected.surv > 3)) {
  # remove low capture years from this vis to make

  selected.surv <- selected.surv %>%
    dplyr::filter(YEAR > 1970) %>%
    dplyr::mutate(decade = ggplot2::cut_width(YEAR, width = 10, center = 1980))


  world <- rnaturalearth::ne_countries(scale = "medium", returnclass = "sf")
  latmin <- (min(selected.surv.centroid$lat) - 1)
  latmax <- (max(selected.surv.centroid$lat) + 1)
  lonmin <- (min(selected.surv.centroid$lon) - 1)
  lonmax <- (max(selected.surv.centroid$lon) + 1)
 
  
  ggplot2::ggplot(data = world) +
    ggplot2::geom_sf() +
    ggplot2::stat_density2d(aes(x = LON, 
                                y = LAT, 
                                fill = ..level..), 
                            alpha = .5, 
                            geom = "polygon", 
                            data = selected.surv) +
    nmfspalette::scale_fill_nmfs(palette = "crustacean", 
                                 discrete = FALSE, 
                                 reverse = TRUE) +
    ggplot2::coord_sf(xlim = c(lonmin - 1, lonmax + 1), 
                      ylim = c(latmin - 1, latmax + 1)) +
    scale_color_gradient(low = "blue", 
                         high = "red", 
                         name = "Year") +
    xlab("Longitude") +
    ylab("Latitude") +
    theme(legend.position = "none") +
    facet_wrap(~decade, ncol = 2)
  
} else {
  ("Not enough data")
}
```


#### Centroid of observations 

The average position of observations across all years of surveys 


```{r, centroid-plot, fig.cap = paste(species, "centroid over time")}
if(nrow(selected.surv.centroid) > 4) {
  world <- rnaturalearth::ne_countries(scale = "medium", returnclass = "sf")
  latmin <- (min(selected.surv.centroid$lat) - 1)
  latmax <- (max(selected.surv.centroid$lat) + 1)
  lonmin <- (min(selected.surv.centroid$lon) - 1)
  lonmax <- (max(selected.surv.centroid$lon) + 1)


  ggplot2::ggplot(data = world) +
    geom_sf() +
    geom_point(data = selected.surv.centroid, 
               aes(x = lon,
                   y = lat,
                   color = as.numeric(YEAR))) +
    ggrepel::geom_label_repel(data = selected.surv.centroid, 
                              aes(x = lon, y = lat, label = YEAR), 
                              max.overlaps = 20) +
    ggplot2::coord_sf(xlim = c(lonmin - 1, lonmax + 1), 
                      ylim = c(latmin - 1, latmax + 1)) +
    scale_color_gradient(low = "blue", 
                         high = "red", 
                         name = "Year") +
    xlab("Longitude") +
    ylab("Latitude")
  
} else {
  ("Not enough data")
}
```

