---
title: "`r params$species_ID`"
author: "Abigail Tyrell"
date: "`r format(Sys.time(), '%d %b %Y')`"
site: bookdown::bookdown_site
output: 
  bookdown::gitbook:
    split_by: section
documentclass: book
github-repo: NOAA-EDAB/esp_data_aggregation

language:
  ui:
    chapter_name: "Section "

params: 
  species_ID: "Acadian redfish" # common name of species
  
  latlong_data: data # strata
  shape: shape # shapefile
  
  asmt_sum_data: data # bbmsy, ffmsy, assessment ratings
  asmt_sum_source: "assessmentdata::stockAssessmentSummary"

  diet_data: data # allfh diet data
  
  survey_data: data # survey data
  survey_source: "survdat"
  
  ricky_survey_data: data # ricky survey data

  rec_data: data # recreational catch (MRIP)
  
  asmt_data: data # assessmentdata:stockAssessmentData, for recruitment and catch
  asmt_source: "assessmentdata::stockAssessmentData"
  
  cond_data: data # Laurel's condition data
  
  risk_data: data # relative risk rankings
  
  com_data: data # commercial catch
  
  swept_data: data # swept area estimates
  
  risk_year_hist_data: data # historical risk by year
  
  risk_year_value_data: data # value risk by year
  
  risk_species_data: data # species risk
  
  cache: FALSE
  
  path: "" # figure path
---

# `r params$species` {-}

Placeholder



<!--chapter:end:index.Rmd-->


# Methods

Placeholder


## Stock identification
## Data collection and presentation
### `assessmentdata` methods
### `survdat` methods
## Risk assessment
### Risk across stocks
#### Suite of indicators
#### Individual indicators
### Risk within stocks

<!--chapter:end:01-methods.Rmd-->

# Habitat information

<!--chapter:end:02-habitat-title-page.Rmd-->

## Distribution

```{r, geo_functions, cache = params$cache}
source(here::here("R/full_report_functions", "get_latlong.R"))
source(here::here("R/full_report_functions", "map_strata_ecsa.R"))
```

```{r, geo_data_species}
latlong_data <- params$latlong_data[params$latlong_data$Species == species, ]
```

### Map of distribution
Strata maps were pulled and compiled using code from [NOAA/EDAB ECSA](https://github.com/NOAA-EDAB/ECSA). Please note, only fall and spring strata are shown on the map.
```{r, ecsa_map}
map_strata_ecsa(data = latlong_data,
                species_name = species)
```

### Latitude and longitude ranges
Latitude and longitude ranges were calculated from NOAA/EDAB ECSA [seasonal species strata](https://github.com/NOAA-EDAB/ECSA/blob/master/data/seasonal_stock_strata.csv) and [Bottom Trawl Survey (BTS) shapefiles](https://github.com/NOAA-EDAB/ECSA/tree/master/data/strata_shapefiles). The coordinate system is WGS84.
```{r, latlong}
data <- get_latlong(species, latlong_data, shapefile = params$shape) %>%
  character_to_factor()

make_html_table(data, col_names = colnames(data))
```

<!--chapter:end:03-latlong.Rmd-->



## Temperature
### Figures
### Summary
### Data

<!--chapter:end:04-temperature.Rmd-->



## Salinity
### Figures
### Summary
### Data

<!--chapter:end:05-salinity.Rmd-->



## Depth
### Figures

<!--chapter:end:06-depth.Rmd-->

## Habitat vulnerability

```{r, vul_functions}
source(here::here("R/full_report_functions", "plot_vulnerability.R"))

data <- read.csv(here::here("data", "Hare_et_al_2016_overall.csv"))
```

Habitat vulnerability information is sourced from the `ecodata` package. 

```{r, hab_vul}
data <- ecodata::habitat_vulnerability %>%
  dplyr::filter(Species == species) %>%
  dplyr::select(-Species)

data <- data[,c(8, 1:7)]

make_html_table(data,
                col_names = c("EPU", "Habitat name", "Habitat vulnerability", 
                             "Importance of habitat to eggs/larvae",
                             "Importance of habitat to juveniles/YOY",
                             "Importance of habitat to adults",
                             "Importance of habitat to spawning adults",
                             "Overall species vulnerability"))

```

<!--chapter:end:07-habitat-vulnerability.Rmd-->

# Biological information

<!--chapter:end:08-biological-title-page.Rmd-->



## Length
### Figures
#### Overview {-}
#### Summary statistics {-}
#### Risk {-}
### Summary
### Data

<!--chapter:end:09-length.Rmd-->



## von Bertalanffy growth curve
### Length at age growth curve

<!--chapter:end:10-von_b.Rmd-->


## L50 maturity


Ricky Tabandera

```{r setup_L50, cache = params$cache}
knitr::opts_chunk$set(echo = TRUE)

suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(nlstools))
library(dbutils)
library(survdat)

```

## Size at first maturity 

L50 or length at which 50%  maturity can be calculated using differing methods. Using Data sourced from `survdat` that consist of visual sexual maturity determination and length measurements. Here, L50 is estimated by fitting a logistic generalized linear model to gonadal maturity and body length measurements and estimating the inflection point of the logistic model. This point represents the size where the individual has 50% odds of being mature. Decade variations of this parameter are investigated utilizing decade as a additive and interaction term in the model specification.

The relationship between sexual maturity by body size is a well established life history parameter. L50 is a useful metric in fisheries to identify sizes that are able to reproduce. This information can be used to inform regulation as to minimum catch size that should allow for a significant portion of the population to spawn and contribute genetic information to sustain the stock. Changes in this parameter can signal population/evolutionary pressures on the stock with reduction in this size potentially indicating excessive fishing pressure.



```{r data wrangling, cache = params$cache, cache.lazy = FALSE}

survdata<-readRDS(here::here("data", "survdat_pull_bio.rds"))

survdata<-survdata %>% mutate(SVSPP=as.numeric(SVSPP))
stock_list_all_strata <- read.csv("https://raw.githubusercontent.com/NOAA-EDAB/ECSA/master/data/stock_list.csv")
stock_list_all_strata<-stock_list_all_strata %>% rename(SVSPP=svspp)
stock_list<-stock_list_all_strata %>% dplyr::distinct(SVSPP,.keep_all =T)

survdata.w.codes<-inner_join(survdata, stock_list, by= "SVSPP" )



selected.surv<-survdata.w.codes %>%
  dplyr::mutate(common_name = common_name %>%
                  stringr::str_to_sentence()) %>%
  dplyr::filter(common_name == species & YEAR>1980)

#con <- connect_to_database(server="sole",uid="RTABANDERA")
#surv.maturity<-get_maturity(con)
#surv_maturity<-surv.maturity$data
#write.csv(surv.maturity, file = "surv_maturity_codes.csv")



#only take records with a maturity status associated

selected.surv.clean<-selected.surv %>% drop_na(MATURITY) 



#recode sexed fish to readable names
#sex codes 1 male, 2 female, 4 transitional, 0 unsexed/unknown   
selected.surv.clean$MATURITY <-recode(selected.surv.clean$MATURITY, "D"="Mature" , "I"="Immature", "R"="Mature", "S"= "Mature", "T"= "Mature", "U"="Mature")
selected.surv.f<- selected.surv.clean %>% filter(SEX=="female" | MATURITY !="X")

# conditionally run models based on having more than 3 rows 
if(nrow(selected.surv.f)>3){



# X= unknown , I = immature, D= developing, R S T U convert to developed


#bin the years of fish into decades to compare if model parameters are chaging 
selected.surv.f<-selected.surv.f %>% mutate(decade=ggplot2::cut_width(YEAR, width = 10, boundary = 1980))
levels(selected.surv.f$decade)<-c( "1980","1990", "2000","2010")
#recode the maturity col into factor
selected.surv.f<-selected.surv.f %>% mutate(MATURITY=as.factor(MATURITY))


maturity.mod<-glm(MATURITY~LENGTH+decade+LENGTH*decade, data=selected.surv.f, family=binomial)



pretty_mod<-papeR::prettify(summary(maturity.mod))

#summary(maturity.mod)  
}else("Not enough data to fit model")
  




```

## Model results

`r if(nrow(selected.surv.f)>3){"If the overall model is significant, there is support for the relationship between sexual maturity and body size. This model can predict the size where there are 50% odds of the indevidual is sexually mature. The results table below displays the L50 for different decades. "}` 

```{r, model table, echo=FALSE, message=FALSE, warning=FALSE}

if(nrow(selected.surv.f)>3){
 
knitr::kable(pretty_mod, digits = 3)
}
```

## L50 across time

`r if(nrow(selected.surv.f)>3){"The significance of differences among decades can be determined by referencing the model summary table above   "}`

```{r model results and L50}
if(nrow(selected.surv.f)>3){
   if(summary(maturity.mod)$coefficients[,4][1]<0.05){


L50.1980<-(-coef(maturity.mod)[1])/coef(maturity.mod)[2] # 1980
L50.1990<-(-(coef(maturity.mod)[1]+coef(maturity.mod)[3]))/(coef(maturity.mod)[2]+coef(maturity.mod)[6]) # 1990
L50.2000<-(-(coef(maturity.mod)[1]+coef(maturity.mod)[4]))/(coef(maturity.mod)[2]+coef(maturity.mod)[7]) # 2000
L50.2010<-(-(coef(maturity.mod)[1]+coef(maturity.mod)[5]))/(coef(maturity.mod)[2]+coef(maturity.mod)[8]) # 2010

L50<-data.frame(L50.1980,L50.1990,L50.2000,L50.2010)
row.names(L50)<-c("Length at 50% maturity")
colnames(L50)<-c("1980","1990","2000","2010")
L50.p<-papeR::prettify(L50)
knitr::kable(L50.p, digits = 3)
}}else("Non-significant")


```




```{r, probability plot}

if(nrow(selected.surv.f)>3){
   if(summary(maturity.mod)$coefficients[,4][1]<0.05){
decade1980<-boot::inv.logit(predict(maturity.mod, data.frame(LENGTH=seq(min(selected.surv.f$LENGTH),max(selected.surv.f$LENGTH)), decade=rep("1980",length(seq(min(selected.surv.f$LENGTH),max(selected.surv.f$LENGTH)))), type="response")))
decade1990<-boot::inv.logit(predict(maturity.mod, data.frame(LENGTH=seq(min(selected.surv.f$LENGTH),max(selected.surv.f$LENGTH)), decade=rep("1990",length(seq(min(selected.surv.f$LENGTH),max(selected.surv.f$LENGTH)))), type="response")))
decade2000<-boot::inv.logit(predict(maturity.mod, data.frame(LENGTH=seq(min(selected.surv.f$LENGTH),max(selected.surv.f$LENGTH)), decade=rep("2000",length(seq(min(selected.surv.f$LENGTH),max(selected.surv.f$LENGTH)))), type="response")))
decade2010<-boot::inv.logit(predict(maturity.mod, data.frame(LENGTH=seq(min(selected.surv.f$LENGTH),max(selected.surv.f$LENGTH)), decade=rep("2010",length(seq(min(selected.surv.f$LENGTH),max(selected.surv.f$LENGTH)))), type="response")))
selected.surv.LENGTH<-seq(min(selected.surv.f$LENGTH),max(selected.surv.f$LENGTH))

selected.surv.simlulation<-data.frame(decade1980, decade1990, decade2000,decade2010,selected.surv.LENGTH)
selected.surv.simlulation<-selected.surv.simlulation %>% pivot_longer(!selected.surv.LENGTH, names_to="decade", values_to="pred")


selected.surv.simlulation %>%
  ggplot(aes(x=selected.surv.LENGTH,y=pred, color=decade))+
  geom_line(size=3)+
  xlab("Length (cm)")+
  ylab("Probability")+
  ggtitle("Length at 50% maturity of" , subtitle = params$species)
  annotate("text", label=c("Mature","Immature"), x=30,y= c(1.05,-.05))+
  geom_hline(yintercept=c(1,0,0.5))+
  scale_color_manual(labels=c("<1980"," 1981-1990","1991-2000",">2010"), values =c( "blue","firebrick","darkviolet","deepskyblue"), name="Decade")+
  theme_minimal()
}}else("Non-significant")


```



<!--chapter:end:11-length50maturity.Rmd-->



## Condition
### Figures
#### Length vs weight
#### Condition factor: Weight-volume
#### Condition factor: Relative weight
### Data
#### Length vs weight with weight-volume condition factor
#### Relative weight condition factor

<!--chapter:end:12-condition.Rmd-->



## Diet
### Figure
### Summary
### Data

<!--chapter:end:13-diet.Rmd-->

# Population information

<!--chapter:end:14-population-title-page.Rmd-->



## Abundance
### Figures
#### Survey abundance (raw measurements)
##### Risk {-}
#### Survey abundance (swept area estimates)
#### Assessment abundance
##### Risk {-}
### Survey summary
### Data
#### Survey data (raw measurements)
#### Survey data (swept area estimates)
#### Assessment data

<!--chapter:end:15-abundance.Rmd-->



## Biomass
### Figures
#### Survey biomass (raw measurements)
##### Risk {-}
#### Survey biomass (swept area estimates)
#### Assessment biomass
##### Risk {-}
### Survey summary
### Data
#### Survey data (raw measurements)
#### Survey data (swept area estimates)
#### Assessment data

<!--chapter:end:16-biomass.Rmd-->



## B/Bmsy 
### Figure
##### Risk {-}
### Data

<!--chapter:end:17-bbmsy.Rmd-->



## Recruitment
### Figure
##### Risk {-}
### Data

<!--chapter:end:18-recruitment.Rmd-->



## Age diversity
### Figures
#### Age diversity {-}
#### Density plots of age {-}

<!--chapter:end:19-age-diversity.Rmd-->



## Climate vulnerability
### Figures
### Data

<!--chapter:end:20-climate-vulnerability.Rmd-->

# Socio-economic information

<!--chapter:end:21-socio-economic-title-page.Rmd-->



## Catch
### Figures
#### Stock assessment catch
##### Risk {-}
#### Recreational catch
##### Risk {-}
#### Commercial catch
##### Risk {-}
#### Commercial vs recreational catch
### Data
#### Stock assessment catch
#### Recreational catch
#### Commercial catch
#### Commercial vs recreational catch
#### Commercial, recreational, and stock assessment catch

<!--chapter:end:22-catch.Rmd-->



## F/Fmsy 
### Figure
##### Risk {-}
### Data

<!--chapter:end:23-ffmsy.Rmd-->



## Revenue 
### Figure
##### Risk {-}
### Data

<!--chapter:end:24-revenue.Rmd-->


# Management information

Placeholder


## Stock assessment and data quality information

<!--chapter:end:25-management.Rmd-->


# Risk assessment

Placeholder


## Figures
### Relative to all other stocks
#### Comprehensive risk assessment {-}
#### Normalized rank of magnitude of change compared to historical value by year {-}
#### Normalized rank of value in each year {-}
### Within a single stock
## Data
### Relative to all other stocks
#### Comprehensive risk assessment {-}
#### Normalized rank of magnitude of change compared to historical value by year {-}
#### Normalized rank of value in each year {-}
### Value within each stock, ranked by year 

<!--chapter:end:26-risk-assessment.Rmd-->

