### Recreational landings


```{r, rec, fig.cap = paste(species, "recreational catch")}
NEesp::plot_rec_catch(rec_data)
```


#### Risk {-}

See Methods for risk calculation details.
```{r}
indicators <- c("rec_catch")
```

```{r, child = system.file("indicator_bookdown_template/_risk-child-doc.Rmd", package = "NEesp")}
```


```{r rec-data-cleaning}
# effort data
rec_effort <- NEesp::rec_effort

# read in recreational landings data
rec_landings_1950_2019 <- NEesp::rec_catch %>%
  dplyr::rename(
    State = st_f,
    Year = year,
    Pounds = lbs_ab1
  )

Selected_sp <- rec_landings_1950_2019 %>%
  dplyr::group_by(Year) %>%
  dplyr::filter(!is.na(Pounds) & Species == params$species) %>%
  dplyr::summarise(total = sum(Pounds, na.rm = FALSE))

selected_by_state <- rec_landings_1950_2019 %>%
  dplyr::filter(!is.na(Pounds) & Species == params$species) %>%
  dplyr::group_by(State, Year)


sp_data_qual <- TRUE
st_data_qual <- TRUE

# testing if data quality is sufficient for each graph
if (length(which(is.na(Selected_sp$total))) / length(Selected_sp$total) > 0.40) {
  sp_data_qual <- FALSE
} else {
  sp_data_qual <- TRUE
}


if (length(which(is.na(selected_by_state$Pounds))) / length(selected_by_state$Pounds) > 0.50) {
  st_data_qual <- FALSE
} else {
  st_data_qual <- TRUE
}
```

#### Recreational catches time series 

This data was sourced from [FOSS - Fisheries One Stop Shop](https://foss.nmfs.noaa.gov/apexfoss/f?p=215:200:4615327020711::NO:::) and is inclusive of `r min(rec_landings_1950_2019$Year,na.rm =TRUE) ` to `r max(rec_landings_1950_2019$Year,na.rm =TRUE)`. The entire data set contains `r length(unique(rec_landings_1950_2019$common_name))` species. Figures produced reflect the coverage of the data. With stocks that have high coverage, a running average is calculated. In low coverage stocks, missing values are excluded and a simple time series is produced. 


```{r total-catch-good-data, eval= sp_data_qual}

Selected_sp %>%
  ggplot2::ggplot(ggplot2::aes(x = Year, y = log(total))) +
  tidyquant::geom_ma(n = 5, lwd = 1.4) +
  ggplot2::geom_point() +
  ggplot2::ggtitle(label = params$species, subtitle = "5 year rolling average") +
  ggplot2::xlim(1981, 2019) +
  ggplot2::xlab(label = "Year") +
  ggplot2::ylab(label = "Log transformed  total catch in pounds (lb) ") +
  ggplot2::theme_minimal()
```


```{r sp-if-data-is-poor, eval= !sp_data_qual}

Selected_sp %>%
  data.table::na.omit() %>%
  ggplot2::ggplot(ggplot2::aes(x = Year, y = log(total))) +
  ggplot2::geom_point(size = 1.3) +
  ggplot2::geom_line() +
  ggplot2::ggtitle(label = params$species, subtitle = "Data coverage is poor") +
  ggplot2::xlim(1981, 2019) +
  ggplot2::xlab(label = "Year") +
  ggplot2::ylab(label = "Log transformed  total catch in pounds (lb)") +
  ggplot2::theme_minimal()
```


#### Recreational landings of `r params$species` across states in the mid-Atlantic and north-east regions

```{r, state-breakdown-data-good,eval=st_data_qual}

selected_by_state %>%
  ggplot2::ggplot(ggplot2::aes(x = Year, y = log(Pounds))) +
  ggplot2::geom_point() +
  tidyquant::geom_ma(ggplot2::aes(color = State, lty = State), n = 5, lwd = 1.4, na.rm = TRUE, show.legend = FALSE) +
  ggplot2::facet_wrap(ggplot2::vars(State)) +
  ggplot2::ggtitle(label = params$species, subtitle = "5 year rolling average") +
  ggplot2::xlim(1981, 2019) +
  ggplot2::xlab(label = "Year") +
  ggplot2::ylab(label = "Log transformed  total catch in pounds (lb) ") +
  ggplot2::theme_minimal()
```

```{r data-poor-graph, eval= !st_data_qual}

selected_by_state %>%
  ggplot2::ggplot(ggplot2::aes(x = Year, y = log(Pounds))) +
  ggplot2::geom_point(ggplot2::aes(color = State), show.legend = FALSE) +
  ggplot2::geom_line(ggplot2::aes(color = State, lty = State), lwd = 1.3, show.legend = FALSE) +
  ggplot2::facet_wrap(ggplot2::vars(State)) +
  ggplot2::xlim(1981, 2019) +
  ggplot2::ggtitle(label = params$species, subtitle = "Data coverage is poor") +
  ggplot2::xlab(label = "Year") +
  ggplot2::ylab(label = "Log transformed  total catch in pounds (lb) ") +
  ggplot2::theme_minimal()
```


#### Visualizing Recreational fishing effort

These records are Recreational Fisheries Statistics sourced from the [ Marine Recreational Information Program (MRIP) ](https://www.st.nmfs.noaa.gov/recreational-fisheries/data-and-documentation/queries/index) via NOAA Office of Science and Technology and are inclusive of all north and mid-Atlantic states. The records span from `r min(rec_effort$Year )` to `r max(rec_effort$Year) ` and are specific to the mode and location of effort. Effort is quantified as angler trips. Mode is divided into `r stringr::str_to_sentence(levels(rec_effort$Fishing.Mode))`.


```{r effor-by-state,fig.width=11,fig.height =8.5}
by_state_mode <- rec_effort %>%
  dplyr::group_by(State, Fishing.Mode, Year) %>%
  dplyr::summarise(total_trips = sum(Angler.Trips))

by_state_mode$Fishing.Mode <- stringr::str_to_sentence(by_state_mode$Fishing.Mode)

by_state_mode %>% ggplot2::ggplot(ggplot2::aes(x = Year, y = (total_trips / 100000))) +
  ggplot2::geom_line(ggplot2::aes(color = Fishing.Mode), size = 1.3) +
  ggplot2::ggtitle("Fishing effort by mode across states") +
  ggplot2::xlab("Year") +
  ggplot2::ylab("Annual recreational trips in 100,000 units") +
  ggplot2::facet_grid(cols = ggplot2::vars(State), rows = ggplot2::vars(Fishing.Mode)) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90), strip.text.x = ggplot2::element_text(size = 7, angle = 30))
```

```{r effort by location, echo=FALSE, message=FALSE,warning=FALSE,fig.width=11,fig.height =8.5}

by_location <- rec_effort %>%
  dplyr::group_by(Year, Fishing.Area, State) %>%
  dplyr::summarise(Total_effort = sum(Angler.Trips))

by_location %>% ggplot2::ggplot(ggplot2::aes(x = Year, y = (Total_effort / 100000))) +
  tidyquant::geom_ma(ggplot2::aes(color = Fishing.Area), size = 1.3, n = 5) +
  ggplot2::ggtitle("Recreational Fishing effort by location ", subtitle = " 4 year rolling average") +
  ggplot2::xlab("Year") +
  ggplot2::ylab("Annual recreational trips in 100,000 units") +
  ggplot2::theme_minimal() +
  ggplot2::facet_grid(rows = ggplot2::vars(Fishing.Area), cols = ggplot2::vars(State)) +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90), strip.text.x = ggplot2::element_text(size = 7, angle = 30))
```

```{r summary by state, echo=FALSE, message=FALSE,error=FALSE,fig.width=11,fig.height =8.5}
# group effort by state mode and year for first summary

by_state <- rec_effort %>%
  dplyr::group_by(State, Year) %>%
  dplyr::summarise(total_trips = sum(Angler.Trips))

by_state %>% ggplot2::ggplot(ggplot2::aes(x = Year, y = (total_trips / 100000))) +
  tidyquant::geom_ma(ggplot2::aes(color = State, linetype = State), size = 1.3, n = 5) +
  ggplot2::ggtitle("Recreational Fishing effort by state ", subtitle = "5 year rolling average") +
  ggplot2::xlab("Year") +
  ggplot2::ylab("Annual recreational trips in 100,000 units")
```
