
### Commercial fishing gear 

Fishing gear can range widely in their catch ability, selectivity, and effective areas of operations. Using `comlandr` to extract reported gear used in the many commercial fisheries in the northeast and any trends and/or shifts that may have occurred as one potential explanatory factor on changes in fisheries landings/revenue.  


```{r gear-wrangling}

com_gear <- NEesp::com_gear
sp_group <- NEesp::sp_group

sp_group <- sp_group %>%
  dplyr::filter(!is.na(SVSPP)) %>%
  dplyr::select(COMNAME, SVSPP, ITISSPP, NESPP3, SCINAME)
sp_group <- sp_group %>% dplyr::mutate_at(c("COMNAME", "SCINAME"), .funs = stringr::str_to_sentence)


key_join <- dplyr::left_join(com_gear, sp_group, by = "NESPP3")
```

#### Gear usage by year

Each landing record included gear type used, size class of gear, ecological production unit, species landed, weight of catch

```{r, gear-raw}

key_join <- key_join %>% tidyr::drop_na()
if (nrow(key_join) > 6) {
  key_join %>%
    dplyr::group_by(YEAR, GEAR) %>%
    dplyr::summarise(gear.count = length(GEAR)) %>%
    ggplot2::ggplot(ggplot2::aes(x = YEAR, y = gear.count, color = GEAR)) +
    ggplot2::geom_point() +
    ggplot2::geom_line() +
    nmfspalette::scale_color_nmfs(palette = "regional web") +
    ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90)) +
    ggplot2::xlab("Year") +
    ggplot2::ylab("# gear type records")
} else {
  ("No Data")
}
```

#### Gear diversity all species 
```{r, gear-diversity}

gear.diversity <- key_join %>%
  dplyr::group_by(YEAR, EPU, GEAR) %>%
  dplyr::summarise(gear.count = (length(GEAR)))

if (nrow(gear.diversity) > 5) {
  gear.diversity <- gear.diversity %>%
    dplyr::group_by(YEAR, EPU) %>%
    dplyr::mutate(gear.prop = (gear.count / sum(gear.count)))


  gear.H <- gear.diversity %>%
    dplyr::group_by(YEAR, EPU) %>%
    dplyr::summarise(shannon = -sum(gear.prop * log(gear.prop)))

  gear.H %>% ggplot2::ggplot(ggplot2::aes(x = YEAR, y = shannon, color = EPU)) +
    ggplot2::geom_point() +
    ggplot2::geom_line() +
    nmfspalette::scale_color_nmfs(palette = "regional web") +
    ggplot2::xlab("Year") +
    ggplot2::ylab("gear diversity index (H')")
} else {
  ("No data")
}
```


#### Commercial gear types used to land `r params$species`

```{r, gear-by-species}

gear.spp <- key_join %>% dplyr::filter(COMNAME == params$species)

if (nrow(gear.spp) > 5) {
  gear.spp %>%
    dplyr::group_by(YEAR, EPU) %>%
    dplyr::mutate(gear.count = length(GEAR)) %>%
    ggplot2::ggplot(ggplot2::aes(x = YEAR, y = gear.count, fill = GEAR)) +
    ggplot2::geom_bar(stat = "identity") +
    ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90)) +
    ggplot2::labs(fill = "Gear Type") +
    nmfspalette::scale_fill_nmfs(palette = "regional web") +
    ggplot2::xlab("Year") +
    ggplot2::ylab("# of landings") +
    ggplot2::ggtitle(label = params$species)
  # scale_fill_discrete(name = "Gear Type")
} else {
  ("No data")
}
```

#### Diversity index of gear type across EPUs


#### Diversity index of gear used to land `r params$species` across EPUs


```{r, gear-diversity-by-species}

spp.diversity <- gear.spp %>%
  dplyr::group_by(YEAR, GEAR, EPU) %>%
  dplyr::summarise(gear.count = (length(GEAR)))

spp.diversity <- spp.diversity %>%
  dplyr::group_by(YEAR, EPU) %>%
  dplyr::mutate(gear.prop = (gear.count / sum(gear.count)))


spp.H <- spp.diversity %>%
  dplyr::group_by(YEAR, EPU) %>%
  dplyr::summarise(shannon = -sum(gear.prop * log(gear.prop)))
if (nrow(spp.H) > 5) {
  spp.H %>%
    ggplot2::ggplot(ggplot2::aes(x = YEAR, y = shannon, color = EPU)) +
    ggplot2::geom_point() +
    ggplot2::geom_line() +
    nmfspalette::scale_color_nmfs(palette = "regional web") +
    ggplot2::xlab("Year") +
    ggplot2::ylab("gear diversity index (H')") +
    ggplot2::ggtitle(label = params$species)
} else {
  ("No data")
}
```

