
### Commercial landings

Ricky Tabandera


```{r com-setup,}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(priceR)
```

```{r, com-data-processing}
# list.files(here::here("commercial"))

# archival copy of data wrangleing of raw foss data used to produce the clean dataset

# #load in the stocks managed in NEFSC
# stock_list_all_strata <- read.csv("https://raw.githubusercontent.com/NOAA-EDAB/ECSA/master/data/stock_list.csv")
# #lowercase all of the names
# stock_list_all_strata$common_name <- stringr::str_to_lower(stock_list_all_strata$common_name)
# #extract out the unique names and rename column
# stock_list<-stock_list_all_strata %>% distinct(common_name, .keep_all=TRUE)
#
#
# #load commercial landings data
# com_landings<-readr::read_csv(here::here("data","foss_ commercial_landings_NE_MA_1950_2019.csv"))
# #renaming cols to aling with stocklist
# com_landings<-com_landings %>% rename(common_name=`NMFS Name`)
# #simplifying the names to be lowercase
# com_landings$common_name<-stringr::str_to_lower(com_landings$common_name)
#
# #removing empty rows due to confidentiality and counting the number of records removed
# confid<-com_landings %>%
#   filter(common_name =="withheld for confidentiality" | Confidentiality == "Confidential" )%>%
#   group_by(Year) %>%
#   summarise(count=n())
# com_landings<-com_landings %>% filter(!common_name =="withheld for confidentiality" , !Confidentiality == "Confidential", !is.na(Pounds), !is.na(Dollars), Dollars >1)
#
#
# # removing aggrigated landings records and counting their occurance
# grouped_landings<-com_landings %>% filter(stringr::str_detect(common_name, "\\*+"))
#
# # rearranging the species names to align with NEFSC conventions
# com_landings<-com_landings %>% filter(!stringr::str_detect(common_name, "\\*+"))
#
# #splitting off the records without a comma
#
#
# single_name<-com_landings %>% filter(!stringr::str_detect(common_name, "\\,\\s"))
# multipe_name<-com_landings %>% filter(stringr::str_detect(common_name, "\\,\\s"))
# split_names<-multipe_name%>% separate (col=common_name,sep ="\\,\\s", into = c("first" ,"second", "third"))
# #fixing plaice to remove flounder
# split_names<- split_names %>% dplyr::mutate(first=replace(first, which(second=="american plaice"), NA))
# #fixing the shark tag for dogfish
# split_names<- split_names %>% dplyr::mutate(first=replace(first, which(second=="dogfish"), NA))
# #longfin loligo into inshore
# split_names<- split_names %>% dplyr::mutate(second=replace(second, which(second=="longfin loligo"), "longfin inshore"))
# #shortfin illex into northern shortfin
# split_names<- split_names %>% dplyr::mutate(second=replace(second, which(second=="shortfin illex"), "northern shortfin"))
# #goosefish into monkfish
# single_name<- single_name %>% dplyr::mutate(common_name=replace(common_name, which(common_name=="goosefish"), "monkfish"))
# #adding flounder to windowpane
# single_name<- single_name %>% dplyr::mutate(common_name=replace(common_name, which(common_name=="windowpane"), "windowpane flounder"))
# split_names<-split_names %>% unite(col = "last", c(third, second),sep = " ", remove = TRUE ,na.rm = TRUE)
# split_names<-split_names %>% unite(col = "common_name", c(last, first),sep = " ", remove = TRUE ,na.rm = TRUE)
#
# joined_names<- dplyr::bind_rows(split_names,single_name)
# #used to check that there are no missing stocks in the data
# #missing_stocks<-dplyr::anti_join(stock_list,joined_names, by= "common_name")
#
#
# #final clean dataset
# com_landings_clean<-dplyr::full_join(stock_list,joined_names, by= "common_name")
# write.csv(com_landings_clean, "FOSS_com_landings_clean_20201222.csv")

com_landings_clean <- read.csv(here::here("data", "FOSS_com_landings_clean_20201222.csv"))
com_landings_clean <- com_landings_clean %>% mutate(common_name = stringr::str_to_sentence(common_name))


selected_com_sp <- com_landings_clean %>%
  dplyr::group_by(Year) %>%
  filter(common_name == params$species) %>%
  summarise(total_pounds = sum(Pounds, na.rm = FALSE), total_dollars = sum(Dollars, na.rm = FALSE))

selected_com_state <- com_landings_clean %>%
  dplyr::group_by(Year, State) %>%
  filter(common_name == params$species) %>%
  summarise(total_pounds = sum(Pounds, na.rm = FALSE), total_dollars = sum(Dollars, na.rm = FALSE))

##########################

# red hake
data_qual <- TRUE

num_states <- selected_com_state %>%
  group_by(State) %>%
  count()

if (min(num_states$n) < 5) {
  data_qual <- FALSE
} else {
  data_qual <- TRUE
}
```


#### Commercial Landings of `r params$species` 

This data was sourced from [FOSS - Fisheries One Stop Shop](https://foss.nmfs.noaa.gov/apexfoss/f?p=215:200:4615327020711::NO:::) and is inclusive of `r min(com_landings_clean$Year) ` to `r max(com_landings_clean$Year)`. The entire data set contains `r length(unique(com_landings_clean$common_name))` species. A total of 9656 aggregated records were removed. These landing reports were aggregations of multiple species and can span several taxonomic groups were removed due to their lack of specificity. 5019 records were withheld due to confidentiality of those specific landings.
Figures produced reflect the coverage of the data. With stocks that have high coverage, a running average is calculated. In low coverage stocks, missing values are excluded and a simple time series is produced. 


```{r, landings-single-stock}


if (nrow(selected_com_sp) > 4) {
  selected_com_sp %>%
    ggplot2::ggplot(aes(x = Year, y = (total_dollars / 10^6))) +
    ggplot2::geom_point() +
    tidyquant::geom_ma(n = 5, lwd = 1.4) +
    ggtitle(label = params$species, subtitle = "5 year rolling average") +
    xlab(label = "Year") +
    ylab(label = "Dollar value of landings in millions ($) ") +
    theme_minimal()
} else {
  ("No data")
}
```

#### Commercial landings of `r params$species` across states in the mid-Atlantic and north-east regions

```{r landings-across-states-good-data,  eval=data_qual}


selected_com_state %>%
  ggplot2::ggplot(aes(x = Year, y = (total_dollars / 10^6))) +
  tidyquant::geom_ma(aes(color = State, lty = State), n = 3, lwd = 1.4) +
  ggtitle(label = params$species, subtitle = "5 year rolling average") +
  xlab(label = "Year") +
  ylab(label = "Dollar value of landings in millions ($)  ") +
  theme_minimal()
```


```{r, landings-across-states-bad-data, eval=!data_qual}


selected_com_state %>%
  ggplot2::ggplot(aes(x = Year, y = (total_dollars / 10^6))) +
  geom_point(aes(color = State)) +
  geom_line(aes(color = State)) +
  ggtitle(label = params$species, subtitle = "Data coverage for some states is poor with >5 years reported") +
  xlab(label = "Year") +
  ylab(label = "Dollar value of landings in millions ($)  ") +
  theme_minimal()
```




#### Non-inflation adjusted price history of `r params$species`
The calculated price is dollars per pound of reported landings. 



```{r Non-inflation-adjusted-price-history  }

if (nrow(selected_com_sp) > 4) {
  selected_com_sp %>%
    dplyr::mutate(price_per_lb = (total_dollars / total_pounds)) %>%
    ggplot2::ggplot(aes(x = Year, y = price_per_lb)) +
    tidyquant::geom_ma(n = 5, lwd = 1.4) +
    ggtitle(label = params$species, subtitle = "5 year rolling average ") +
    xlab(label = "Year") +
    ylab(label = "Price per pound (lb/$) ") +
    theme_minimal()
} else {
  ("Not enough data")
}
```

#### Non-inflation adjusted price history of `r params$species` across states in the mid-Atlantic and north-east regions
The calculated price is dollars per pound of reported landings. 


```{r Non-inflation-adjusted-price-history-across-states-good-data , eval=data_qual}

if (nrow(selected_com_state) > 4) {
  selected_com_state %>%
    dplyr::mutate(price_per_lb = (total_dollars / total_pounds)) %>%
    ggplot2::ggplot(aes(x = Year, y = price_per_lb)) +
    tidyquant::geom_ma(aes(color = State, lty = State), n = 5, lwd = 1.4, show.legend = FALSE) +
    facet_wrap(vars(State)) +
    ggtitle(label = params$species, subtitle = "5 year rolling average ") +
    xlab(label = "Year") +
    ylab(label = "Price per pound (lb/$) ") +
    theme_minimal()
} else {
  ("Not enough data")
}
```

```{r Non-inflation-adjusted-price-history-across-states-poor-data , eval=!data_qual}

if (nrow(selected_com_state) > 4) {
  selected_com_state %>%
    dplyr::mutate(price_per_lb = (total_dollars / total_pounds)) %>%
    ggplot2::ggplot(aes(x = Year, y = price_per_lb)) +
    geom_point(aes(color = State), show.legend = FALSE) +
    geom_line(aes(color = State), show.legend = FALSE) +
    facet_wrap(vars(State)) +
    ggtitle(label = params$species, subtitle = "Data coverage for some states is poor with >5 years reported") +
    xlab(label = "Year") +
    ylab(label = "Price per pound (lb/$) ") +
    theme_minimal()
} else {
  ("Not enough data")
}
```

#### inflation adjusted price history of `r params$species`
The price history  adjusted to 2019 values using data from the World Bank inflation data as informed by Inflation adjustment calculations in Principles of Macroeconomics by Gregory Mankiw et al (2014). Adjustments performed using the packages [priceR](https://github.com/stevecondylios/priceR).

```{r, inflation-adjusted-price-history }


stock_price <- selected_com_state %>%
  group_by(Year) %>%
  dplyr::mutate(price_per_lb = (total_dollars / total_pounds))
country <- "US"
stock_price$adj_price <- adjust_for_inflation(stock_price$price_per_lb, stock_price$Year, country, to_date = 2019)


if (nrow(stock_price) > 5) {
  stock_price %>%
    group_by(Year) %>%
    summarise(adj_ave_price = mean(adj_price)) %>%
    ggplot2::ggplot(aes(x = Year, y = adj_ave_price)) +
    tidyquant::geom_ma(n = 5, lwd = 1.4) +
    ggtitle(label = params$species, subtitle = "5 year rolling average ") +
    xlab(label = "Year") +
    ylab(label = "Price per pound (lb/$) in 2019 value ") +
    theme_minimal()
} else {
  ("Not enough data")
}
```


#### inflation adjusted price history of `r params$species` across states in the mid-Atlantic and north-east regions 



```{r, inflation-adjusted-price-across-states-good-data, eval=data_qual}
if (nrow(stock_price) > 3) {
  stock_price %>%
    ggplot2::ggplot(aes(x = Year, y = adj_price)) +
    tidyquant::geom_ma(aes(color = State), n = 5, lwd = 1.4, show.legend = FALSE) +
    facet_wrap(vars(State)) +
    ggtitle(label = params$species, subtitle = "5 year rolling average ") +
    xlab(label = "Year") +
    ylab(label = "Price per pound (lb/$) in 2019 value ") +
    theme_minimal()
} else {
  ("Not enough data")
}
```





```{r, inflation-adjusted-price-across-states-poor-data, eval=!data_qual}
if (nrow(stock_price) > 3) {
  stock_price %>%
    ggplot2::ggplot(aes(x = Year, y = adj_price)) +
    geom_point(aes(color = State), show.legend = FALSE) +
    geom_line(aes(color = State), show.legend = FALSE) +
    facet_wrap(vars(State)) +
    ggtitle(label = params$species, subtitle = "Data coverage for some states is poor with >5 years reported ") +
    xlab(label = "Year") +
    ylab(label = "Price per pound (lb/$) in 2019 value") +
    theme_minimal()
} else {
  ("Not enough data")
}
```

