## Age diversity

```{r data_wrangling_age}
selected.spp <- NEesp::common_names_survdat("bio") %>%
  dplyr::filter(common_name == params$species)
```

Diversity in age measurements of a stock is a useful indicator of several factors relating to fishing pressure and recruitment. A decrease in diversity can be due to either truncation, the lack of older or younger ages. Diversity changes as a function of an increase of a single/few ages relative to the usual stock age structure or as more ages become less represented. Diagnostic plots of age are constructed below using fisheries independent data from `survdat`.


### Age diversity 

`r if(params$species == "Monkfish"){"The age determination method for monkfish has not been validated, and the anatomic structure used has changed through time. In addition these stocks display a high degree of sexual dimorphisms making fitting of single growth curves unreliable.  This results a high degree of uncertainty in assessing the age structure of the stock and the effects of fishing pressure."}`


```{r, age-diversity-calculations, fig.cap = paste(species, "age diversity")}
NEesp::plot_age_diversity(selected.spp, species = species)
```


### Density plots of age 

Age distribution across years of survey data of `r params$species`. These plots can help identify strong year classes of recruits and how these classes persist in the fishery. 


```{r age-density, fig.height = 12, fig.cap = paste(species, "age density")}
NEesp::plot_age_density(selected.spp)
```
