## Distribution

Using fisheries independent data from bottom trawls in `survdat`. Several metrics of distribution in observed catches are calculated. Changes in distribution can indicate changes in the geographic range of a species due to a variety of constraints such as food availability or physiologic limitations. 

### Map of seasonal ranges

Strata maps were pulled and compiled using code from [NOAA/EDAB ECSA](https://github.com/NOAA-EDAB/ECSA).


```{r, ecsa-map, fig.cap = paste(species, "distribution")}
NEesp::map_strata(
  strata = latlong_data,
  common_name = species
)
```


### Density and distribution estimation

The density of observations for `r params$species` was visualized using two-dimensional kernel density estimation on a grid as documented in `MASS::kde2d`. This density estimation is then visualized to indicate areas of greater or lesser probability of occurrence.


```{r geo-conc-data-wrangling-a}
survdata.w.codes <- NEesp::bio_survey %>%
  dplyr::mutate(SVSPP = SVSPP %>% as.numeric()) %>%
  dplyr::left_join(NEesp::species_key, by = "SVSPP")

selected.surv <- survdata.w.codes %>%
  dplyr::filter(Species == params$species)
```


```{r, fig.cap = paste(species, "density and distribution by decade"), fig.height = 12}
if (nrow(selected.surv) > 3) {
  # remove low capture years from this vis to make

  selected.surv <- selected.surv %>%
    dplyr::filter(YEAR > 1970) %>%
    dplyr::mutate(decade = ggplot2::cut_width(YEAR, width = 10, center = 1980))

  world <- rnaturalearth::ne_countries(scale = "medium", 
                                       returnclass = "sf")
  latmin <- (min(selected.surv$LAT) - 1)
  latmax <- (max(selected.surv$LAT) + 1)
  lonmin <- (min(selected.surv$LON) - 1)
  lonmax <- (max(selected.surv$LON) + 1)
 
  ggplot2::ggplot(data = world) +
    ggplot2::geom_sf() +
    ggplot2::stat_density2d(ggplot2::aes(x = LON, 
                                y = LAT, 
                                fill = ..level..), 
                            alpha = .5, 
                            geom = "polygon", 
                            data = selected.surv) +
    nmfspalette::scale_fill_nmfs(palette = "crustacean", 
                                 discrete = FALSE, 
                                 reverse = TRUE) +
    ggplot2::coord_sf(xlim = c(lonmin - 1, lonmax + 1), 
                      ylim = c(latmin - 1, latmax + 1)) +
    ggplot2::scale_color_gradient(low = "blue", 
                         high = "red", 
                         name = "Year") +
    ggplot2::xlab("Longitude") +
    ggplot2::ylab("Latitude") +
    ggplot2::theme(legend.position = "none") +
    ggplot2::facet_wrap(~decade, ncol = 2)
  
} else {
  ("Not enough data")
}
```


### Centroid of observations 

The unweighted centroid is one metric that describes the geometric center of the observed range of `r params$species` in a given year. Calculated as the mean of latitude and longitude of all survey tows that captured the species of interest.


```{r, centroid-plot-a, fig.cap = paste(species, "centroid over time")}
if(nrow(selected.surv) > 3) {
  selected.surv.centroid <- selected.surv %>%
    dplyr::group_by(YEAR) %>%
    dplyr::summarise(lon = mean(LON), lat = mean(LAT))
} else {
  ("Not enough data")
}

if(nrow(selected.surv) > 3) {
  if(nrow(selected.surv.centroid) > 4) {
    world <- rnaturalearth::ne_countries(scale = "medium", 
                                         returnclass = "sf")
    latmin <- (min(selected.surv.centroid$lat) - 1)
    latmax <- (max(selected.surv.centroid$lat) + 1)
    lonmin <- (min(selected.surv.centroid$lon) - 1)
    lonmax <- (max(selected.surv.centroid$lon) + 1)
  
    ggplot2::ggplot(data = world) +
      ggplot2::geom_sf() +
      ggplot2::geom_point(data = selected.surv.centroid, 
                 ggplot2::aes(x = lon,
                     y = lat,
                     color = as.numeric(YEAR))) +
      ggrepel::geom_label_repel(data = selected.surv.centroid, 
                                ggplot2::aes(x = lon, y = lat, label = YEAR), 
                                max.overlaps = 20) +
      ggplot2::coord_sf(xlim = c(lonmin - 1, lonmax + 1), 
                        ylim = c(latmin - 1, latmax + 1)) +
      ggplot2::scale_color_gradient(low = "blue", 
                           high = "red", 
                           name = "Year") +
      ggplot2::xlab("Longitude") +
      ggplot2::ylab("Latitude")
  
    }
} else {
  ("Not enough data")
}
```


### Latitude and longitude ranges

Latitude and longitude ranges were calculated from NOAA/EDAB ECSA [seasonal species strata](https://github.com/NOAA-EDAB/ECSA/blob/master/data/seasonal_stock_strata.csv) and [Bottom Trawl Survey (BTS) shapefiles](https://github.com/NOAA-EDAB/ECSA/tree/master/data/strata_shapefiles). The coordinate system is WGS84.
```{r, latlong}
data <- NEesp::get_latlong(species, latlong_data, shapefile = NEesp::shape) %>%
  NEesp::character_to_factor()

NEesp::make_html_table(data, 
                       type = params$file,
                       col_names = colnames(data))
```
