## Geographic distribution

### Map of distribution
Strata maps were pulled and compiled using code from [NOAA/EDAB ECSA](https://github.com/NOAA-EDAB/ECSA).


```{r, ecsa-map, fig.cap = paste(species, "distribution")}
NEesp::map_strata(
  strata = latlong_data,
  common_name = species
)
```


### Latitude and longitude ranges
Latitude and longitude ranges were calculated from NOAA/EDAB ECSA [seasonal species strata](https://github.com/NOAA-EDAB/ECSA/blob/master/data/seasonal_stock_strata.csv) and [Bottom Trawl Survey (BTS) shapefiles](https://github.com/NOAA-EDAB/ECSA/tree/master/data/strata_shapefiles). The coordinate system is WGS84.
```{r, latlong}
data <- NEesp::get_latlong(species, latlong_data, shapefile = NEesp::shape) %>%
  NEesp::character_to_factor()

NEesp::make_html_table(data, col_names = colnames(data))
```


Using fisheries independent data from bottom trawls in `survdat`. Several metrics of distribution in observed catches are calculated. Changes in distribution can indicate changes in the geographic range of a species due to a variety of constraints such as food availability or physiologic limitations.The unweighted centroid is one metric that describes the geometric center of the observed range of `r params$species` in a given year. Calculated as the mean of latitude and longitude of all tows that captured the species of interest.The density of observations for `r params$species` was visualized using two-dimensional kernel density estimation on a grid as documented in `MASS::kde2d`. This density estimation is then visualized to indicate areas of greater or lesser probability of occurrence 


```{r}
# load all survdat data
survdata.w.codes <- NEesp::common_names_survdat("all")

selected.surv <- survdata.w.codes %>%
  dplyr::filter(common_name == params$species)

if (nrow(selected.surv) > 3) {
  selected.surv.centroid <- selected.surv %>%
    dplyr::group_by(YEAR) %>%
    dplyr::summarise(lon = mean(LON), lat = mean(LAT))
} else {
  ("Not enough data")
}
```

### Figures

#### Density and distribution estimation by season
```{r, density-plot-setup-spring}

selected.surv.temp <- selected.surv %>% dplyr::filter(YEAR < 1970, SEASON == "SPRING")

if (nrow(selected.surv) > 3) {
  if (nrow(selected.surv.temp) < 200) {
    # remove low capture years from this vis to make it consistently run

    selected.surv.temp <- selected.surv %>% dplyr::filter(YEAR > 1970)


    selected.surv.1970.plus <- selected.surv %>%
      dplyr::filter(YEAR > 1970, SEASON == "SPRING") %>%
      dplyr::mutate(decade = ggplot2::cut_width(YEAR, width = 10, center = 1980))


    latmin <- (min(selected.surv.centroid$lat) - 3)
    latmax <- (max(selected.surv.centroid$lat) + 3)
    lonmin <- (min(selected.surv.centroid$lon) - 3)
    lonmax <- (max(selected.surv.centroid$lon) + 3)


    ggplot2::ggplot(data = ecodata::coast) +
      ggplot2::geom_sf() +
      ggplot2::stat_density2d(ggplot2::aes(
        x = LON,
        y = LAT,
        fill = ..level..
      ),
      alpha = .5,
      geom = "polygon",
      data = selected.surv.1970.plus
      ) +
      nmfspalette::scale_fill_nmfs(
        palette = "seagrass",
        discrete = FALSE,
        reverse = TRUE
      ) +
      ggplot2::coord_sf(
        xlim = c(lonmin, lonmax),
        ylim = c(latmin, latmax)
      ) +
      ggplot2::scale_color_gradient(
        low = "blue",
        high = "red",
        name = "Year"
      ) +
      ggplot2::xlab("Longitude") +
      ggplot2::ylab("Latitude") +
      ggplot2::theme(legend.position = "none") +
      ggplot2::facet_wrap(~decade) +
      ggplot2::ggtitle("Spring")
  } else {

    # if there are  more than 200 observation pre 1970 then this visualization will capture that data
    selected.surv.1970 <- selected.surv %>%
      dplyr::filter(SEASON == "SPRING") %>%
      dplyr::mutate(decade = ggplot2::cut_width(YEAR, width = 10, center = 1980))


    latmin <- (min(selected.surv.centroid$lat) - 3)
    latmax <- (max(selected.surv.centroid$lat) + 3)
    lonmin <- (min(selected.surv.centroid$lon) - 3)
    lonmax <- (max(selected.surv.centroid$lon) + 3)


    ggplot2::ggplot(data = ecodata::coast) +
      ggplot2::geom_sf() +
      ggplot2::stat_density2d(ggplot2::aes(
        x = LON,
        y = LAT,
        fill = ..level..
      ),
      alpha = .5,
      geom = "polygon",
      data = selected.surv.1970
      ) +
      nmfspalette::scale_fill_nmfs(
        palette = "seagrass",
        discrete = FALSE,
        reverse = TRUE
      ) +
      ggplot2::coord_sf(
        xlim = c(lonmin, lonmax),
        ylim = c(latmin, latmax)
      ) +
      ggplot2::scale_color_gradient(
        low = "blue",
        high = "red",
        name = "Year"
      ) +
      ggplot2::xlab("Longitude") +
      ggplot2::ylab("Latitude") +
      ggplot2::theme(legend.position = "none") +
      ggplot2::facet_wrap(~decade)
  }
} else {
  ("No data")
}
```


```{r, density-plot-setup-fall}

selected.surv.temp.f <- selected.surv %>% dplyr::filter(YEAR < 1970, SEASON == "FALL")

if (nrow(selected.surv) > 3) {
  if (nrow(selected.surv.temp.f) < 200) {
    # remove low capture years from this vis to make it consistantly run

    selected.surv.temp <- selected.surv %>% dplyr::filter(YEAR > 1970)


    selected.surv.1970.plus <- selected.surv %>%
      dplyr::filter(YEAR > 1970, SEASON == "FALL") %>%
      dplyr::mutate(decade = ggplot2::cut_width(YEAR, width = 10, center = 1980))


    latmin <- (min(selected.surv.centroid$lat) - 3)
    latmax <- (max(selected.surv.centroid$lat) + 3)
    lonmin <- (min(selected.surv.centroid$lon) - 3)
    lonmax <- (max(selected.surv.centroid$lon) + 3)


    ggplot2::ggplot(data = ecodata::coast) +
      ggplot2::geom_sf() +
      ggplot2::stat_density2d(ggplot2::aes(x = LON, y = LAT, fill = ..level..), alpha = .5, geom = "polygon", data = selected.surv.1970.plus) +
      nmfspalette::scale_fill_nmfs(palette = "crustacean", discrete = FALSE, reverse = TRUE) +
      ggplot2::coord_sf(xlim = c(lonmin, lonmax), ylim = c(latmin, latmax)) +
      ggplot2::scale_color_gradient(low = "blue", high = "red", name = "Year") +
      ggplot2::xlab("Longitude") +
      ggplot2::ylab("Latitude") +
      ggplot2::theme(legend.position = "none") +
      ggplot2::facet_wrap(~decade) +
      ggplot2::ggtitle("Fall")
  } else {

    # if there are  more than 200 observation pre 1970 then this visualization will capture that data
    selected.surv.1970 <- selected.surv %>%
      dplyr::filter(SEASON == "FALL") %>%
      dplyr::mutate(decade = ggplot2::cut_width(YEAR, width = 10, center = 1980))


    latmin <- (min(selected.surv.centroid$lat) - 3)
    latmax <- (max(selected.surv.centroid$lat) + 3)
    lonmin <- (min(selected.surv.centroid$lon) - 3)
    lonmax <- (max(selected.surv.centroid$lon) + 3)


    ggplot2::ggplot(data = ecodata::coast) +
      ggplot2::geom_sf() +
      ggplot2::stat_density2d(ggplot2::aes(x = LON, y = LAT, fill = ..level..), alpha = .5, geom = "polygon", data = selected.surv.1970) +
      nmfspalette::scale_fill_nmfs(palette = "crustacean", discrete = FALSE, reverse = TRUE) +
      ggplot2::coord_sf(xlim = c(lonmin, lonmax), ylim = c(latmin, latmax)) +
      ggplot2::scale_color_gradient(low = "blue", high = "red", name = "Year") +
      ggplot2::xlab("Longitude") +
      ggplot2::ylab("Latitude") +
      ggplot2::theme(legend.position = "none") +
      ggplot2::facet_wrap(~decade) +
      ggplot2::ggtitle("Fall")
  }
} else {
  ("No data")
}
```


#### Centroid of observations 

The average position of observations across all years of surveys 

```{r, centroid-plot}

if (nrow(selected.surv.centroid) > 4) {
  latmin <- (min(selected.surv.centroid$lat) - 1)
  latmax <- (max(selected.surv.centroid$lat) + 1)
  lonmin <- (min(selected.surv.centroid$lon) - 1)
  lonmax <- (max(selected.surv.centroid$lon) + 1)


  ggplot2::ggplot(data = ecodata::coast) +
    ggplot2::geom_sf() +
    ggplot2::geom_point(data = selected.surv.centroid, ggplot2::aes(x = lon, y = lat, color = as.numeric(YEAR))) +
    ggrepel::geom_label_repel(data = selected.surv.centroid, ggplot2::aes(x = lon, y = lat, label = YEAR), max.overlaps = 20) +
    ggplot2::coord_sf(xlim = c(lonmin, lonmax), ylim = c(latmin, latmax)) +
    ggplot2::scale_color_gradient(low = "blue", high = "red", name = "Year") +
    ggplot2::xlab("Longitude") +
    ggplot2::ylab("Latitude")
} else {
  ("Not enough data")
}
```

