## Geographic distribution

Ricky Tabandera

```{r geography_setup, include=FALSE, eval = FALSE}
# avoid using `library()` in a package
knitr::opts_chunk$set(echo = TRUE)
suppressPackageStartupMessages(library(tidyverse))
library(dbutils)
library(survdat)
library(sf)
library("rnaturalearth")
```


Using fisheries independant data from bottom trawls in `survdat`. Several metrics of distribution in observed catches are calculated. Changes in distributuion can indicate changes in the geographic range of a species due to a variety of constraints such as food availability or physiologic limitations.The unweighted centroid is one metric that describes the geometric center of the observed range of `r params$species` in a given year. Calculated as the mean of latitude and longitude of all tows that captured the species of interest.The density of observations for `r params$species` was visualized using two-dimensional kernel density estimation on a grid as documented in `MASS::kde2d`. This density estimation is then visualized to indicate areas of greater or lesser probability of occurrence 


```{r geo-conc-data-wrangling }
# source(here::here("R","common_names_survdat.R")) # loaded in `life_history_functions.R`
```

```{r}
survdata.w.codes<-Common_names_survdat("bio")

```

```{r}
selected.surv<-survdata.w.codes %>%
   dplyr::filter(common_name == params$species)

# selected.surv<-survdata.w.codes %>% 
#    dplyr::filter(common_name == "Pollock")

```

```{r}

if(nrow(selected.surv)>3){
  
selected.surv.centroid<-selected.surv %>% dplyr::group_by(YEAR) %>% dplyr::summarise(lon=mean(LON),lat=mean(LAT))

}else("Not enough data")


```
### Figures


#### Density and distribution estimation by season
```{r, density-plot-setup-spring}
# underscores in chunk name mess up figure numbers (bookdown bug)

selected.surv.temp<-selected.surv %>%  dplyr::filter(YEAR <1970, SEASON=="SPRING") 

if(nrow(selected.surv)>3){
  
  if(nrow(selected.surv.temp)<200){
  #remove low capture years from this vis to make it consistantly run
  
selected.surv.temp<-selected.surv%>%  dplyr::filter(YEAR >1970)  
  
  
selected.surv.1970.plus<-selected.surv %>%
  dplyr::filter(YEAR >1970, SEASON=="SPRING")%>%
  dplyr::mutate(decade=ggplot2::cut_width(YEAR, width = 10, center =  1980 ))



latmin<-(min(selected.surv.centroid$lat)-3)
latmax<-(max(selected.surv.centroid$lat)+3)
lonmin<-(min(selected.surv.centroid$lon)-3)
lonmax<-(max(selected.surv.centroid$lon)+3)



ggplot2::ggplot(data = ecodata::coast) +
  ggplot2::geom_sf() +
  ggplot2::stat_density2d(ggplot2::aes(x=LON,
                                      y=LAT,
                                      fill = ..level..), 
                          alpha = .5,
                          geom = "polygon", 
                          data = selected.surv.1970.plus) +
  nmfspalette::scale_fill_nmfs(palette="seagrass",
                               discrete=FALSE, 
                               reverse=TRUE)+
  ggplot2::coord_sf(xlim=c(lonmin,lonmax), 
                    ylim=c(latmin,latmax))+
  ggplot2::scale_color_gradient(low = "blue", 
                               high = "red", 
                               name="Year")+
  ggplot2::xlab("Longitude")+
  ggplot2::ylab("Latitude")+
  ggplot2::theme(legend.position = 'none')+
  ggplot2::facet_wrap(~decade)+
  ggplot2::ggtitle("Spring")
  
}else{
  
#if there are  more than 200 observation pre 1970 then this visualization will capture that data  
selected.surv.1970<-selected.surv %>% filter(SEASON=="SPRING") %>%
  dplyr::mutate(decade=ggplot2::cut_width(YEAR, width = 10, center =  1980 ))



latmin<-(min(selected.surv.centroid$lat)-3)
latmax<-(max(selected.surv.centroid$lat)+3)
lonmin<-(min(selected.surv.centroid$lon)-3)
lonmax<-(max(selected.surv.centroid$lon)+3)



ggplot2::ggplot(data = ecodata::coast) +
  ggplot2::geom_sf() +
  ggplot2::stat_density2d(ggplot2::aes(x=LON,
                                      y=LAT,
                                      fill = ..level..), 
                          alpha = .5,
                          geom = "polygon", 
                          data = selected.surv.1970) +
  nmfspalette::scale_fill_nmfs(palette="seagrass",
                               discrete=FALSE, 
                               reverse=TRUE)+
  ggplot2::coord_sf(xlim=c(lonmin,lonmax), 
                    ylim=c(latmin,latmax))+
  ggplot2::scale_color_gradient(low = "blue", 
                               high = "red", 
                               name="Year")+
  ggplot2::xlab("Longitude")+
  ggplot2::ylab("Latitude")+
  ggplot2::theme(legend.position = 'none')+
  ggplot2::facet_wrap(~decade)
  
}
  
}else("No data")
  










```




```{r, density-plot-setup-fall}
# underscores in chunk name mess up figure numbers (bookdown bug)

selected.surv.temp.f<-selected.surv %>%  dplyr::filter(YEAR <1970, SEASON=="FALL") 

if(nrow(selected.surv)>3){
  
  if(nrow(selected.surv.temp.f)<200){
  #remove low capture years from this vis to make it consistantly run
  
selected.surv.temp<-selected.surv%>%  dplyr::filter(YEAR >1970)  
  
  
selected.surv.1970.plus<-selected.surv %>%
  dplyr::filter(YEAR >1970, SEASON=="FALL")%>%
  dplyr::mutate(decade=ggplot2::cut_width(YEAR, width = 10, center =  1980 ))



latmin<-(min(selected.surv.centroid$lat)-3)
latmax<-(max(selected.surv.centroid$lat)+3)
lonmin<-(min(selected.surv.centroid$lon)-3)
lonmax<-(max(selected.surv.centroid$lon)+3)



ggplot2::ggplot(data = ecodata::coast) +
  ggplot2::geom_sf() +
  ggplot2::stat_density2d(ggplot2::aes(x=LON,y=LAT,fill = ..level..), alpha = .5,geom = "polygon", data = selected.surv.1970.plus) +
  nmfspalette::scale_fill_nmfs(palette="crustacean",discrete=FALSE, reverse=TRUE)+
  ggplot2::coord_sf(xlim=c(lonmin,lonmax), ylim=c(latmin,latmax))+
  ggplot2::scale_color_gradient(low = "blue", high = "red", name="Year")+
  ggplot2::xlab("Longitude")+
  ggplot2::ylab("Latitude")+
  ggplot2::theme(legend.position = 'none')+
  ggplot2::facet_wrap(~decade)+
  ggplot2::ggtitle("Fall")

ggplot2::ggplot(data = ecodata::coast) +
  ggplot2::geom_sf() +
  ggplot2::stat_density2d(ggplot2::aes(x=LON,y=LAT,fill = ..level..), alpha = .5,geom = "polygon", data = selected.surv.1970.plus) +
  ggplot2::geom_sf(data = ecodata::coast) +
  nmfspalette::scale_fill_nmfs(palette="crustacean",discrete=FALSE, reverse=TRUE)+
  ggplot2::coord_sf(xlim=c(lonmin,lonmax), ylim=c(latmin,latmax))+
  ggplot2::scale_color_gradient(low = "blue", high = "red", name="Year")+
  ggplot2::xlab("Longitude")+
  ggplot2::ylab("Latitude")+
  ggplot2::theme(legend.position = 'none')+
  ggplot2::facet_wrap(~decade)+
  ggplot2::ggtitle("Fall")
  
}else{
  
#if there are  more than 200 observation pre 1970 then this visualization will capture that data  
selected.surv.1970<-selected.surv %>% filter(SEASON=="FALL") %>%
  dplyr::mutate(decade=ggplot2::cut_width(YEAR, width = 10, center =  1980 ))



latmin<-(min(selected.surv.centroid$lat)-3)
latmax<-(max(selected.surv.centroid$lat)+3)
lonmin<-(min(selected.surv.centroid$lon)-3)
lonmax<-(max(selected.surv.centroid$lon)+3)



ggplot2::ggplot(data = ecodata::coast) +
  ggplot2::geom_sf() +
  ggplot2::stat_density2d(ggplot2::aes(x=LON,y=LAT,fill = ..level..), alpha = .5,geom = "polygon", data = selected.surv.1970) +
  nmfspalette::scale_fill_nmfs(palette="crustacean",discrete=FALSE, reverse=TRUE)+
  ggplot2::coord_sf(xlim=c(lonmin,lonmax), ylim=c(latmin,latmax))+
  ggplot2::scale_color_gradient(low = "blue", high = "red", name="Year")+
  ggplot2::xlab("Longitude")+
  ggplot2::ylab("Latitude")+
  ggplot2::theme(legend.position = 'none')+
  ggplot2::facet_wrap(~decade)+
  ggplot2::ggtitle("Fall")
  
}
  
}else("No data")
  








```


#### Centroid of observations 

The average position of observations across all years of surveys 

```{r, centroid-plot}
# spaces in chunk name mess up figure numbers (bookdown bug)

if(nrow(selected.surv.centroid)<4){
  
latmin<-(min(selected.surv.centroid$lat)-1)
latmax<-(max(selected.surv.centroid$lat)+1)
lonmin<-(min(selected.surv.centroid$lon)-1)
lonmax<-(max(selected.surv.centroid$lon)+1)


ggplot2::ggplot(data = ecodata::coast ) +
  geom_sf() +
  geom_point(data = selected.surv.centroid, ggplot2::aes(x=lon,y=lat,color=as.numeric(YEAR)))+
  ggrepel::geom_label_repel(data = selected.surv.centroid,aes(x=lon,y=lat,label=YEAR),max.overlaps =20 )+
  ggplot2::coord_sf(xlim=c(lonmin,lonmax), ylim=c(latmin,latmax))+
  ggplot2::scale_color_gradient(low = "blue", high = "red", name="Year")+
  ggplot2::xlab("Longitude")+
  ggplot2::ylab("Latitude")
}else("Not enough data")

```




