## Length

Length data were pulled from ``r params$survey_source``. Only years with more than 10 fish lengths were considered for analysis. 

### Overview 


```{r, length-freq, fig.cap = paste(species, "length frequency distribution")}
NEesp::plot_len_hist(survey_data)
```


### Summary statistics 
Separate `geom_gls()` functions were fit for the minimum, mean, and maximum lengths; trend lines are only shown when the trend was statistically significant, so some plots may have fewer than three trend lines. Please note, sometimes the survey observed a small number of fish outside of the defined stock area.


```{r, length, fig.cap = paste(species, "length")}
NEesp::generate_len_plot(survey_data)
```


### Risk 

See Methods for risk calculation details.
```{r}
indicators <- c(
  "max_length_spring",
  "max_length_fall",
  "avg_length_spring",
  "avg_length_fall"
)
```

```{r, child = system.file("indicator_bookdown_template/_risk-child-doc.Rmd", package = "NEesp")}
```

### Summary
```{r, length_summary}
NEesp::generate_len_table(survey_data)
```

### Data
```{r, length_data}
data <- NEesp::get_len_data2(survey_data) %>%
  NEesp::character_to_factor()

NEesp::make_html_table(data,
  type = params$file,
  col_names = c(
    "Year", "Season", "Region", "Number of fish",
    "Mean length", "Minimum length", "Maximum length"
  )
)
```
