## Length at first maturity

### Size at first maturity 

L50 or length at which 50%  maturity can be calculated using differing methods. Using Data sourced from `survdat` that consist of visual sexual maturity determination and length measurements. Here, L50 is estimated by fitting a logistic generalized linear model to gonadal maturity and body length measurements and estimating the inflection point of the logistic model. This point represents the size where the individual has 50% odds of being mature. Decade variations of this parameter are investigated utilizing decade as a additive and interaction term in the model specification.

The relationship between sexual maturity by body size is a well established life history parameter. L50 is a useful metric in fisheries to identify sizes that are able to reproduce. This information can be used to inform regulation as to minimum catch size that should allow for a significant portion of the population to spawn and contribute genetic information to sustain the stock. Changes in this parameter can signal population/evolutionary pressures on the stock with reduction in this size potentially indicating excessive fishing pressure.

```{r, L50-data-wrangling}

# only take records with a maturity status associated
selected.surv.clean <- NEesp::bio_survey %>%
  dplyr::mutate(SVSPP = as.numeric(SVSPP)) %>%
  dplyr::left_join(NEesp::species_key, by = "SVSPP") %>%
  dplyr::filter(Species == species) %>%
  tidyr::drop_na(MATURITY) %>%
  dplyr::mutate(MATURITY = MATURITY %>% stringr::str_to_upper()) %>%
  dplyr::filter(MATURITY %in% c("D", "I", "R", "S", "T", "U", "E"))

# recode the maturity col into factor

# recode sexed fish to readable names
# sex codes 1 male, 2 female, 4 transitional, 0 unsexed/unknown
selected.surv.clean$MATURITY <- dplyr::recode(selected.surv.clean$MATURITY, "D" = "Mature", "I" = "Immature", "R" = "Mature", "S" = "Mature", "T" = "Mature", "U" = "Mature", "E" = "Mature")
# recode sex to male or female
selected.surv.clean$SEX <- dplyr::recode(selected.surv.clean$SEX, "1" = "male", "2" = "female", "0" = "unknown", "4" = "transitional")

selected.surv.clean <- selected.surv.clean %>%
  dplyr::mutate(
    SEX = as.factor(selected.surv.clean$SEX),
    MATURITY = as.factor(MATURITY)
  )

selected.surv.f <- selected.surv.clean %>%
  dplyr::filter(SEX %in% c("female", "male") & MATURITY != "X")
```

```{r, mature_model}
# conditionally run models based on having more than 3 rows
eval_model <- nrow(selected.surv.f) > 3 &
  selected.surv.f$MATURITY %>%
    unique() %>%
    length() > 1 # more than 1 maturity category

if (eval_model) {

  # X= unknown , I = immature, D= developing, R S T U convert to developed

  # #bin the years of fish into decades to compare if model parameters are chaging
  # selected.surv.f<-selected.surv.f %>% mutate(decade=ggplot2::cut_width(YEAR, width = 10, boundary = 1980))
  # levels(selected.surv.f$decade)<-c( "1980","1990", "2000","2010")

  # model with sex as additive and multiplicative term
  if (selected.surv.f$SEX %>%
    unique() %>%
    length() == 2) {
    maturity.mod <- glm(MATURITY ~ LENGTH + SEX + LENGTH * SEX, data = selected.surv.f, family = "binomial")
  }

  # handle cases with only 1 sex
  if (selected.surv.f$SEX %>%
    unique() %>%
    length() == 1) {
    maturity.mod <- glm(MATURITY ~ LENGTH, data = selected.surv.f, family = "binomial")
  }

  pretty_mod <- papeR::prettify(summary(maturity.mod))

  # summary(maturity.mod)
} else {
  ("Not enough data to fit model")
}
```


### Maturity classification 

The gonadal development stage can vary between immature and mature. Once mature, there are several stages that represent phases in the spawning sequence. These stages and phases can vary across body length and the proportion of the population in each of these categories can help identify spawning size and what seasonal effects there are on development. 

```{r, maturity-classification}
#########################
# filter for maturity codes
selected_surv_mat_clean <- NEesp::bio_survey %>%
  dplyr::mutate(SVSPP = as.numeric(SVSPP)) %>%
  dplyr::left_join(NEesp::species_key, by = "SVSPP") %>%
  dplyr::filter(Species == species) %>%
  tidyr::drop_na(MATURITY) %>%
  dplyr::mutate(MATURITY = MATURITY %>% stringr::str_to_upper())

# recode maturity codes to readable names
selected_surv_mat_clean$MATURITY <- dplyr::recode(selected_surv_mat_clean$MATURITY, "D" = "DEVELOPING", "I" = "IMMATURE", "R" = "RIPE", "S" = "SPENT", "T" = "RESTING", "U" = "RIPE AND RUNNING", "E" = "EYED", "X" = "UNKNOWN", "i" = "IMMATURE", "r" = "RIPE", "d" = "DEVELOPING", "t" = "RESTING", "1" = "UNKNOWN")
```


```{r, maturity-classification-plot, fig.cap = paste(species, "maturity classification")}

if (nrow(selected_surv_mat_clean > 5)) {
  mat_sum_season <- selected_surv_mat_clean %>%
    dplyr::group_by(LENGTH, MATURITY, SEASON) %>%
    dplyr::summarise(mat.code.count = length(MATURITY)) # problems with dplyr::n()

  dat <- mat_sum_season %>%
    dplyr::group_by(LENGTH, SEASON) %>%
    dplyr::mutate(
      mat.sum = sum(mat.code.count),
      prop = (mat.code.count / mat.sum)
    ) %>%
    dplyr::group_by(LENGTH)

  if (nrow(dat) > 0) {
    fig <- dat %>%
      ggplot2::ggplot(ggplot2::aes(x = LENGTH, y = prop * 100, fill = MATURITY)) +
      ggplot2::geom_bar(position = "stack", stat = "identity") +
      nmfspalette::scale_fill_nmfs(palette = "regional web", discrete = TRUE, reverse = FALSE) +
      ggplot2::xlab("Length (cm)") +
      ggplot2::ylab("Percent (%)") +
      ggplot2::ggtitle("Maturity classification across body length of", subtitle = species)

    if (dat$SEASON %>% unique() %>% length() > 1) {
      fig <- fig +
        ggplot2::facet_wrap(ggplot2::vars(SEASON))
    }

    return(fig)
  } else {
    ("Not enough data")
  }
}
```


### Model results

`r if(nrow(selected.surv.f)>3){"If the overall model is significant, there is support for the relationship between sexual maturity and body size. This model can predict the size where there are 50% odds of the indevidual is sexually mature. The results table below displays the L50 for Male and female indeviduals. "}` 

```{r, model-table}
if (eval_model) {
  knitr::kable(pretty_mod, digits = 3)
}
```

### L50 differences between sexes

`r if(nrow(selected.surv.f)>3){"The significance of differences between sex can be determined by referencing the model summary table above differences in intercept indicate differences in mean size between groups. Differences in betta coefficents indicate increases or decreases to the degree to which probability of maturity increases across body lengths   "}`

```{r model-results-and-L50}
if (eval_model) {
  if (summary(maturity.mod)$coefficients[, 4][1] < 0.05) {

    # intercept devided by beta
    L50.f <- (-coef(maturity.mod)[1]) / (coef(maturity.mod)[2]) # female
    L50.m <- (-(coef(maturity.mod)[1] + coef(maturity.mod)[3])) / (coef(maturity.mod)[2] + coef(maturity.mod)[4]) # male

    L50 <- data.frame(L50.f, L50.m)
    row.names(L50) <- c("Length (cm) at 50% maturity")
    colnames(L50) <- c("Female", "Male")
    L50.p <- papeR::prettify(L50)
    knitr::kable(L50.p, digits = 3)
  }
} else {
  ("Non-significant")
}
```


```{r, probability-plot, fig.cap = paste(species, "probability of maturity")}

if (eval_model) {
  if (summary(maturity.mod)$coefficients[, 4][1] < 0.05) {
    selected.surv.LENGTH <- seq(1:max(selected.surv.f$LENGTH))

    sel.length <- length(selected.surv.LENGTH)
    sel.min <- min(selected.surv.f$LENGTH)
    sel.max <- max(selected.surv.f$LENGTH)

    L50.sim.f <- boot::inv.logit(predict.glm(maturity.mod,
      newdata = data.frame(
        LENGTH = seq(1:sel.max),
        SEX = rep("female", length(seq(1:sel.max))),
        type = "response"
      )
    ))

    L50.sim.m <- boot::inv.logit(predict.glm(maturity.mod,
      newdata = data.frame(
        LENGTH = seq(1:sel.max),
        SEX = rep("male", length(seq(1:sel.max))),
        type = "response"
      )
    ))


    selected.surv.simlulation <- data.frame(L50.sim.f, L50.sim.m, selected.surv.LENGTH)
    selected.surv.simlulation <- selected.surv.simlulation %>%
      tidyr::pivot_longer(!selected.surv.LENGTH, names_to = "sex", values_to = "pred")


    selected.surv.simlulation %>%
      ggplot2::ggplot(ggplot2::aes(x = selected.surv.LENGTH, y = pred, color = sex)) +
      ggplot2::geom_line(size = 3) +
      ggplot2::xlab("Length (cm)") +
      ggplot2::ylab("Probability") +
      ggplot2::annotate("text", label = c("Mature", "Immature"), x = 30, y = c(1.05, -.05)) +
      ggplot2::geom_hline(yintercept = c(1, 0)) +
      ggplot2::geom_vline(
        xintercept = c(L50.f, L50.m),
        colour = c("#0093D0", "#FF4438"),
        size = 1.4,
        show.legend = TRUE
      ) +
      ggplot2::scale_color_manual(
        labels = c("Female", "Male"),
        name = "Sex",
        values = c("#0093D0", "#FF4438")
      )
  }
} else {
  ("Non-significant")
}
```
