---
title: "Preliminary ESP reports"
author: "Abigail Tyrell"
date: "`r format(Sys.time(), '%d %b %Y')`"

output: 
  html_document:
    toc: TRUE
    number_sections: TRUE
    toc_float: 
      collapsed: FALSE
    toc_depth: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

`%>%` <- dplyr::`%>%`

report_pages <- list.files(here::here("docs/reports"), pattern = ".html")

rec_pages <- list.files(here::here("docs/Recreational landings"), pattern = ".html")

rec_summary_pages <- rec_pages[stringr::str_which(rec_pages, pattern = "recreational")]

rec_species_pages <- rec_pages[-stringr::str_which(rec_pages, pattern = "recreational")]

com_pages <- list.files(here::here("docs/Commercial landings"), pattern = ".html")

```

# Preliminary reports
These reports are a preliminary analysis of ecological and socioeconomic data sourced from `assessmentdata`, `survdat`, and other public data sources.

```{r}
purrr::map(report_pages, 
           ~paste("<a href='https://rtabandera.github.io/esp_data_aggregation/reports/", 
                  .x, "'>",
                  stringr::str_split_fixed(.x, pattern = "_", n = 2)[,1], 
                  "</a>", sep = "")) %>%
  unlist() %>%
  knitr::kable(col.names = "Species name")

```

# Recreational data
These reports are a preliminary analysis of recreational data only.

```{r}
purrr::map(rec_summary_pages, 
           ~paste("<a href='https://rtabandera.github.io/esp_data_aggregation/Recreational landings/", 
                  .x, "'>",
                  stringr::str_split_fixed(.x, pattern = ".h", n = 2)[,1] %>% 
                    stringr::str_to_sentence(), 
                  "</a>", sep = "")) %>%
  unlist() %>%
  knitr::kable(col.names = "File name")

purrr::map(rec_species_pages, 
           ~paste("<a href='https://rtabandera.github.io/esp_data_aggregation/Recreational landings/", 
                  .x, "'>",
                  stringr::str_split_fixed(.x, pattern = "_", n = 2)[,1] %>% 
                    stringr::str_to_sentence(), 
                  "</a>", sep = "")) %>%
  unlist() %>%
  knitr::kable(col.names = "Species name")

```

# Commercial data
These reports are a preliminary analysis of commercial data only.

```{r}
purrr::map(com_pages, 
           ~paste("<a href='https://rtabandera.github.io/esp_data_aggregation/Commercial landings/", 
                  .x, "'>",
                  stringr::str_split_fixed(.x, pattern = "_", n = 2)[,1] %>% 
                    stringr::str_to_sentence(), 
                  "</a>", sep = "")) %>%
  unlist() %>%
  knitr::kable(col.names = "Species name")

```

# Risk analysis
This is a preliminary risk analysis visualization.

```{r}
paste("<a href='https://rtabandera.github.io/esp_data_aggregation/risk_ranking/", 
                  "plot_all_risk", "'>",
                  "Preliminary risk analysis", 
                  "</a>", sep = "") %>%
  knitr::kable(col.names = "File name")

```