---
title: "Black_sea_bass_foodhabits"
author: "Ricky Tabandera"
date: "5/12/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(gridExtra)
library(grid)
```

## Black sea bass food habits

Black sea bass are diurnal visual predators usually associated with complex hard structured bottoms and are observed to eat a wide variety of prey. Investigations into the food habits of black sea bass as sampled by three bottom trawl surveys (NEFSC, NEAMAP, and CHESAMAP).


```{r data_organization}

#all food habits from condition work +

load(url("https://github.com/Laurels1/Condition/raw/master/data/allfh.RData"))
#saveRDS(allfh, file="allfh_20200526.rds")
#key of strata from ECSA <- read.csv("https://raw.githubusercontent.com/NOAA-EDAB/ECSA/master/data/seasonal_stock_strata.csv")


CHESAMAP_2022<-read.csv("BSB_CHESAMAP_2022.csv")
NEAMAP_2022<-read.csv("BSB_NEAMAP_2022.csv" )

#col definitions from data dict
#PDLEN	PREDATOR LENGTH	Predator length in CM, rounded up.


#filter all food habits down to just black sea bass svspp code 141
#NEesp::species_key used to get svspp key

#Well-digested prey (unidentifiable animal remains) labeled as WDP.

#removing pre 1991 years as they lack a unique predator id making reconstructing stomachs difficult 
#removing rows without prey weight ~ 400 records removed 
# & year>1990
BSB_allfh<- allfh %>% filter(svspp=="141"  & !is.na(pyamtw))


#select only used variables PYPERI
BSB_small<-BSB_allfh %>% select(c(pynam, gencat, gensci, analcat,  analsci, 
                                  pyamtw,pywgti,pdid, pdlen,pdsex, pdwgt,pynum, 
                                  geoarea,
                                  stratum, surftemp, bottemp, season,month, day, year,decade ,cruise6))
#remove SAB as contains only ~20 records
BSB_small<-BSB_small %>% filter(!geoarea=="SAB" )


# north vs south
# geo regions in the dataset "MAB" "SNE" "GoM" "GB"  "SAB"
# refactor GOM and GB into SNE as northern subunit, and removed South atlantic bight

BSB_small$subunit <- BSB_small$geoarea

BSB_small$subunit[BSB_small$geoarea=="MAB"]<-"Southern"
BSB_small$subunit[BSB_small$geoarea=="SNE"]<-"Northern"
BSB_small$subunit[BSB_small$geoarea=="GoM"]<-"Northern"
BSB_small$subunit[BSB_small$geoarea=="GB"]<-"Northern"



```

## Frequency of occurance

NEFSC bottom trawl survey food habits contain a fraction of fish collected in a tow for stomach removal and analysis. Each entry is the identified contents of indevidual black sea bass stomachs associated with a unique predatorID. This dataset contains `r length(unique(BSB_ID$pdid))  ` stomachs containing `r length(BSB_ID$pdid) ` prey records. Data from prior to 1991 are 

```{r frequency_occ, echo=FALSE}

# PYSPP	Prey species name.

#removing 


 


N<-BSB_small %>% filter(subunit =="Northern" & !gensci%in%c("EMPTY","","OTHER") ) %>% group_by(gensci) %>%
  summarise(count=n(),freq_occ=(count/length(BSB_small$gensci)),weight=sum(pyamtw))


N_weight<-BSB_small %>% filter(subunit =="Northern"& !gensci%in%c("EMPTY","","OTHER")) %>% group_by(gensci) %>% summarise(weight=sum(pyamtw))
N_weight<- N_weight %>% mutate(perc_weight=weight/sum(weight)*100)  

N_order<-N %>%  mutate(gensci=forcats::fct_reorder(gensci, (count)))


#N_feq_occ_plot<-
N1<-N %>%
  mutate(gensci=forcats::fct_reorder(gensci, (count))) %>%
  ggplot(aes(x=gensci,y=count))+
  geom_bar(stat = "identity")+
  theme(axis.text.x = element_text(angle = 90),
        panel.border= element_blank(),
        panel.background= element_rect(fill = "white"),
        axis.title.y=element_blank()
        )+
   coord_flip()+
  xlab("Prey category")+
  ylab("Frequency of occurrence")

N2<-N_weight %>%
  mutate(gensci=forcats::fct_reorder(gensci, desc(perc_weight))) %>%
  ggplot(aes(x=N_order$gensci,y=weight))+geom_bar(stat = "identity")+
  theme(axis.text.x = element_text(angle = 90),
        panel.border= element_blank(),
        panel.background= element_rect(fill = "white"),
        axis.title.y=element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        )+
  scale_y_reverse()+
  coord_flip()+
  ylab("Weight (g)")


grid.arrange(N2,N1,ncol=2)





```

```{r By_weight, echo=FALSE, message=FALSE, warning=FALSE}

S<-BSB_small %>% filter(subunit =="Southern"& !gensci%in%c("EMPTY","","OTHER")) %>% group_by(gensci) %>%
  summarise(count=n(),freq_occ=(count/length(BSB_small$gensci)),weight=sum(pyamtw))

   

S_weight<-BSB_small %>% filter(subunit =="Southern"& !gensci==c("EMPTY","")) %>% group_by(gensci) %>% summarise(weight=sum(pyamtw))
S_weight<- S_weight %>% mutate(perc_weight=weight/sum(weight)*100)  


S_feq_occ_plot<-S %>% mutate(gensci=forcats::fct_reorder(gensci, desc(count))) %>%
  ggplot(aes(x=gensci,y=count))+geom_bar(stat = "identity")+
  theme(axis.text.x = element_text(angle = 90))+
  xlab("Prey category")+
  ylab("Frequency of occurrence")+
  theme_minimal()



S_weight_plot<-S_weight %>% mutate(gensci=forcats::fct_reorder(gensci, desc(perc_weight))) %>%
  ggplot(aes(x=gensci,y=perc_weight))+geom_bar(stat = "identity")+
  theme(axis.text.x = element_text(angle = 90))

print(N_weight_plot)
print(S_weight_plot)

```

