---
title: "Black_sea_bass_foodhabits"
author: "Ricky Tabandera"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(gridExtra)
library(grid)
```

## Black sea bass food habits

Black sea bass are diurnal visual predators usually associated with complex hard structured bottoms and are observed to eat a wide variety of prey. Investigations into the food habits of black sea bass as sampled by three bottom trawl surveys (NEFSC, NEAMAP, and CHESAMAP).


```{r data_organization, warning=FALSE, include=FALSE}

#all food habits from condition work +

#load(url("https://github.com/Laurels1/Condition/raw/master/data/allfh.RData"))
# load("C:\\Users\\ricky.tabandera\\Documents\\EDAB\\esp_data_aggregation\\black-sea-bass\\allfh_20200526.rds")
# saveRDS(allfh, file="allfh_20200526.RDS")
load("~/EDAB/esp_data_aggregation/black-sea-bass/allfh_20200526.RDATA")
#key of strata from ECSA <- read.csv("https://raw.githubusercontent.com/NOAA-EDAB/ECSA/master/data/seasonal_stock_strata.csv")


CHESAMAP_2022<-read.csv("~/EDAB/esp_data_aggregation/black-sea-bass/BSB_CHESAMAP_2022.csv")
NEAMAP_2022<-read.csv("~/EDAB/esp_data_aggregation/black-sea-bass/BSB_NEAMAP_2022.csv" )

#col definitions from data dict
#PDLEN	PREDATOR LENGTH	Predator length in CM, rounded up.


#filter all food habits down to just black sea bass svspp code 141
#NEesp::species_key used to get svspp key

#Well-digested prey (unidentifiable animal remains) labeled as WDP.

#removing pre 1991 years as they lack a unique predator id making reconstructing stomachs difficult 
#removing rows without prey weight ~ 400 records removed 
# & year>1990
BSB_allfh<- allfh %>% filter(svspp=="141"  )


#select only used variables PYPERI
BSB_small<-BSB_allfh %>% select(c(pynam, gencat, gensci, analcat,  analsci, 
                                  pyamtw,pywgti,pdid, pdlen,pdsex, pdwgt,pynum, 
                                  geoarea,station,stratum, surftemp, bottemp,
                                  season,month, day, year,decade ,cruise6,tow))

#remove SAB as contains only ~20 records
BSB_small<-BSB_small %>% filter(!geoarea=="SAB")






# north vs south
# geo regions in the dataset "MAB" "SNE" "GoM" "GB"  "SAB"
# refactor GOM and GB into SNE as northern subunit, and removed South atlantic bight

BSB_small$subunit <- BSB_small$geoarea

BSB_small$subunit[BSB_small$geoarea=="MAB"]<-"Southern"
BSB_small$subunit[BSB_small$geoarea=="SNE"]<-"Northern"
BSB_small$subunit[BSB_small$geoarea=="GoM"]<-"Northern"
BSB_small$subunit[BSB_small$geoarea=="GB"]<-"Northern"

#removing pre 1991 years as they lack a unique predator id making reconstructing stomachs difficult used for freq
BSB_small_ID<-BSB_small  %>% filter(year>1990)



```

```{r data_sumary, echo=FALSE, message=FALSE, warning=FALSE}

list.files(here::here("black-sea-bass","FWdp_shiney_data"))

prey_70<-read.csv(here::here("black-sea-bass","FWdp_shiney_data","DC.Black sea bass (Centropristis striata).1970s.2023-04-09.csv" )) %>%    mutate(Decade = 1970)   

prey_80<-read.csv(here::here("black-sea-bass","FWdp_shiney_data","DC.Black sea bass (Centropristis striata).1980s.2023-04-09.csv" )) %>%    mutate(Decade = 1980)
 
prey_80<-read.csv(here::here("black-sea-bass","FWdp_shiney_data","DC.Black sea bass (Centropristis striata).1980s.2023-04-09.csv" )) %>%    mutate(Decade = 1980)

prey_00<-read.csv(here::here("black-sea-bass","FWdp_shiney_data","DC.Black sea bass (Centropristis striata).2000s.2023-04-09.csv"   )) %>%    mutate(Decade = 2000)

prey_10<-read.csv(here::here("black-sea-bass","FWdp_shiney_data","DC.Black sea bass (Centropristis striata).2010s.2023-04-09.csv"  )) %>%    mutate(Decade = 2010)             
   
prey_20<-read.csv(here::here("black-sea-bass","FWdp_shiney_data","DC.Black sea bass (Centropristis striata).2020s.2023-04-09.csv"  )) %>%    mutate(Decade = 2020)               
  
BSB_prey_dec<-rbind(prey_70,prey_80,prey_00,prey_10,prey_20)


library(Polychrome)

prey_colors<-glasbey.colors(33)

py_names<-c("Ammodytes sp"   ,       "Anthozoa"     ,         "Aphroditidae"       ,   "Bivalvia",     
 "Cancer sp"     ,        "Centropristis striata", "Cephalopoda"         ,  "Clupeidae"   ,         
"Crustacea"       ,      "Crustacean shrimp",     "Decapoda crab"         ,"Decapoda shrimp",      
 "Euphausiacea"    ,      "Gammaridea"       ,     "Gastropoda"           , "Illex sp"       ,      
"Invertebrata"      ,    "Larvaceans"         ,   "Loligo pealeii"      ,  "Loligo sp"        ,    
"Miscellaneous"      ,   "Mollusca"            ,  "Myctophidae"         ,  "Mysidacea"         ,   
"Octopoda"            ,  "Ophiuroidea"          , "Paguroidea"           , "Pandalidae"         ,  
 "Polychaeta"          ,  "Prionotus carolinus"  , "Stenotomus chrysops"  , "Unidentified fish"  ,  
 "WDP")



names(prey_colors)<-py_names[1:length(prey_colors)]

prey_all<-read.csv(here::here("black-sea-bass","FWdp_shiney_data","DC.Black sea bass (Centropristis striata).All.2023-04-09.csv"   )) 


prey_m<-read.csv(here::here("black-sea-bass","FWdp_shiney_data","DC.Black sea bass (Centropristis striata).M.2023-04-09.csv"  ))                 


prey_s<-read.csv(here::here("black-sea-bass","FWdp_shiney_data","DC.Black sea bass (Centropristis striata).S.2023-04-09.csv"  ))    

prey_MAB<-read.csv(here::here("black-sea-bass","FWdp_shiney_data","DC.Black sea bass (Centropristis striata).MID-ATLANTIC BIGHT.2023-04-09.csv"  )) 

prey_SNE<-read.csv(here::here("black-sea-bass","FWdp_shiney_data","DC.Black sea bass (Centropristis striata).SOUTHERN NEW ENGLAND.2023-04-09.csv"  )) 


BSB_prey_dec %>%
  # mutate(Pct.m=forcats::fct_reorder2(Prey, Decade, Pct.m, .desc = TRUE))%>%
  ggplot(aes(x=as.factor(Decade), y=Pct.m ,fill=Prey))+
  geom_col(stat = "identity",show.legend = TRUE)+
  scale_fill_manual(values=prey_colors)+
  theme(axis.text.y=element_blank())+
  xlab("Decade")+
  ylab("Percent by weight")




BSB_prey_dec %>% filter(Decade==1970) %>% mutate(Pct.m=forcats::fct_reorder(Prey, Pct.m, .desc = TRUE)) %>%
  ggplot(aes(x=Pct.m,y=Prey, fill= Prey))+
  geom_bar(stat = "identity")+
  theme(axis.text.x = element_text(angle = 90),
        panel.border= element_blank(),
        panel.background= element_rect(fill = "white"),
        axis.title.y=element_blank(),
        legend.position="none"
        )+
   coord_flip()+
  xlab("Prey name")+
  ylab("Percent by weight")+
  scale_fill_manual(values=prey_colors)



```




## Frequency of occurance and weight of prey items

NEFSC bottom trawl survey food habits contain a fraction of fish collected in a tow for stomach removal and analysis. Each entry is the identified contents of indevidual black sea bass stomachs associated with a unique predatorID. This dataset contains `r length(unique(BSB_small$pdid))  ` stomachs containing `r length(BSB_small$pdid) ` prey records. Data coverage begins in 1977 though 2020.


```{r}
#create a new index of indevidual bsb as pdid appears to not be reliable at IDing fish early in the dataset
#strata cruise tow station pdid ==== to make key
BSB_small_ID<- BSB_small %>% unite("ID",c(month, day, year,stratum,cruise6, tow, station, pdid),remove =FALSE)

#used to get total number of stomachs by year
ID_counts<-BSB_small_ID %>% group_by(year) %>% summarise(id_count=length(unique(ID)))
#used to get counts of empty stomachs
ID_counts.1<-BSB_small_ID %>% group_by(year) %>% filter(analsci=="EMPTY") %>% summarise(empty.stom=length(unique(ID)))

#used to get proportion of stomachs that are empty
empty.count.table<-left_join(ID_counts.1, ID_counts, by=c("year"))
empty.count.table<-empty.count.table %>% mutate(pct.empty=(empty.stom/id_count))
# remove the blown stomachs as they are indetermanant 
BSB_small_ID_clean<-BSB_small_ID %>% filter(!analsci=="BLOWN")




#generate counts of each prey catagory by year and subunit
prey_dec_sub<-BSB_small_ID_clean %>% group_by(pynam ) %>% summarise(count=n()) %>% arrange(desc(count))

prey_dec_sub.ani<-BSB_small_ID_clean %>% group_by(analsci ) %>% summarise(count=n()) %>% arrange(desc(count))


#get a count of blown stomachs and empty by decade and subunit
prey_dec_sub_blown<-BSB_small_ID_clean %>% group_by(decade,subunit ) %>% summarise(count=n())
BSB_small_ID %>% group_by(decade,subunit) %>% filter(analsci=="BLOWN") %>% summarise(count=n())



# weighted average by prey item

# tow id for 
BSB_small_ID<-BSB_small_ID %>% unite("tow_ID",c(cruise6, tow,station, stratum),remove =FALSE)
#verify we have the correct number of tows for BSB
tow_ID_list<-BSB_small_ID %>%  summarise( tow_count=unique(tow_ID))
# count the number of predator stomachs in each tow
n_pred_per_tow<-BSB_small_ID %>% group_by(tow_ID) %>% summarise(n_stom=length(unique(ID)))
# calculate the total number of stomachs by decade 
n_pred_per_tow_dec<-BSB_small_ID %>% group_by(decade) %>% summarise(n_stom_dec=length(unique(ID)))
BSB_small_ID %>% group_by(decade) %>% summarise(n_tow_dec=length(unique(ID)))

# merge total number of stomachs by decade into original dataset 
BSB_small_ID<-left_join(BSB_small_ID,n_pred_per_tow, by="tow_ID")

n_tow_dec_table<-BSB_small_ID %>% group_by(decade) %>% summarise(n_tow_dec=length(unique(tow_ID)))
  
BSB_small_ID<-left_join(BSB_small_ID,n_tow_dec_table,by="decade")
# calculate weighted average by decade 

t1<-BSB_small_ID %>% filter() %>% group_by( decade,pynam) %>% summarise(w_ijs=(n_stom*mean(pywgti)/n_tow_dec))



t1 %>% ggplot(aes(x=decade,y=w_ijs, color=pynam))+geom_bar(position="dodge", stat="identity")



 
t2 %>%
  filter(count>10, !analsci%in%c("","BLOWN", "EMPTY"))%>%
  mutate(analsci=forcats::fct_reorder(analsci, f_o, .desc = TRUE))%>%
  ggplot()+
  geom_bar(aes(x=analsci, y=f_o),stat = "identity")+
  theme(axis.text.x = element_text(angle = 90),
        panel.border= element_blank(),
        panel.background= element_rect(fill = "white"),
        axis.title.y=element_blank(),
        legend.position="none"
        )









 

```


```{r prey trial}
t0<- BSB_small %>% filter(!analsci%in%c("BLOWN"))

t1<- t0 %>% unite("ID",c(month, day, year,cruise6, pdlen, tow),remove =FALSE)

length(unique(t1$ID))

#generate counts of each prey catagory by year and subunit
t2<-t1 %>% group_by( pynam ) %>% summarise(count=n())

t2<-t2 %>% mutate(f_o=count/length(unique(t1$ID)))
#count each identified predator for each subunit and year
t3<-t1 %>%group_by( pynam )%>%  summarise(pd_count=length(unique(ID)))
#join the above counts to get frequency of occurance in stomachs by region and decade
t4<-dplyr::full_join( t2,t3, by= "analsci")

t4<-t4 %>% mutate(freq_o=count/pd_count)

#weighted mean 
 





t2 %>%
  filter(count>20,!pynam%in%c("EMPTY"))%>%
  mutate(analsci=forcats::fct_reorder(pynam, f_o, .desc = TRUE))%>%
  ggplot()+
  geom_bar(aes(x=pynam, y=f_o),stat = "identity")+
  theme(axis.text.x = element_text(angle = 90),
        panel.border= element_blank(),
        panel.background= element_rect(fill = "white"),
        axis.title.y=element_blank(),
        legend.position="none"
        )

```



```{r northern_raw, echo=FALSE}

# PYSPP	Prey species name.

#removing 




N<-BSB_small %>% filter(subunit =="Northern" , !gensci%in%c("EMPTY","","OTHER","BLOWN") & !is.na(pyamtw)) %>% group_by(gensci)%>%
  summarise(count=n(),freq_occ=(count/length(BSB_small$gensci)),weight=sum(pyamtw))


N_weight<-BSB_small %>%
  filter(subunit =="Northern"& !gensci%in%c("EMPTY","","OTHER")) %>%
  group_by(gensci)%>%
  summarise(weight=sum(pyamtw))


N_weight<- N_weight %>% filter(weight>0)%>%
  mutate(perc_weight=weight/sum(weight)*100)  

N_order<-N %>%  mutate(gensci=forcats::fct_reorder(gensci, (count)))


N1<-N %>%
  mutate(gensci=forcats::fct_reorder(gensci, (count))) %>%
  ggplot(aes(x=gensci,y=count, fill= gensci))+
  geom_bar(stat = "identity")+
  theme(axis.text.x = element_text(angle = 90),
        panel.border= element_blank(),
        panel.background= element_rect(fill = "white"),
        axis.title.y=element_blank(),
        legend.position="none"
        )+
   coord_flip()+
  xlab("Prey category")+
  ylab("count")+
  nmfspalette::scale_fill_nmfs(palette = "regional web")

N2<-N_weight %>%
  mutate(gensci=forcats::fct_reorder(gensci, desc(perc_weight))) %>%
  ggplot(aes(x=N_order$gensci,y=weight, fill=N_order$gensci))+geom_bar(stat = "identity")+
  theme(axis.text.x = element_text(angle = 90),
        panel.border= element_blank(),
        panel.background= element_rect(fill = "white"),
        axis.title.y=element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        legend.position="none"
        )+
  scale_y_reverse()+
  coord_flip()+
  ylab("Weight (g)")+
  nmfspalette::scale_fill_nmfs(palette = "regional web")


plot_N<-ggpubr::ggarrange(N2,N1,ncol=2)

ggpubr::annotate_figure(plot_N,top = ggpubr::text_grob("Northern area food habits", 
               color = "red", face = "bold", size = 14))



```

```{r southern_raw, echo=FALSE, message=FALSE, warning=FALSE}





S<-BSB_small %>% filter(subunit =="Southern" & !gensci%in%c("EMPTY","","OTHER") & !is.na(pyamtw)) %>% group_by(gensci) %>%
  summarise(count=n(),freq_occ=(count/length(BSB_small$gensci)),weight=sum(pyamtw))


S_weight<-BSB_small %>% 
  filter(subunit =="Southern"& !gensci%in%c("EMPTY","","OTHER")) %>%
  group_by(gensci) %>% 
  summarise(weight=sum(pyamtw))

S_weight<- S_weight %>%
  filter(weight>0)%>% 
  mutate(perc_weight=weight/sum(weight)*100)  

S_order<-S %>%  mutate(gensci=forcats::fct_reorder(gensci, (count)))


S1<-S %>%
  mutate(gensci=forcats::fct_reorder(gensci, (count))) %>%
  ggplot(aes(x=gensci,y=count, fill= gensci))+
  geom_bar(stat = "identity")+
  theme(axis.text.x = element_text(angle = 90),
        panel.border= element_blank(),
        panel.background= element_rect(fill = "white"),
        axis.title.y=element_blank(),
        legend.position="none"
        )+
   coord_flip()+
  xlab("Prey category")+
  ylab("count")+
  nmfspalette::scale_fill_nmfs(palette = "regional web")

S2<-S_weight %>%
  mutate(gensci=forcats::fct_reorder(gensci, desc(perc_weight))) %>%
  ggplot(aes(x=S_order$gensci,y=weight, fill=S_order$gensci))+geom_bar(stat = "identity")+
  theme(axis.text.x = element_text(angle = 90),
        panel.border= element_blank(),
        panel.background= element_rect(fill = "white"),
        axis.title.y=element_blank(),
        axis.text.y = element_blank(),
        axis.ticks.y = element_blank(),
        legend.position="none"
        )+
  scale_y_reverse()+
  coord_flip()+
  ylab("Weight (g)")+
  nmfspalette::scale_fill_nmfs(palette = "regional web")


plot_S<-ggpubr::ggarrange(S2,S1,ncol=2)

ggpubr::annotate_figure(plot_S,top = ggpubr::text_grob("Southern area food habits", 
               color = "red", face = "bold", size = 14))

```


```{r Frequency_occ, echo=FALSE, message=FALSE, warning=FALSE}
#removing 
t1<-BSB_small_ID %>% filter(year==1991)
t2<-BSB_small_ID %>% filter(year==1992)
t3<-BSB_small_ID %>% filter(year==2012)
  
t3 %>% filter(pdid==177479)
  # group_by(year) %>%
  # summarise(nstom=unique(pdid))
  
  
  t4<-t3 %>% group_by(pdid) %>% summarise(count=n())
  


BSB_small_ID %>%
  group_by(pynam) %>%
  summarise(count=n())%>%
  filter(count>10) %>%
  mutate(pynam=forcats::fct_reorder(pynam, desc(count))) %>%
  ggplot(aes(x=pynam, y=count))+geom_bar(stat = "identity")+
  theme(axis.text.x = element_text(angle = 90) )

```



```{r}



```




## NEAMAP

```{r NEAMAP_RAW, echo=FALSE, message=FALSE, warning=FALSE}
# total prey items 8227, 
NEAMAP_2022 %>%
  filter(Frequency>100)%>%
  mutate(Common_order=forcats::fct_reorder(COMMON, Frequency)) %>%
  ggplot(aes(x=Common_order,y=Frequency, fill= COMMON))+geom_bar(stat = "identity")+
  coord_flip()+
  theme(
        panel.border= element_blank(),
        panel.background= element_rect(fill = "white"),
        axis.title.y=element_blank(),
        legend.position="none"
        )+
  ggtitle("NEAMAP raw freq")

```

##CHESAMAP

```{r CHESAMAP_RAW, echo=FALSE, message=FALSE, warning=FALSE}

CHESAMAP_2022 %>%
  filter(Frequency>15)%>%
  mutate(Common_order=forcats::fct_reorder(COMMON, Frequency)) %>%
  ggplot(aes(x=Common_order,y=Frequency, fill= COMMON))+geom_bar(stat = "identity")+
  coord_flip()+
  theme(
        panel.border= element_blank(),
        panel.background= element_rect(fill = "white"),
        axis.title.y=element_blank(),
        legend.position="none"
        )+
  ggtitle("CHESAMAP raw freq")

```


```{r Diet diversity, echo=FALSE, message=FALSE, warning=FALSE}

BSB_small_ID %>% group_by()





```

