---
title: "CPUE"
author: "Ricky Tabandera"
date: "5/11/2021"
output: html_document
---

```{r cpue-setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(sf)
library(here)
library(mgcv)
```

## R Markdown

"Hilborn and Walters (1992)
identified two major situations that affect interpretation of mean CPUE as an index of stock
abundance: 1) hyperstability (when abundance declines faster than CPUE declines), and 2)
hyperdepletion (when abundance declines slower than CPUE declines). In general, a fishery with
a situation of hyperstability could be attributed to schooling behavior or changes in the species
spatial distribution rather than changes in abundance (van der Lee 2012). Hyperdepletion, on the
other hand, can occur when a more vulnerable portion of the population is easily caught,
followed by a more cryptic portion that avoids fishing mortality (van der Lee 2012).
Hyperstability can cause overfishing to go undetected, while hyperdepletion can result in
foregone yields when adopted management strategies such as catch limits are based on
inaccurate estimates of abundance (Erisman et al. 2011, van der Lee 2012). "


```{r cpue-data}

source(here::here("r","DDMMSS_to_DD.R"))



OBSPP<-rio::import(here::here("data","OBDBS" ,"OBSPP.rds"))

OBSPEC<-read.csv(here::here("data","OBDBS", "OBSPEC.csv"))

OB_key<-OBSPEC %>% select(SPECIES_ITIS,NESPP4 )%>% rename("ITISSPP" ="SPECIES_ITIS")

OB_key_join<- dplyr::left_join(OBSPP,OB_key, by= "NESPP4")

SP_group<-read.csv(here::here("data","species_groupings_V2.csv"))

SP_key<-SP_group  %>% select( COMNAME, SVSPP, ITISSPP, NESPP3, SCINAME) %>% mutate(COMNAME= stringr::str_to_sentence(COMNAME))

OB_sp_join<-dplyr::left_join(OB_key_join,SP_key, by= "ITISSPP")

OB_selected<-OB_sp_join %>% filter(COMNAME == "Black sea bass")




```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}

OB_selected %>% ggplot(aes(x=YEAR,y=HAILWT))+geom_bar(stat="identity")
OB_selected <- DDMMSS_to_DD(OB_selected, "LATHBEG","LONHBEG")
#remove rounded data that doesnt play well with conversion 
OB_selected <- OB_selected %>% filter(LON.DD>30)

t1<-data.frame("type"= seq(1:2),"lat"= c(400000,400001),"lon"=c(730118,730112))
t2<-DDMMSS_to_DD(t1,"lat","lon")

OB_selected %>% dplyr::group_by(YEAR) %>% dplyr::summarise(haul_per_yr=length(unique(LINK3)))%>% ggplot(aes(x=YEAR,y=haul_per_yr))+geom_bar(stat="identity")

library(sf)

OB_selected %>% ggplot(aes(x=LAT.DD))+geom_histogram()

ggplot(data=ecodata::coast)+
  geom_sf()+
  geom_point(aes(x=LON.DD,y=LAT.DD),data=OB_selected)


```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

```{r}
#R package mgcv using the argument ‘select’ equals ‘TRUE
OB_selected <- OB_selected %>% mutate("AREA"=as.factor(AREA), "WGTTYPE"=as.factor(WGTTYPE))


# make sure to shift the plot to reflect the actual range of values using seWithMean = TRUE, shift = coef(gam_model)[1]


gam_mod <- gam(HAILWT ~ s(YEAR)+s(MONTH)+te(LON.DD,LAT.DD)+WGTTYPE, data = OB_selected, method = "REML", select=TRUE)
coef(gam_mod)
summary(gam_mod)
plot(gam_mod, pages = 2,rug = TRUE,residuals = TRUE,
     pch = 1, cex = 1,shade = TRUE)
install.packages("mgcViz")
library(mgcViz)
t1<-mgcViz::getViz(gam_mod)
print(plot(t1), ask = FALSE)

```














