## Abundance

```{r, child = system.file("indicator_bookdown_template/_legend-child-doc.Rmd", package = "NEesp")}
```

Abundance data were pulled from ``r params$survey_source`` and ``r params$asmt_source``. 

### Survey abundance (raw measurements)

Separate geom_gls() functions were fit for fall and spring measurements; trend lines are only shown when the trend was statistically significant, so some plots may have fewer than two trend lines. Fall has solid trend lines, and spring has dashed trend lines. Please note, sometimes the survey observed a small number of fish outside of the defined stock area.


```{r, a-survey, fig.cap = paste(species, "survey abundance")}
NEesp::generate_plot(survey_data,
  variable = "ABUNDANCE",
  ytitle = "Abundance"
)
```


#### Risk {-}

See Methods for risk calculation details.
```{r}
indicators <- c(
  "abundance_spring",
  "abundance_fall"
)
```

```{r, child = system.file("indicator_bookdown_template/_risk-child-doc.Rmd", package = "NEesp")}
```

### Survey abundance (swept area estimates)

Please note, these estimates are not parsed by region Swept area estimates are based on spring and fall surveys only. The shaded gray region indicates +/- two standard errors.


```{r, abunance-survey-swept, fig.cap = paste(species, "swept area estimate from survey abundance")}
NEesp::plot_swept(swept_data, var = "abundance")
```


### Assessment abundance
```{r}
if (nrow(asmt_data %>% dplyr::filter(Metric == "Abundance")) > 0) {
  ndesc <- asmt_data %>%
    dplyr::filter(Metric == "Abundance") %>%
    # dplyr::mutate(options = paste(Description, Units)) %>%
    dplyr::select(Age) %>%
    unique() %>%
    nrow()
} else {
  ndesc <- 1
}
```


```{r, abun-asmt, fig.height = ndesc * 4, fig.cap = paste(species, "assessment abundance")}
NEesp::plot_asmt(asmt_data,
  metric = "Abundance",
  ytitle = "Abundance",
  lin = lines,
  col = colors
)
```


#### Risk {-}

See Methods for risk calculation details.
```{r}
indicators <- c("asmt_abundance")
```


```{r, child = system.file("indicator_bookdown_template/_risk-child-doc.Rmd", package = "NEesp")}
```


### Survey summary
```{r, abun_summary}
NEesp::generate_table(survey_data,
  variable = "ABUNDANCE",
  cap = "Measured abundance",
  type = params$file
)
```

### Data

#### Survey data (raw measurements) {-}
```{r, surv_abun_data}
data <- survey_data %>%
  NEesp::get_var_data(variable = "ABUNDANCE") %>%
  dplyr::mutate(variable = variable %>%
    format(big.mark = ",")) %>%
  NEesp::character_to_factor()

NEesp::make_html_table(data, 
                       type = params$file,
                       col_names = c("Year", "Season", "Region", "Abundance"))
```

#### Survey data (swept area estimates) {-}
```{r, swept_abun_data}
data <- swept_data %>%
  dplyr::select(YEAR, tot.abundance, tot.abund.SE) %>%
  dplyr::mutate(
    YEAR = as.numeric(YEAR),
    tot.abundance = tot.abundance %>%
      format(big.mark = ","),
    tot.abund.SE = tot.abund.SE %>%
      round(digits = 1) %>%
      format(big.mark = ",")
  )

NEesp::make_html_table(data, 
                       type = params$file,
                       col_names = c("Year", "Abundance", "Standard error"))
```

#### Assessment data {-}
```{r, asmt_abun_data}
data <- asmt_data %>%
  dplyr::filter(Metric == "Abundance") %>%
  dplyr::select(-Species, -Metric) %>%
  dplyr::mutate(Value = Value %>%
    round(digits = 0)) %>%
  dplyr::mutate(Value = Value %>%
    format(big.mark = ",")) %>%
  NEesp::character_to_factor()

NEesp::make_html_table(data, 
                       type = params$file)
```

