## Condition

Condition information comes from [diet data](https://github.com/Laurels1/Condition/blob/master/data/allfh.RData); only regions and seasons with more than 10 fish observations were considered. We calculated a rough condition factor as: Weight / Length^3, and relative weight was [previously calculated](https://github.com/Laurels1/Condition/tree/master/data).

### Length vs weight 

Please note, no trend lines were fit, points are jittered to reduce overlap, and sometimes the survey observed a small number of fish outside of the defined stock area.


```{r, lw, fig.cap = paste(species, "length vs weight")}
NEesp::plot_lw(diet_data)
```


### Condition factor: Weight-volume 

If there were more than 30 years of data, a geom_gls() regression was fit. In order to fit the geom_gls() regression, we calculated the mean condition factor for each year and plotted the geom_gls() through those points. Please note, points are jittered to reduce overlap, and sometimes the survey observed a small number of fish outside of the defined stock area.


```{r, condition, fig.cap = paste(species, "weight-volume condition")}
NEesp::plot_cond(diet_data)
```


#### Condition factor: Relative weight 

Please note, this data is aggregated by Ecological Protection Unit (EPU), which may differ slightly from the stock assessment regions.


```{r, laurel-cond, fig.cap = paste(species, "relative weight")}
NEesp::plot_relw(cond_data)
```


### Data

#### Length vs weight with weight-volume condition factor {-}
```{r, lw_data}
data <- diet_data %>%
  dplyr::filter(
    is.na(pdlen) == FALSE,
    is.na(pdwgt) == FALSE
  ) %>%
  dplyr::select(pdlen, pdwgt, season, Region, fish_id, year) %>%
  dplyr::distinct() %>% # remove duplicates
  dplyr::group_by(Region, season) %>%
  dplyr::mutate(n_fish = length(pdlen)) %>%
  dplyr::filter(n_fish > 10) %>% # only region-season with >10 fish)
  dplyr::mutate(condition = pdwgt / (pdlen^3)) %>%
  dplyr::mutate(condition = condition %>%
    round(digits = 4)) %>%
  dplyr::select(Region, year, season, pdlen, pdwgt, condition) %>%
  NEesp::character_to_factor()

NEesp::make_html_table(data,
                       type = params$file,
                       col_names = c(
  "Region", "Year", "Season", "Fish length (cm)",
  "Fish weight (g)", "Condition (g/cm^3)"
))
```

#### Relative weight condition factor {-}

Please note, this data is aggregated by Ecological Protection Unit (EPU), which may differ slightly from the stock assessment regions.
```{r, laurel_cond_data}
data <- cond_data %>%
  dplyr::select(-Species, -n) %>%
  dplyr::mutate(MeanCond = MeanCond %>%
    round(digits = 1)) %>%
  NEesp::character_to_factor()

NEesp::make_html_table(data,
                       type = params$file,
                       col_names = c(
  "EPU", "Fish sex", "Year", "Mean condition",
  "Number of fish"
))
```
