## von Bertalanffy growth curve

```{r data_wrangling}
selected.surv.clean <- NEesp::bio_survey %>%
  dplyr::mutate(SVSPP = as.numeric(SVSPP)) %>%
  dplyr::left_join(NEesp::species_key, by = "SVSPP") %>%
  dplyr::filter(Species == species) %>% 
  tidyr::drop_na(AGE)
```

```{r, vonb_model}
selected.surv.vonb <- vonb_model(selected.surv.clean)
```

### Length at age growth curve

The predicted von Bertalanffy growth curve for NMFS managed fish species. Growth parameters of `Linf` (Length infinty), `K` (growth coefficient), and `t0` (size at time 0) were estimated using non-linear least square model. The starting point for model building is accomplished using `FSA::vbStarts`. Age and length data sourced from `survdat` and spans all years and survey areas. 

`r if(species == "Monkfish"){"The age determination method for monkfish has not been validated, and the anatomic structure used has changed through time. In addition these stocks display a high degree of sexual dimorphisms making fitting of single growth curves unreliable.  This results a high degree of uncertainty in assessing the age structure of the stock and the effects of fishing pressure."}`


```{r single-growth-curve, fig.cap = paste(species, "length at age growth curve")}
plot_la(data = selected.surv.clean,
        vonb = selected.surv.vonb,
        species = species)
```
