## von Bertalanffy growth curve

```{r setup_vonb, cache = params$cache}
# install.packages("ggrepel")
# install.packages("ggpubr")
# devtools::install_github("james-thorson/FishLife")

suppressPackageStartupMessages(library(FishLife))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(nlstools))
library(rfishbase)
library(ggpubr)
library(FSA)
```

```{r data_wrangling, cache = params$cache, cache.lazy = FALSE}

#survdata <- readRDS(here::here("data", "survdat_pull_bio.rds"))
survdata <- bio_survey

survdata <- survdata %>% mutate(SVSPP = as.numeric(SVSPP))
stock_list_all_strata <- read.csv("https://raw.githubusercontent.com/NOAA-EDAB/ECSA/master/data/stock_list.csv")
stock_list_all_strata <- stock_list_all_strata %>% rename(SVSPP = svspp)
stock_list <- stock_list_all_strata %>% dplyr::distinct(SVSPP, .keep_all = T)

survdata.w.codes <- inner_join(survdata, stock_list, by = "SVSPP")

selected.surv <- survdata.w.codes %>%
  dplyr::mutate(common_name = common_name %>%
    stringr::str_to_sentence()) %>%
  dplyr::filter(common_name == species)

# only take records with age associated
selected.surv.clean <- selected.surv %>% drop_na(AGE)
```

```{r, vonb_model}
if (selected.surv.clean$AGE %>% unique() %>% length() >= 3) {
  # define the type of von b model
  vb <- vbFuns(param = "Typical")
  # define starting parameters based on the avalible data
  f.starts <- FSA::vbStarts(LENGTH ~ AGE, data = selected.surv.clean, methLinf = "Walford")

  # fit a non-linear least squares model based on the data and starting values
  f.fit <- nls(LENGTH ~ vb(AGE, Linf, K, t0), data = selected.surv.clean, start = f.starts)
  # store the fit parameters for later investigation
  f.fit.summary <- summary(f.fit, correlation = TRUE)

  # define the range of age values that will be used to generate points from the fitted model
  # roughly by 0.2 year steps
  newages <- data.frame(AGE = seq(0, 50, length = 250))
  # predict(f.fit,newdata=newages) this funtion uses the model from f.fit to generate new lengths
  # make a dataset with the values from the model
  selected.surv.vonb <- data.frame(AGE = seq(1, 50, length = 250), LENGTH = predict(f.fit, newdata = newages))
} else {
  print("NOT ENOUGH DATA TO FIT A CURVE")
}
```

### Length at age growth curve

The predicted von Bertalanffy growth curve for NMFS managed fish species. Growth parameters of `Linf` (Length infinty), `K` (growth coefficient), and `t0` (size at time 0) were estimated using non-linear least square model. The starting point for model building is accomplished using `FSA::vbStarts`. Age and length data sourced from `survdat` and spans all years and survey areas. 

`r if(params$species == "Monkfish"){"The age determination method for monkfish has not been validated, and the anatomic structure used has changed through time. In addition these stocks display a high degree of sexual dimorphisms making fitting of single growth curves unreliable.  This results a high degree of uncertainty in assessing the age structure of the stock and the effects of fishing pressure."}`


```{r single-growth-curve, fig.cap = paste(species, "length at age growth curve")}
# palette
if (nrow(selected.surv.clean) > 0) {
  fig <- ggplot(
    data = selected.surv.clean,
    aes(
      x = AGE,
      y = LENGTH,
      color = YEAR %>% as.numeric()
    )
  ) +
    geom_jitter(alpha = 0.5) +
    scale_color_gradientn(
      colors = nmfspalette::nmfs_palette("regional web")(4),
      name = "Year"
    ) +
    xlim(0, (1.5 * max(selected.surv.clean$AGE))) +
    ylim(0, (1.5 * max(selected.surv.clean$LENGTH, na.rm = TRUE))) +
    xlab("Age (jittered)") +
    ylab(" Total length (cm) (jittered)") +
    ggtitle(species, subtitle = "Length at age") +
    theme_minimal()

  if (nrow(selected.surv.vonb) > 0) {
    fig <- fig +
      geom_line(
        data = selected.surv.vonb,
        inherit.aes = FALSE,
        mapping = aes(
          x = AGE,
          y = LENGTH
        ),
        color = "blue",
        size = 1.4
      )
  }

  return(fig)
} else {
  print("NO DATA")
}
```
