---
title: "assessmentdata_testing"
author: "Abigail Tyrell"
date: "`r format(Sys.time(), '%d %b %Y')`"
output: html_document
params: 
  species_ID: "Acadian redfish - Gulf of Maine / Georges Bank"
  data_source: data
---

```{r, setup, include = FALSE}
knitr::opts_chunk$set(echo = F, message = F, warning = F, fig.align = 'center')
```

```{r, functions, include = FALSE}
library(ggplot2)

`%>%` <- dplyr::`%>%`

plot_bbmsy <- function(x, ytitle) {

  # use geom_gls only if there are 30 or more years of data
  if(length(x$`B Year`) >= 30){
    fig <- ggplot(x,
         aes(x = `B Year`,
             y = `B/Bmsy`))+
    geom_point()+
    ecodata::geom_gls()+
    theme_bw()+
    xlab("Year")+
    ylab(ytitle)
  } else {
    fig <- ggplot(x,
         aes(x = `B Year`,
             y = `B/Bmsy`))+
    geom_point()+
    theme_bw()+
    xlab("Year")+
    ylab(ytitle)
  }

  return(fig)
}

plot_ffmsy <- function(x, ytitle) {

  # use geom_gls only if there are 30 or more years of data
  if(length(x$`B Year`) >= 30){
    fig <- ggplot(x,
         aes(x = `F Year`,
             y = `F/Fmsy`))+
    geom_point()+
    ecodata::geom_gls()+
    theme_bw()+
    xlab("Year")+
    ylab(ytitle)
  } else {
    fig <- ggplot(x,
         aes(x = `F Year`,
             y = `F/Fmsy`))+
    geom_point()+
    theme_bw()+
    xlab("Year")+
    ylab(ytitle)
  }

  return(fig)
}


```

```{r, data, include = FALSE}
species <- params$species_ID
data <- params$data_source
plot_data <- data[data$`Stock Name` == species,]

# TO-DO
## find average from past 5 years and compare to long-term average?
## or some other comparison metric
```

## Test of assesssmentdata data, species stock name `r species`

This is a test report of `assessmentdata` data. 

This report is pulling information on the species with the stock name: `r species`.
B/Bmsy and F/Fmsy were plotted against time. If there were 30 or more years of data, a geom_gls regression line was fit (yellow = significant increase; purple = significant decrease; no line = no significant trend). If there were fewer than 30 years of data, no regression was fit. 

```{r status}
b_status <- if((plot_data$'B/Bmsy' %>% tail(n = 1)) > 1) "DANGER" else "GOOD"
f_status <- if((plot_data$'F/Fmsy' %>% tail(n = 1)) > 1) "DANGER" else "GOOD"
```

### B/Bmsy
The most recent status of B/Bmsy is: `r b_status`
```{r, bbmsy} 

plot_bbmsy(plot_data, ytitle = "B / Bmsy")
```

### F/Fmsy
The most recent status of F/Fmsy is: `r f_status`
```{r, ffmsy}
plot_ffmsy(plot_data, ytitle = "F / Fmsy")
```

